---
title: "Exploratory: Group 1 Specificity Tests"
author: "Solomon Johnson"
date: "`r Sys.Date()`"
output-file: "phase2_exploratory_specificity.html"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  cache: false
  freeze: auto
---

<!--
============================================================
EXPLORATORY ANALYSIS — NOT PART OF VERSIONED PIPELINE
Two scratch tests run against v1.7 cohort (no re-extraction):

  TEST 1 — Strict anatomic definition
           Restricts wound codes to S01.1/S01.3/S01.4/S01.5/S01.81
           (anatomically concordant with facial nerve territory),
           drops S01.0 (scalp), S01.2 (nose), S01.9 (unspecified),
           and all S11.* (neck). Restricts to 7th char A only
           (drops 7th char D subsequent encounters). Excludes
           parotid/cranial nerve neoplasm + surgical complication.

  TEST 2 — Neoplasm/iatrogenic exclusion only
           Keeps the v1.7 primary definition intact, only removes
           encounters carrying parotid/cranial-nerve neoplasm
           (C07, D11.0, C49.0, D33.3, C72.4) or surgical complication
           (T85.4*, Y83.*) codes. Tests whether removing the
           single largest false-positive category alone changes
           the headline numbers.

This file is exploratory. Findings inform a future formal version
bump (e.g. v2.0 with strict primary, or addition of Sensitivity
Analysis 4 to the main pipeline) but do not modify any existing
output.
============================================================
-->

```{r setup}
#| cache: false
library(haven);     library(dplyr);     library(purrr);    library(furrr)
library(future);    library(tidyr);     library(stringr);  library(survey)
library(knitr);     library(kableExtra);library(gt)

# ── File paths — identical to v1.7 ───────────────────────────────────────────
d16_23 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/NASS_HCUP Database/RAW Data Files HNASS"
d21 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2021"
d22 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2022"
d23 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2023"

enc_files <- list(
  "2016"=file.path(d16_23,"NASS_2016_Encounter-002.dta"),
  "2017"=file.path(d16_23,"NASS_2017_Encounter-003.dta"),
  "2018"=file.path(d16_23,"NASS_2018_Encounter-001.dta"),
  "2019"=file.path(d16_23,"NASS_2019_Encounter-005.dta"),
  "2020"=file.path(d16_23,"NASS_2020_Encounter.dta"),
  "2021"=file.path(d21,   "NASS_2021_Encounter.dta"),
  "2022"=file.path(d22,   "NASS_2022_Encounter.dta"),
  "2023"=file.path(d23,   "NASS_2023_Encounter.dta")
)
hosp_files <- list(
  "2016"=file.path(d16_23,"NASS_2016_Hospital.dta"),
  "2017"=file.path(d16_23,"NASS_2017_Hospital.dta"),
  "2018"=file.path(d16_23,"NASS_2018_Hospital.dta"),
  "2019"=file.path(d16_23,"NASS_2019_Hospital.dta"),
  "2020"=file.path(d16_23,"NASS_2020_Hospital.dta"),
  "2021"=file.path(d21,   "NASS_2021_Hospital.dta"),
  "2022"=file.path(d22,   "NASS_2022_Hospital.dta"),
  "2023"=file.path(d23,   "NASS_2023_Hospital.dta")
)

# ── Code constants — identical to v1.7 ───────────────────────────────────────
cpt_codes         <- c("64864","64865","64885","64886")
dx_stem           <- "S045"
s01_stem          <- "S01"
s11_stem          <- "S11"
ext_cause_pattern <- "^(W2[5-7]|W45|X99|Y28|V[0-9]{2})"
not_sequela       <- function(cd) !str_ends(cd, "S")

# ── Extraction function — v1.7 logic + dx_raw carried forward ────────────────
extract_year_expl <- function(year, enc_fp, hosp_fp) {
  message("  Loading: ", year)
  if (!file.exists(enc_fp))  { warning("Missing: ", enc_fp);  return(NULL) }
  enc  <- read_dta(enc_fp)
  ec   <- names(enc)
  get_col <- function(pat) grep(pat, ec, value=TRUE, ignore.case=TRUE)[1]
  dx_cols   <- grep("^i10_dx",   ec, value=TRUE, ignore.case=TRUE)
  cpt_cols  <- grep("^cpt[0-9]", ec, value=TRUE, ignore.case=TRUE)
  discwt_col <- get_col("^discwt$")
  hosp_id_e  <- get_col("^hosp_nass$")

  # v1.7 flags
  dx_s045  <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, dx_stem) & not_sequela(cd) })) > 0
  s01_flag <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, s01_stem) & not_sequela(cd) })) > 0
  s11_flag <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, s11_stem) & not_sequela(cd) })) > 0
  ext_flag <- rowSums(sapply(dx_cols, function(col)
    str_detect(trimws(as.character(enc[[col]])), ext_cause_pattern))) > 0
  cpt_flag <- rowSums(sapply(cpt_cols, function(col)
    trimws(as.character(enc[[col]])) %in% cpt_codes)) > 0

  crit_a <- dx_s045 & (s01_flag | s11_flag)
  crit_b <- cpt_flag & (s01_flag | s11_flag)
  g1     <- crit_a | crit_b
  g2     <- cpt_flag & g1
  keep   <- g1
  if (sum(keep) == 0) return(NULL)

  # Carry raw dx codes for post-hoc specificity tests
  clean_codes <- function(mat) {
    apply(mat, 1, function(row) {
      codes <- trimws(as.character(row))
      codes <- codes[nzchar(codes) & codes != "." & !is.na(codes) & codes != "0"]
      paste(codes, collapse="|")
    })
  }
  dx_raw_col <- clean_codes(enc[keep, dx_cols, drop=FALSE])

  tibble(
    year       = year,
    had_repair = as.integer(g2[keep]),
    discwt     = if (!is.na(discwt_col))
                   as.numeric(enc[[discwt_col]][keep]) else rep(1, sum(keep)),
    hosp_id    = if (!is.na(hosp_id_e))
                   as.character(enc[[hosp_id_e]][keep])
                 else as.character(seq_len(sum(keep))),
    flag_crit_a = crit_a[keep],
    flag_crit_b = crit_b[keep],
    flag_extcause = ext_flag[keep],
    dx_raw     = dx_raw_col
  )
}

# ── Run extraction (parallel) ─────────────────────────────────────────────────
message("Starting extraction for exploratory specificity tests...")
plan(multisession, workers = max(1, availableCores()-1))
enc_all <- bind_rows(Filter(Negate(is.null),
  future_map2(names(enc_files), enc_files,
    ~ extract_year_expl(.x, .y, hosp_files[[.x]]),
    .options = furrr_options(seed=TRUE))))
plan(sequential)

message("enc_all loaded: ", nrow(enc_all), " rows")

# ── Reference numbers from v1.7 primary ──────────────────────────────────────
primary_n_uw  <- nrow(enc_all)
primary_n_w   <- round(sum(enc_all$discwt, na.rm=TRUE))
primary_g2_uw <- sum(enc_all$had_repair == 1)
primary_g2_w  <- round(sum(enc_all$discwt[enc_all$had_repair==1], na.rm=TRUE))
primary_rate  <- round(100 * primary_g2_w / primary_n_w, 1)

cat("v1.7 primary cohort: ", format(primary_n_uw, big.mark=","),
    " unweighted, ", format(primary_n_w, big.mark=","), " weighted\n",
    "Group 2: ", primary_g2_uw, " unweighted (",
    primary_rate, "% repair rate)\n", sep="")
```

# Test 1 — Strict Anatomic Definition

```{r test1-strict-anatomic}
#| cache: false

# ── Helper: parse pipe-joined dx_raw into per-row code list ───────────────────
parse_codes <- function(s) strsplit(s, "\\|")[[1]]

# ── Strict-anatomic flags ─────────────────────────────────────────────────────
# Anatomically concordant wound codes only (drop S01.0 scalp, S01.2 nose,
# S01.9 unspecified, all S11.* neck)
anatomic_wound_re <- "^S01[13458]"   # matches S011, S013, S014, S015, S018

# 7th character A only (drop D subsequent encounters)
not_acute <- function(cd) !str_ends(cd, "A")

is_strict_acute_s045 <- function(codes) {
  any(str_starts(codes, "S045") & str_ends(codes, "A"))
}
is_strict_anatomic_wound <- function(codes) {
  any(str_detect(codes, anatomic_wound_re) & str_ends(codes, "A"))
}

# ── Neoplasm and iatrogenic exclusion regex ──────────────────────────────────
# C07*    parotid malignant neoplasm
# D11.0   benign parotid neoplasm
# C49.0*  connective tissue head/face/neck
# D33.3   benign cranial nerve neoplasm
# C72.4*  malignant cranial nerve
# T85.4*  mechanical complication of prosthetic device
# Y83*    surgical operation as cause of abnormal reaction
exclusion_re <- "^(C07|D110|C490|D333|C724|T854|Y83)"

is_excluded <- function(codes) {
  any(str_detect(codes, exclusion_re))
}

# ── Apply tests row-wise ──────────────────────────────────────────────────────
message("Parsing dx_raw across ", nrow(enc_all), " rows...")
code_lists <- lapply(enc_all$dx_raw, parse_codes)

flag_strict_s045  <- vapply(code_lists, is_strict_acute_s045,    logical(1))
flag_strict_wound <- vapply(code_lists, is_strict_anatomic_wound, logical(1))
flag_excluded     <- vapply(code_lists, is_excluded,             logical(1))

# Strict Criterion A: acute S04.5 + anatomic wound
strict_crit_a <- flag_strict_s045 & flag_strict_wound

# Strict Criterion B: had repair CPT + anatomic wound
strict_crit_b <- (enc_all$had_repair == 1) & flag_strict_wound

# Strict Group 1 = (A or B) AND not excluded
strict_g1 <- (strict_crit_a | strict_crit_b) & !flag_excluded

# Test 1 cohort
t1 <- enc_all[strict_g1, ]
t1_n_uw  <- nrow(t1)
t1_n_w   <- round(sum(t1$discwt, na.rm=TRUE))
t1_g2_uw <- sum(t1$had_repair == 1)
t1_g2_w  <- round(sum(t1$discwt[t1$had_repair==1], na.rm=TRUE))
t1_rate  <- if (t1_n_w > 0) round(100 * t1_g2_w / t1_n_w, 1) else NA_real_

# ── Loss accounting ───────────────────────────────────────────────────────────
loss_to_anatomic_wound <- sum(!flag_strict_wound)
loss_to_acute_only     <- sum(flag_strict_wound & !flag_strict_s045 &
                               enc_all$had_repair == 0)
loss_to_exclusion      <- sum((strict_crit_a | strict_crit_b) & flag_excluded)

cat("\n=== TEST 1 — Strict Anatomic Definition ===\n",
    "Cohort: ", format(t1_n_uw, big.mark=","), " unweighted, ",
                format(t1_n_w,  big.mark=","), " weighted\n",
    "Group 2: ", t1_g2_uw, " unweighted (", t1_rate, "% repair rate)\n",
    "Excluded by neoplasm/iatrogenic codes: ", loss_to_exclusion, " rows\n",
    "Excluded by anatomic-wound restriction: ",
        primary_n_uw - sum(strict_crit_a | strict_crit_b), " rows\n", sep="")

# ── Side-by-side summary table ────────────────────────────────────────────────
test1_summary <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Repair rate (weighted)",
    "Cohort \u0394 vs primary (unweighted)",
    "Excluded by anatomic-wound restriction",
    "Excluded by neoplasm/iatrogenic codes"
  ),
  `v1.7 Primary` = c(
    format(primary_n_uw, big.mark=","),
    format(primary_n_w,  big.mark=","),
    format(primary_g2_uw,big.mark=","),
    format(primary_g2_w, big.mark=","),
    paste0(primary_rate, "%"),
    "(reference)",
    "(reference)",
    "(reference)"
  ),
  `Test 1: Strict` = c(
    format(t1_n_uw, big.mark=","),
    format(t1_n_w,  big.mark=","),
    format(t1_g2_uw,big.mark=","),
    format(t1_g2_w, big.mark=","),
    paste0(t1_rate, "%"),
    paste0(sprintf("%+d", t1_n_uw - primary_n_uw),
           " (", sprintf("%+.1f", 100*(t1_n_uw-primary_n_uw)/primary_n_uw),
           "%)"),
    paste0(primary_n_uw - sum(strict_crit_a | strict_crit_b), " rows"),
    paste0(loss_to_exclusion, " rows")
  )
)

test1_summary %>%
  kable(caption = paste0(
    "Test 1: Strict anatomic-wound definition + neoplasm/iatrogenic exclusion. ",
    "Wound codes restricted to S01.1, S01.3, S01.4, S01.5, S01.81 ",
    "(facial-nerve territory) with 7th char A only. Exclusion codes: ",
    "C07, D11.0, C49.0, D33.3, C72.4 (parotid/cranial nerve neoplasm); ",
    "T85.4*, Y83.* (surgical complication).")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

# Test 2 — Neoplasm/Iatrogenic Exclusion Only

```{r test2-exclusion-only}
#| cache: false
# Same v1.7 definition, only remove encounters carrying parotid/cranial-nerve
# neoplasm or surgical-complication codes.

t2_keep <- !flag_excluded
t2 <- enc_all[t2_keep, ]
t2_n_uw  <- nrow(t2)
t2_n_w   <- round(sum(t2$discwt, na.rm=TRUE))
t2_g2_uw <- sum(t2$had_repair == 1)
t2_g2_w  <- round(sum(t2$discwt[t2$had_repair==1], na.rm=TRUE))
t2_rate  <- if (t2_n_w > 0) round(100 * t2_g2_w / t2_n_w, 1) else NA_real_

# Breakdown of which exclusion codes hit
hits <- list(
  `C07 parotid malignant`   = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "C07")))),
  `D11.0 parotid benign`     = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "D110")))),
  `C49.0 connective tissue head/neck` = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "C490")))),
  `D33.3 cranial nerve benign neoplasm` = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "D333")))),
  `C72.4 cranial nerve malignant`       = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "C724")))),
  `T85.4* surgical mesh complication`   = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "T854")))),
  `Y83.* surgical complication`          = sum(sapply(code_lists,
    function(cs) any(str_starts(cs, "Y83"))))
)

cat("\n=== TEST 2 — Neoplasm/Iatrogenic Exclusion Only ===\n",
    "Cohort: ", format(t2_n_uw, big.mark=","), " unweighted, ",
                format(t2_n_w,  big.mark=","), " weighted\n",
    "Group 2: ", t2_g2_uw, " unweighted (", t2_rate, "% repair rate)\n",
    "Total rows excluded: ", primary_n_uw - t2_n_uw, "\n", sep="")

cat("\nHits per exclusion code (rows can match multiple):\n")
for (nm in names(hits)) cat("  ", nm, ": ", hits[[nm]], "\n", sep="")

test2_summary <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Repair rate (weighted)",
    "Cohort \u0394 vs primary (unweighted)",
    "Total rows excluded"
  ),
  `v1.7 Primary` = c(
    format(primary_n_uw, big.mark=","),
    format(primary_n_w,  big.mark=","),
    format(primary_g2_uw,big.mark=","),
    format(primary_g2_w, big.mark=","),
    paste0(primary_rate, "%"),
    "(reference)",
    "(reference)"
  ),
  `Test 2: Exclusion-only` = c(
    format(t2_n_uw, big.mark=","),
    format(t2_n_w,  big.mark=","),
    format(t2_g2_uw,big.mark=","),
    format(t2_g2_w, big.mark=","),
    paste0(t2_rate, "%"),
    paste0(sprintf("%+d", t2_n_uw - primary_n_uw),
           " (", sprintf("%+.1f", 100*(t2_n_uw-primary_n_uw)/primary_n_uw),
           "%)"),
    paste0(primary_n_uw - t2_n_uw, " rows")
  )
)

test2_summary %>%
  kable(caption = paste0(
    "Test 2: v1.7 primary definition with parotid/cranial-nerve neoplasm and ",
    "surgical-complication exclusions only. Anatomic wound restriction NOT applied. ",
    "Tests whether removing the largest false-positive category alone moves ",
    "the headline numbers.")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)

# Per-code hits table
hits_tbl <- tibble(
  `Exclusion code`           = names(hits),
  `Rows hit (may overlap)`   = unlist(hits)
)
hits_tbl %>%
  kable(caption = "Test 2: per-code hit counts. Rows may match more than one exclusion code.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

# Combined Comparison

```{r combined-summary}
#| cache: false
combined <- tibble(
  Metric = c("Group 1 unweighted N",
             "Group 1 weighted N",
             "Group 2 unweighted N",
             "Group 2 weighted N",
             "Repair rate (weighted)"),
  `v1.7 Primary` = c(format(primary_n_uw, big.mark=","),
                     format(primary_n_w,  big.mark=","),
                     format(primary_g2_uw,big.mark=","),
                     format(primary_g2_w, big.mark=","),
                     paste0(primary_rate, "%")),
  `Test 1: Strict` = c(format(t1_n_uw, big.mark=","),
                       format(t1_n_w,  big.mark=","),
                       format(t1_g2_uw,big.mark=","),
                       format(t1_g2_w, big.mark=","),
                       paste0(t1_rate, "%")),
  `Test 2: Exclusion only` = c(format(t2_n_uw, big.mark=","),
                                format(t2_n_w,  big.mark=","),
                                format(t2_g2_uw,big.mark=","),
                                format(t2_g2_w, big.mark=","),
                                paste0(t2_rate, "%"))
)

combined %>%
  kable(caption = "Side-by-side comparison: v1.7 primary vs Test 1 (strict) vs Test 2 (exclusion-only).") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(5, bold = TRUE, background = "#EBF4FB")
```

# Interpretation Notes

```{r interpretation-notes}
#| cache: false
#| echo: false

cat("Interpretation guide:\n\n")

# Test 1 interpretation
delta_t1   <- t1_n_uw - primary_n_uw
rate_t1    <- t1_rate - primary_rate
cat("TEST 1 (strict anatomic): cohort change = ", sprintf("%+d", delta_t1),
    " unweighted (", sprintf("%+.1f", 100*delta_t1/primary_n_uw), "%); ",
    "repair rate change = ", sprintf("%+.1f", rate_t1), " percentage points.\n",
    sep="")

if (abs(rate_t1) < 3) {
  cat("  → Repair rate moved <3 pp. Strict definition produces similar rate.\n",
      "    v1.7 primary is robust to anatomic restriction.\n", sep="")
} else if (rate_t1 > 0) {
  cat("  → Repair rate increased substantially under strict definition.\n",
      "    Cases dropped by anatomic restriction were enriched for non-repair.\n",
      "    Suggests S01.0/S01.9/S11.* (less anatomic-specific wounds) carry\n",
      "    more conservatively-managed cases.\n", sep="")
} else {
  cat("  → Repair rate decreased substantially under strict definition.\n",
      "    Investigate; this is unexpected.\n", sep="")
}

cat("\n")

# Test 2 interpretation
delta_t2   <- t2_n_uw - primary_n_uw
rate_t2    <- t2_rate - primary_rate
cat("TEST 2 (exclusion only): cohort change = ", sprintf("%+d", delta_t2),
    " unweighted (", sprintf("%+.1f", 100*delta_t2/primary_n_uw), "%); ",
    "repair rate change = ", sprintf("%+.1f", rate_t2), " percentage points.\n",
    sep="")

if (abs(delta_t2) < 10) {
  cat("  → Few rows hit the exclusion criteria. Neoplasm/iatrogenic\n",
      "    contamination of the v1.7 cohort is minimal.\n", sep="")
} else {
  cat("  → ", abs(delta_t2), " rows excluded. Worth promoting to a\n",
      "    permanent exclusion in a future v2.0.\n", sep="")
}
```

# Decision Framework

The numbers above feed three possible next steps:

1. **If Test 1 produces results within 2-3 percentage points of v1.7** for both cohort size proportionally and repair rate, the current definition is robust to anatomic tightening — no change needed; document as a sensitivity in methods.

2. **If Test 1 differs meaningfully but Test 2 alone produces no change** — the anatomic-wound restriction is doing the work, not the neoplasm/iatrogenic exclusion. Decide whether to elevate strict definition to primary in v2.0 or report as a permanent Sensitivity Analysis 4.

3. **If Test 2 alone meaningfully shifts numbers** — the v1.7 cohort has substantive neoplasm/iatrogenic contamination. Add the exclusion codes to the extraction function and re-extract for v2.0.
