---
title: "Traumatic Facial Nerve Laceration: HCUP NASS 2016–2023 (Phase 2)"
author: "Solomon Johnson"
date: "`r Sys.Date()`"
output-file: "nass_phase2_analysis_2_0.html"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
  freeze: auto
---

<!--
============================================================
CHANGELOG
v2.0 — 2026-04-30 — NIS chunk redesigned to align with the research
       question (acute traumatic facial nerve laceration and repair):
       (1) NIS Group 1 now defined as (A OR B):
           A. S04.5* (7th char A or D) co-occurring with
              S01.* or S11.* (7th char A or D)
           B. Facial nerve repair PCS code (01QM*, 00QM* miscoded,
              01UM07Z/0KZ/0JZ grafting) co-occurring with
              S01.* or S11.* (7th char A or D)
           Grafting codes (01UM07Z/0KZ/0JZ) included for forward
           compatibility with future NIS years even though 2023
           returns zero of them.
       (2) NEW chunk: nis-mechanism — applies the same V/W/X/Y
           hierarchical mechanism-of-injury classification used in
           the NASS chunk to the NIS Group 1 cohort.
       (3) any_match_start fixed to handle NA values in NIS DX
           columns (was returning NA across all DX flags).
v1.9 — 2026-04-30 — Three changes:
       (1) Fixed NIS parallel cohort error — flags were referenced
           outside the survey design data; now added as integer
           columns to nis before svydesign so svytotal can resolve
           them. Wrapped svytotal calls with as.numeric() to coerce
           the named coefficient vector to scalar.
       (2) Data paths migrated from OneDrive to local Desktop alias
           (~/Desktop/Otolaryngology - Head and Neck Surgery alias/
           HCUP Data/). Substantially faster reads vs OneDrive
           CloudStorage. Output paths remain on OneDrive.
       (3) Explicit multicore optimization: set options(mc.cores)
           and use future plan(multisession) with
           parallel::detectCores() throughout. Cache MUST be
           cleared before re-rendering due to path changes.
v1.8 — 2026-04-30 — Three additions:
       (1) Mechanism-of-injury descriptive analysis using V/W/X/Y
           external cause codes within Group 1. New chunk
           cooccur-mechanism produces an Excel sheet and gt Word
           table classifying encounters by transport, assault,
           cutting, fall, animal, self-harm, undetermined, or
           unspecified mechanism.
       (2) Sheet 4 footnote updated to note that the apparent
           year trend (p=0.016) likely reflects ICD-10-CM coding
           maturation rather than a true practice change.
       (3) Neoplasm/iatrogenic exclusion added to extract_year:
           encounters carrying C07, D11.0, C49.0, D33.3, C72.4,
           T85.4*, or Y83.* codes are dropped from Group 1.
           Cache MUST be cleared before re-rendering.
       (4) New NIS exploratory chunk (cache: false) computes a
           parallel inpatient cohort using the same v1.7 inclusion
           logic adapted for ICD-10-PCS: includes both PNS-correct
           (01QM) and CNS-miscoded (00QM) procedure codes since
           v1.1 confirmed both are real facial nerve repairs.
v1.7 — 2026-04-28 — Refine NIS 2023 sheet using completed three-phase
       NIS analysis (v1.0 frequency, v1.1 miscoding confirmation,
       v1.2 trauma classification). Total inpatient facial nerve
       repair burden = 285 weighted; the TRAUMATIC subset is 26
       unweighted / 130 weighted (45.6% of all). Non-traumatic
       remainder = 31 / 155. Miscoding hypothesis confirmed at
       75.9% (threshold 70%). Adds traumatic-vs-nontraumatic split
       and cell-suppression compliance language. Footnote elevated
       from "data-quality observation" to "confirmed novel finding."
v1.6 — 2026-04-27 — Update NIS 2023 sheet to reflect expanded inpatient query
       results. Now includes 10 ICD-10-PCS codes (3 PNS Repair, 3 Supplement,
       1 Reposition, 3 CNS Repair). Total rises from 3/15 to 57/285. Adds
       novel data-quality finding: 94.7% of inpatient facial nerve repairs
       are coded to the CNS section (00QM) rather than the correct PNS
       section (01QM). Footnote updated across HTML, Excel, and gt outputs
       to document this miscoding pattern. NASS-derived analyses unchanged;
       no cache clear required (only NIS chunks affected, all cache: false).
v1.5 — 2026-04-27 — Tighten Group 1 definition for greater specificity.
       (1) Apply 7th-character filter (A/D only, exclude S sequela)
           to S01.* and S11.* in addition to S04.5*. All wound codes
           must reflect acute mechanism, not late effects.
       (2) Drop external cause from Criterion A — Group 1 now requires
           S01/S11 wound code (Criterion A) or CPT + S01/S11 (Criterion B).
           External cause codes remain available for Sensitivity Analysis 2
           and 3 as optional confirmation, not primary inclusion.
       Cache must be cleared before re-rendering.
v1.4 — 2026-04-27 — Exclude S04.5* sequela encounters (7th character S)
       from Criterion A. Sequela codes represent late effects of prior
       injury, not acute traumatic lacerations, and should not count
       toward the study denominator. Subsequent encounters (7th char D)
       retained — late nerve repairs in NASS may legitimately carry D.
       7th-character distribution table added to co-occurrence output
       for transparency. Cache must be cleared before re-rendering.
v1.3 — 2026-04-27 — Add co-occurrence exploratory analysis for Group 2.
       Extraction now carries dx_raw and cpt_raw (pipe-joined code strings)
       on every enc_all row. New cooccur-explore chunk produces: top 40
       co-occurring ICD-10-CM codes, top 30 co-occurring CPT codes,
       S04.5* subcode breakdown, and open wound subcode breakdown —
       all in Group 2 encounters. Results in HTML and Excel sheet
       "Co-occurrence (Exploratory)". Cache must be cleared.
v1.2 — 2026-04-27 — Expand Group 2 CPT codes to include nerve graft
       codes 64885 and 64886 alongside 64864 and 64865. All footnotes,
       figure labels, and draft methods text updated. copy-to-drive
       self-copy bug fixed. Cache must be cleared before re-rendering.
v1.1 — 2026-04-27 — Add Sensitivity Analysis 3 (maximum specificity).
       Adds laterality-specific nerve flags (S04.51*, S04.52*) and
       anatomic wound flags (S01.1*, S01.3*, S01.4*, S01.5*, S01.81*)
       to extraction. SA3 definition: laterality-specified nerve injury
       + anatomically concordant wound + external cause code.
       Cache must be cleared before re-rendering (extraction changed).
v1.0 — 2026-04-27 — Initial Phase 2 analysis. New Group 1
       definition (Criterion A: S04.5 + open wound or external
       cause; Criterion B: CPT + open wound). Adds two
       sensitivity analyses, Phase 1 vs Phase 2 comparison
       sheet, and an automated draft methods paragraph.
       Output redirected from Desktop to OneDrive Phase 2
       folder with subfolder routing (Figures, Excel, Word
       Tables, Quarto Versions).
============================================================
-->

```{r setup}
#| cache: false
pkgs <- c("haven","dplyr","purrr","furrr","future","readr","tidyr",
          "stringr","survey","openxlsx","ggplot2","scales","patchwork",
          "ggrepel","DiagrammeR","DiagrammeRsvg","rsvg",
          "WeightIt","cobalt","knitr","kableExtra","maps","mapproj",
          "gt","officer","flextable")
for (p in pkgs) if (!requireNamespace(p, quietly=TRUE)) install.packages(p)

library(haven);     library(dplyr);     library(purrr);    library(furrr)
library(future);    library(readr);     library(tidyr);    library(stringr)
library(survey);    library(openxlsx);  library(ggplot2);  library(scales)
library(patchwork); library(ggrepel);   library(WeightIt); library(cobalt)
library(knitr);     library(kableExtra);library(maps);     library(mapproj)
library(DiagrammeR);library(DiagrammeRsvg);library(rsvg)
library(gt);

# ── Multicore / parallelization configuration ──────────────────────────────
# Use all available cores minus 1 by default. parallel::detectCores() returns
# the total number of logical CPUs; we leave one for the OS and other tasks.
n_cores <- max(1L, parallel::detectCores() - 1L)
options(mc.cores = n_cores)
message("Multicore configured: ", n_cores, " of ",
        parallel::detectCores(), " cores will be used")

# Phase 2 output directory — OneDrive only, no Desktop output
output_dir <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Analysis/Phase 2"

if (!dir.exists(path.expand(output_dir))) {
  stop("Phase 2 output directory not found: ", output_dir)
}

fig_dir   <- file.path(path.expand(output_dir), "Figures")
xlsx_dir  <- file.path(path.expand(output_dir), "Excel")
word_dir  <- file.path(path.expand(output_dir), "Word Tables")
qmd_dir   <- file.path(path.expand(output_dir), "Quarto Versions")

for (d in c(fig_dir, xlsx_dir, word_dir, qmd_dir)) {
  if (!dir.exists(d)) {
    stop("Required Phase 2 subfolder not found: ", d)
  }
}

message("Phase 2 output_dir confirmed: ", output_dir)
```

# Configuration

```{r config}
# ── Data location ────────────────────────────────────────────────────────────
# v1.9: data files migrated to local Desktop alias for faster reads.
# OneDrive paths retained as fallback in case local files are unavailable.
local_data_root <- path.expand(
  "~/Desktop/Otolaryngology - Head and Neck Surgery alias/HCUP Data"
)
onedrive_data_root <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data"

if (dir.exists(local_data_root)) {
  data_root <- local_data_root
  message("Using LOCAL data root: ", data_root)
} else {
  data_root <- onedrive_data_root
  message("Local data root not found; falling back to OneDrive: ", data_root)
}

# Subdirectories (match the existing OneDrive structure)
d16_23 <- file.path(data_root, "NASS_HCUP Database", "RAW Data Files HNASS")
d21    <- file.path(data_root,
            "HCUP Original Data Files 21 through 23",
            "HCUP NASS 21 through 23", "NASS_2021")
d22    <- file.path(data_root,
            "HCUP Original Data Files 21 through 23",
            "HCUP NASS 21 through 23", "NASS_2022")
d23    <- file.path(data_root,
            "HCUP Original Data Files 21 through 23",
            "HCUP NASS 21 through 23", "NASS_2023")

# Sanity-check: warn early if any subdirectory is missing
for (d in c(d16_23, d21, d22, d23)) {
  if (!dir.exists(d)) warning("NASS data subdirectory not found: ", d)
}

enc_files <- list(
  "2016" = file.path(d16_23, "NASS_2016_Encounter-002.dta"),
  "2017" = file.path(d16_23, "NASS_2017_Encounter-003.dta"),
  "2018" = file.path(d16_23, "NASS_2018_Encounter-001.dta"),
  "2019" = file.path(d16_23, "NASS_2019_Encounter-005.dta"),
  "2020" = file.path(d16_23, "NASS_2020_Encounter.dta"),
  "2021" = file.path(d21,    "NASS_2021_Encounter.dta"),
  "2022" = file.path(d22,    "NASS_2022_Encounter.dta"),
  "2023" = file.path(d23,    "NASS_2023_Encounter.dta")
)

hosp_files <- list(
  "2016" = file.path(d16_23, "NASS_2016_Hospital.dta"),
  "2017" = file.path(d16_23, "NASS_2017_Hospital.dta"),
  "2018" = file.path(d16_23, "NASS_2018_Hospital.dta"),
  "2019" = file.path(d16_23, "NASS_2019_Hospital.dta"),
  "2020" = file.path(d16_23, "NASS_2020_Hospital.dta"),
  "2021" = file.path(d21,    "NASS_2021_Hospital.dta"),
  "2022" = file.path(d22,    "NASS_2022_Hospital.dta"),
  "2023" = file.path(d23,    "NASS_2023_Hospital.dta")
)

cpt_codes  <- c("64864","64865","64885","64886")
dx_stem    <- "S045"

# Phase 2 inclusion logic — broader open wound matching
s01_stem   <- "S01"   # open wound of head, any subcode
s11_stem   <- "S11"   # open wound of neck, any subcode

# External cause codes: W25-W27, W45, X99, Y28, V01-V99
# Matched against the leading 3 characters of any dx code.
ext_cause_pattern <- "^(W2[5-7]|W45|X99|Y28|V(0[1-9]|[1-9][0-9]))"

age_breaks <- c(18, 35, 45, 55, 65, Inf)
age_labels <- c("18-34","35-44","45-54","55-65","65+")

assign_age_group <- function(x) {
  cut(as.numeric(x), breaks=age_breaks, labels=age_labels,
      right=FALSE, include.lowest=TRUE)
}

# HCUP NASS variable coding (verified per official NASS Notes documentation)
# HOSP_LOCATION:  0=Rural, 1=Urban
# HOSP_TEACH:     0=Nonteaching, 1=Teaching
# HOSP_LOCTEACH:  1=Rural nonteaching, 2=Urban nonteaching, 3=Urban teaching
# HOSP_REGION:    1=Northeast, 2=Midwest, 3=South, 4=West
# FEMALE:         0=Male, 1=Female
# PAY1:           1=Medicare,2=Medicaid,3=Private,4=Self-pay,5=No charge,6=Other
# RACE:           1=White,2=Black,3=Hispanic,4=Asian/PI,5=Native American,6=Other
# ZIPINC_QRTL:    1=0-25th,2=26-50th,3=51-75th,4=76-100th
```

# Data Extraction

```{r extract-function}
extract_year <- function(year, enc_fp, hosp_fp) {
  message("  Loading: ", year)
  if (!file.exists(enc_fp))  { warning("Missing: ", enc_fp);  return(NULL) }
  if (!file.exists(hosp_fp)) { warning("Missing: ", hosp_fp); return(NULL) }

  enc  <- read_dta(enc_fp)
  hosp <- read_dta(hosp_fp)
  ec   <- names(enc); hc <- names(hosp)

  get_col <- function(pat, cols) grep(pat, cols, value=TRUE, ignore.case=TRUE)[1]

  dx_cols  <- grep("^i10_dx",   ec, value=TRUE, ignore.case=TRUE)
  cpt_cols <- grep("^cpt[0-9]", ec, value=TRUE, ignore.case=TRUE)

  age_col    <- get_col("^age$",          ec)
  female_col <- get_col("^female$",       ec)
  race_col   <- get_col("^race$",         ec)
  pay1_col   <- get_col("^pay1$",         ec)
  zip_col    <- get_col("^zipinc_qrtl$",  ec)
  totchg_col <- get_col("^totchg",        ec)
  discwt_col <- get_col("^discwt$",       ec)
  hosp_id_e  <- get_col("^hosp_nass$",    ec)
  hosp_id_h  <- get_col("^hosp_nass$",    hc)
  region_col <- get_col("^hosp_region$",  hc)
  loc_col    <- get_col("^hosp_location$",hc)
  teach_col  <- get_col("^hosp_teach$|^hosp_locteach$", hc)

  # Phase 2 flag construction (v1.5)
  # All injury-code flags exclude 7th character S (sequela / late effect)
  # because sequela codes represent late effects of prior injury, not
  # acute traumatic events. 7th character A (initial) and D (subsequent)
  # are retained — late primary nerve repairs may carry D in NASS.
  not_sequela <- function(cd) !str_ends(cd, "S")

  dx_s045 <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, dx_stem) & not_sequela(cd)
  })) > 0
  s01_flag <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, s01_stem) & not_sequela(cd)
  })) > 0
  s11_flag <- rowSums(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc[[col]]))
    str_starts(cd, s11_stem) & not_sequela(cd)
  })) > 0
  ext_cause_flag <- rowSums(sapply(dx_cols, function(col)
    str_detect(trimws(as.character(enc[[col]])), ext_cause_pattern))) > 0
  cpt_flag <- rowSums(sapply(cpt_cols, function(col)
    trimws(as.character(enc[[col]])) %in% cpt_codes)) > 0

  # v1.8: Neoplasm and iatrogenic exclusion flag
  # C07     parotid malignant neoplasm
  # D110    benign parotid neoplasm
  # C490    connective tissue head/face/neck
  # D333    benign cranial nerve neoplasm
  # C724    cranial nerve malignant
  # T854*   mechanical complication of prosthetic device
  # Y83*    surgical operation as cause of abnormal reaction
  exclusion_pattern <- "^(C07|D110|C490|D333|C724|T854|Y83)"
  exclusion_flag <- rowSums(sapply(dx_cols, function(col)
    str_detect(trimws(as.character(enc[[col]])), exclusion_pattern))) > 0

  # Phase 2 Group 1 (v1.5+v1.8) = (Criterion A OR Criterion B) AND NOT excluded
  #   Criterion A: S04.5* (A or D) co-occurring with S01.* or S11.* (A or D)
  #   Criterion B: CPT 64864/64865/64885/64886 co-occurring with S01.* or S11.* (A or D)
  # v1.8: drop encounters carrying parotid/cranial-nerve neoplasm or
  # surgical-complication codes — these are oncologic or iatrogenic, not acute trauma.
  # External cause is NOT a primary inclusion path; it remains for SA2/SA3 only.
  crit_a <- dx_s045 & (s01_flag | s11_flag)
  crit_b <- cpt_flag & (s01_flag | s11_flag)
  g1     <- (crit_a | crit_b) & !exclusion_flag
  g2     <- cpt_flag & g1
  keep   <- g1
  if (sum(keep) == 0) return(NULL)

  cpt_assigned <- rep(NA_character_, nrow(enc))
  for (code in cpt_codes) {
    hit <- rowSums(sapply(cpt_cols, function(col)
      trimws(as.character(enc[[col]])) == code)) > 0
    cpt_assigned[hit & is.na(cpt_assigned)] <- code
  }

  out <- tibble(
    year       = year,
    had_repair = as.integer(g2[keep]),
    cpt_code   = cpt_assigned[keep],
    discwt     = if (!is.na(discwt_col))
                   as.numeric(enc[[discwt_col]][keep]) else rep(1, sum(keep)),
    hosp_id    = if (!is.na(hosp_id_e))
                   as.character(enc[[hosp_id_e]][keep])
                 else as.character(seq_len(sum(keep))),
    totchg     = if (!is.na(totchg_col))
                   as.numeric(enc[[totchg_col]][keep]) else NA_real_,
    # Phase 2 — preserve per-row flags so sensitivity analyses can subset
    # enc_all directly without a re-extraction pass
    flag_s045      = dx_s045[keep],
    flag_s01       = s01_flag[keep],
    flag_s11       = s11_flag[keep],
    flag_extcause  = ext_cause_flag[keep],
    flag_cpt       = cpt_flag[keep],
    flag_crit_a    = crit_a[keep],
    flag_crit_b    = crit_b[keep],
    # Sensitivity Analysis 3 flags — laterality-specific nerve and anatomic wound
    flag_sa3_nerve = (rowSums(sapply(dx_cols, function(col)
      str_starts(trimws(as.character(enc[[col]])), "S0451") |
      str_starts(trimws(as.character(enc[[col]])), "S0452"))) > 0)[keep],
    flag_sa3_wound = (rowSums(sapply(dx_cols, function(col) {
      cd <- trimws(as.character(enc[[col]]))
      str_starts(cd, "S011") | str_starts(cd, "S013") |
      str_starts(cd, "S014") | str_starts(cd, "S015") |
      str_starts(cd, "S0181")
    })) > 0)[keep]
  )

  # Carry raw dx and cpt codes as pipe-joined strings for co-occurrence analysis.
  # Each element is a "|"-separated list of non-empty codes for that encounter.
  clean_codes <- function(mat) {
    apply(mat, 1, function(row) {
      codes <- trimws(as.character(row))
      codes <- codes[nzchar(codes) & codes != "." & codes != "NA" &
                     !is.na(codes) & codes != "0"]
      paste(codes, collapse="|")
    })
  }
  out$dx_raw  <- clean_codes(enc[keep, dx_cols,  drop=FALSE])
  out$cpt_raw <- if (length(cpt_cols) > 0)
    clean_codes(enc[keep, cpt_cols, drop=FALSE]) else rep("", sum(keep))

  out$age_group <- if (!is.na(age_col))
    as.character(assign_age_group(enc[[age_col]][keep]))
  else NA_character_

  if (!is.na(female_col)) {
    f <- as.numeric(enc[[female_col]][keep])
    out$sex <- case_when(f==1~"Female", f==0~"Male", TRUE~NA_character_)
  } else out$sex <- NA_character_

  if (!is.na(race_col)) {
    r <- as.character(as.numeric(enc[[race_col]][keep]))
    out$race <- case_when(
      r=="1" ~ "White", r=="2" ~ "Black", r=="3" ~ "Hispanic",
      r %in% c("4","5","6","7") ~
        "Asian or Pacific Islander or Native American or Other",
      TRUE ~ NA_character_)
  } else out$race <- NA_character_

  if (!is.na(pay1_col)) {
    p <- as.character(as.numeric(enc[[pay1_col]][keep]))
    out$insurance <- case_when(
      p=="1"~"Medicare", p=="2"~"Medicaid", p=="3"~"Private insurance",
      p=="4"~"Self-pay", p %in% c("5","6")~"Other", TRUE~NA_character_)
  } else out$insurance <- NA_character_

  if (!is.na(zip_col)) {
    z <- as.character(as.numeric(enc[[zip_col]][keep]))
    out$income_qrtl <- case_when(
      z=="1"~"0-25th", z=="2"~"26th to 50th",
      z=="3"~"51st to 75th", z=="4"~"76th to 100th", TRUE~NA_character_)
  } else out$income_qrtl <- NA_character_

  if (!is.na(hosp_id_h)) {
    hosp_slim <- hosp %>%
      mutate(hosp_id = as.character(.data[[hosp_id_h]]))

    hosp_slim$hosp_region <- if (!is.na(region_col))
      case_when(as.numeric(hosp[[region_col]])==1~"Northeast",
                as.numeric(hosp[[region_col]])==2~"Midwest",
                as.numeric(hosp[[region_col]])==3~"South",
                as.numeric(hosp[[region_col]])==4~"West",
                TRUE~NA_character_)
    else NA_character_

    hosp_slim$hosp_location <- if (!is.na(loc_col))
      case_when(as.numeric(hosp[[loc_col]])==0~"Rural",
                as.numeric(hosp[[loc_col]])==1~"Urban",
                TRUE~NA_character_)
    else NA_character_

    if (!is.na(teach_col)) {
      t_raw <- as.numeric(hosp[[teach_col]])
      hosp_slim$hosp_teach <- if (grepl("locteach", tolower(teach_col)))
        case_when(t_raw==3~"Teaching", t_raw %in% c(1,2)~"Nonteaching",
                  TRUE~NA_character_)
      else
        case_when(t_raw==1~"Teaching", t_raw==0~"Nonteaching",
                  TRUE~NA_character_)
    } else hosp_slim$hosp_teach <- NA_character_

    out <- left_join(out,
                     hosp_slim %>% select(hosp_id, hosp_region,
                                          hosp_location, hosp_teach),
                     by="hosp_id")
  } else {
    out$hosp_region   <- NA_character_
    out$hosp_location <- NA_character_
    out$hosp_teach    <- NA_character_
  }
  out
}
```

```{r run-extraction}
#| cache: true
message("Starting parallel extraction 2016-2023 (Phase 2 criteria)...")
message("Workers: ", n_cores, " (parallel::detectCores() = ",
        parallel::detectCores(), ")")
plan(multisession, workers = n_cores)
enc_all <- bind_rows(Filter(Negate(is.null),
  future_map2(names(enc_files), enc_files,
    ~ extract_year(.x, .y, hosp_files[[.x]]),
    .options = furrr_options(seed=TRUE))))
plan(sequential)

enc_all <- enc_all %>%
  mutate(discwt = as.numeric(discwt), had_repair = as.integer(had_repair))

message("Group 1 (unweighted): ", format(nrow(enc_all), big.mark=","))
message("Group 2 repair:       ", format(sum(enc_all$had_repair), big.mark=","))
```

# Survey Design

```{r survey-design}
svy <- svydesign(ids=~hosp_id, weights=~discwt, data=enc_all, nest=TRUE)

total_repair_w   <- sum(enc_all$discwt[enc_all$had_repair==1], na.rm=TRUE)
total_norepair_w <- sum(enc_all$discwt[enc_all$had_repair==0], na.rm=TRUE)
total_g1_w       <- sum(enc_all$discwt, na.rm=TRUE)

message("Weighted Group 1: ", format(round(total_g1_w),       big.mark=","))
message("Weighted Group 2: ", format(round(total_repair_w),   big.mark=","))
message("Weighted Group 3: ", format(round(total_norepair_w), big.mark=","))
```

# Helper Functions

```{r helpers}
fmt_p <- function(p) {
  sapply(p, function(x) {
    if (is.na(x)) return(NA_character_)
    if (x < 0.001) return("<0.001")
    sprintf("%.3f", round(x, 3))
  })
}

wt_cell <- function(var, level, repair_val) {
  idx   <- enc_all$had_repair==repair_val &
           !is.na(enc_all[[var]]) & enc_all[[var]]==level
  n_w   <- sum(enc_all$discwt[idx], na.rm=TRUE)
  denom <- if (repair_val==1) total_repair_w else total_norepair_w
  paste0(format(round(n_w), big.mark=","),
         " (", round(100*n_w/denom, 1), "%)")
}

wt_cell_g1 <- function(var, level) {
  idx <- !is.na(enc_all[[var]]) & enc_all[[var]]==level
  n_w <- sum(enc_all$discwt[idx], na.rm=TRUE)
  paste0(format(round(n_w), big.mark=","),
         " (", round(100*n_w/total_g1_w, 1), "%)")
}

repair_rate <- function(var, level) {
  idx_g1     <- !is.na(enc_all[[var]]) & enc_all[[var]]==level
  idx_repair <- idx_g1 & enc_all$had_repair==1
  n_g1  <- sum(enc_all$discwt[idx_g1],    na.rm=TRUE)
  n_rep <- sum(enc_all$discwt[idx_repair], na.rm=TRUE)
  if (n_g1 == 0) return(NA_character_)
  paste0(round(100*n_rep/n_g1, 1), "%")
}

svy_chisq_p <- function(var) {
  d <- enc_all %>%
    filter(!is.na(.data[[var]]),
           !.data[[var]] %in% c("Unknown","Missing/Unknown"))
  if (length(unique(d[[var]])) < 2) return(NA_real_)
  s <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  r <- tryCatch(
    svychisq(as.formula(paste0("~had_repair+",var)), design=s, statistic="Chisq"),
    error=function(e) NULL)
  if (is.null(r)) return(NA_real_)
  r$p.value
}

# Survey-weighted ANOVA on log-charges within a repair group
# Uses grp column to avoid .data[[]] formula issues in svyglm
svy_anova_charges_p <- function(repair_val, group_var) {
  d <- enc_all %>%
    filter(had_repair==repair_val, !is.na(totchg), totchg>0,
           !is.na(.data[[group_var]]),
           !.data[[group_var]] %in% c("Unknown","Missing/Unknown")) %>%
    mutate(grp = factor(.data[[group_var]]))
  if (length(unique(d$grp)) < 2 || nrow(d) < 10) return(NA_real_)
  s   <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  fit <- tryCatch(svyglm(log(totchg) ~ grp, design=s, family=gaussian()),
                  error=function(e) NULL)
  if (is.null(fit)) return(NA_real_)
  an  <- tryCatch(regTermTest(fit, "grp"), error=function(e) NULL)
  if (is.null(an)) return(NA_real_)
  an$p
}

# Survey-weighted t-test on log-charges between two levels
svy_ttest_charges_p <- function(repair_val, group_var) {
  d <- enc_all %>%
    filter(had_repair==repair_val, !is.na(totchg), totchg>0,
           !is.na(.data[[group_var]]),
           !.data[[group_var]] %in% c("Unknown","Missing/Unknown")) %>%
    mutate(grp = factor(.data[[group_var]]))
  if (length(unique(d$grp)) != 2 || nrow(d) < 10) return(NA_real_)
  s   <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  fit <- tryCatch(svyglm(log(totchg) ~ grp, design=s, family=gaussian()),
                  error=function(e) NULL)
  if (is.null(fit)) return(NA_real_)
  coefs <- summary(fit)$coefficients
  if (nrow(coefs) < 2) return(NA_real_)
  coefs[2, "Pr(>|t|)"]
}

wt_median_iqr <- function(data, repair_val) {
  d <- data %>% filter(had_repair==repair_val)
  if (nrow(d)==0) return(NA_character_)
  s <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  q <- tryCatch(
    svyquantile(~totchg, s, quantiles=c(0.25,0.5,0.75), na.rm=TRUE),
    error=function(e) NULL)
  if (is.null(q)) return(NA_character_)
  v <- as.numeric(q$totchg)
  paste0("$",format(round(v[2]),big.mark=","),
         " ($",format(round(v[1]),big.mark=","),
         "\u2013$",format(round(v[3]),big.mark=","),")")
}

# Weighted mean ± SD using age group midpoints
mean_sd_cell <- function(repair_val) {
  d   <- enc_all[which(enc_all$had_repair==as.integer(repair_val) &
                       !is.na(enc_all$age_group)), ]
  mid <- c("18-34"=26, "35-44"=39.5, "45-54"=49.5, "55-65"=60, "65+"=72)
  ages <- mid[d$age_group]
  wts  <- d$discwt
  m    <- weighted.mean(ages, wts, na.rm=TRUE)
  v    <- sum(wts*(ages-m)^2, na.rm=TRUE)/sum(wts, na.rm=TRUE)
  paste0(round(m,1)," \u00b1 ",round(sqrt(v),1))
}

mean_sd_cell_g1 <- function() {
  d   <- enc_all[which(!is.na(enc_all$age_group)), ]
  mid <- c("18-34"=26, "35-44"=39.5, "45-54"=49.5, "55-65"=60, "65+"=72)
  ages <- mid[d$age_group]
  wts  <- d$discwt
  m    <- weighted.mean(ages, wts, na.rm=TRUE)
  v    <- sum(wts*(ages-m)^2, na.rm=TRUE)/sum(wts, na.rm=TRUE)
  paste0(round(m,1)," \u00b1 ",round(sqrt(v),1))
}

make_rows <- function(var, levels_ordered, label) {
  hdr <- tibble(
    Characteristic    = label,
    `No Repair (n, %)` = NA_character_,
    `Repair (n, %)`    = NA_character_,
    `Total (n)`        = NA_character_,
    `Repair Rate (%)`  = NA_character_
  )
  sub <- map_dfr(levels_ordered, function(lev) {
    tibble(
      Characteristic     = paste0("   ", lev),
      `No Repair (n, %)` = wt_cell(var, lev, 0),
      `Repair (n, %)`    = wt_cell(var, lev, 1),
      `Total (n)`        = wt_cell_g1(var, lev),
      `Repair Rate (%)`  = repair_rate(var, lev)
    )
  })
  bind_rows(hdr, sub)
}
```

# Sheet 1 — National Estimates Table

```{r sheet1}
#| cache: false
total_rate <- paste0(round(100*total_repair_w/total_g1_w, 1), "%")

year_rows <- bind_rows(
  tibble(
    Characteristic     = "Total cases",
    `No Repair (n, %)` = paste0(format(round(total_norepair_w),big.mark=",")," (100%)"),
    `Repair (n, %)`    = paste0(format(round(total_repair_w),  big.mark=",")," (100%)"),
    `Total (n)`        = format(round(total_g1_w), big.mark=","),
    `Repair Rate (%)`  = total_rate
  ),
  map_dfr(as.character(2016:2023), function(yr) {
    idx_yr  <- enc_all$year==yr
    idx_rep <- idx_yr & enc_all$had_repair==1
    n_yr    <- sum(enc_all$discwt[idx_yr],  na.rm=TRUE)
    n_rep   <- sum(enc_all$discwt[idx_rep], na.rm=TRUE)
    tibble(
      Characteristic     = paste0("   ", yr),
      `No Repair (n, %)` = wt_cell("year", yr, 0),
      `Repair (n, %)`    = wt_cell("year", yr, 1),
      `Total (n)`        = format(round(n_yr), big.mark=","),
      `Repair Rate (%)`  = paste0(round(100*n_rep/n_yr, 1), "%")
    )
  })
)

sheet1 <- bind_rows(
  year_rows,
  tibble(
    Characteristic     = "Age, mean \u00b1 SD",
    `No Repair (n, %)` = mean_sd_cell(0),
    `Repair (n, %)`    = mean_sd_cell(1),
    `Total (n)`        = mean_sd_cell_g1(),
    `Repair Rate (%)`  = NA_character_
  ),
  make_rows("sex",         c("Male","Female"),                "Sex"),
  make_rows("age_group",   age_labels,                        "Age group"),
  make_rows("race",
    c("White","Black","Hispanic",
      "Asian or Pacific Islander or Native American or Other"), "Race"),
  make_rows("income_qrtl",
    c("0-25th","26th to 50th","51st to 75th","76th to 100th"),
    "Median household income of patients zipcode by quartile"),
  make_rows("insurance",
    c("Medicare","Medicaid","Private insurance","Self-pay","Other"),
    "Insurance type"),
  make_rows("hosp_region", c("Northeast","Midwest","South","West"),
            "Hospital region"),
  make_rows("hosp_location",c("Rural","Urban"),               "Hospital location"),
  make_rows("hosp_teach",  c("Nonteaching","Teaching"),       "Hospital teach")
)

sheet1 %>%
  kable(caption=paste0(
    "National Estimates of Patient and Hospital Characteristics by ",
    "Facial Nerve Repair Status, Including Repair Rates. ",
    "HCUP NASS 2016\u20132023. Weighted national estimates (DISCWT). ",
    "Total = All patients with traumatic facial nerve laceration nationally.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  row_spec(which(is.na(sheet1$`No Repair (n, %)`)), bold=TRUE)
```

# Sheet 2 — Multivariable Logistic Regression

```{r sheet2}
#| cache: true
reg_data <- enc_all %>%
  filter(!is.na(sex), !is.na(age_group), !is.na(race),
         !is.na(income_qrtl), !is.na(insurance),
         !is.na(hosp_region), !is.na(hosp_location), !is.na(hosp_teach)) %>%
  mutate(
    had_repair    = as.integer(had_repair),
    discwt        = as.numeric(discwt),
    sex           = relevel(factor(sex),           ref="Male"),
    age_group     = relevel(factor(age_group, levels=age_labels), ref="18-34"),
    race          = relevel(factor(race),           ref="White"),
    income_qrtl   = relevel(factor(income_qrtl,
                      levels=c("0-25th","26th to 50th",
                               "51st to 75th","76th to 100th")),
                      ref="76th to 100th"),
    insurance     = relevel(factor(insurance),      ref="Private insurance"),
    hosp_region   = relevel(factor(hosp_region),   ref="South"),
    hosp_location = relevel(factor(hosp_location), ref="Urban"),
    hosp_teach    = relevel(factor(hosp_teach),     ref="Nonteaching")
  )

svy_reg <- svydesign(ids=~hosp_id, weights=~discwt, data=reg_data, nest=TRUE)

model <- svyglm(
  had_repair ~ sex + age_group + race + income_qrtl + insurance +
               hosp_region + hosp_location + hosp_teach,
  design=svy_reg, family=quasibinomial(link="logit")
)

coef_df      <- as.data.frame(summary(model)$coefficients)
coef_df$term <- rownames(coef_df)

or_df <- coef_df %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR    = exp(Estimate),
    lower = exp(Estimate - 1.96 * `Std. Error`),
    upper = exp(Estimate + 1.96 * `Std. Error`),
    OR_CI = paste0(sprintf("%.2f",OR)," (",
                   sprintf("%.2f",lower),"\u2013",
                   sprintf("%.2f",upper),")"),
    p_fmt = fmt_p(`Pr(>|t|)`)
  )

term_map <- c(
  "sexFemale"="Female",
  "age_group35-44"="35-44","age_group45-54"="45-54",
  "age_group55-65"="55-65","age_group65+"="65+",
  "raceBlack"="Black","raceHispanic"="Hispanic",
  "raceAsian or Pacific Islander or Native American or Other"=
    "Asian or Pacific Islander or Native American or Other",
  "income_qrtl0-25th"="0-25th",
  "income_qrtl26th to 50th"="26th to 50th",
  "income_qrtl51st to 75th"="51st to 75th",
  "insuranceMedicare"="Medicare","insuranceMedicaid"="Medicaid",
  "insuranceSelf-pay"="Self-pay","insuranceOther"="Other",
  "hosp_regionNortheast"="Northeast","hosp_regionMidwest"="Midwest",
  "hosp_regionWest"="West",
  "hosp_locationRural"="Rural",
  "hosp_teachTeaching"="Teaching"
)

build_or_group <- function(label, levels_ord, ref) {
  hdr  <- tibble(Characteristic=label,
                 `Adjusted OR (95% CI)`=NA_character_,
                 `p value`=NA_character_)
  rows <- map_dfr(levels_ord, function(lev) {
    if (lev==ref) {
      tibble(Characteristic=paste0("   ",lev),
             `Adjusted OR (95% CI)`="Reference", `p value`=NA_character_)
    } else {
      key <- names(term_map)[term_map==lev]
      r   <- or_df %>% filter(term %in% key)
      tibble(Characteristic=paste0("   ",lev),
             `Adjusted OR (95% CI)`=if(nrow(r)>0) r$OR_CI[1] else NA_character_,
             `p value`=if(nrow(r)>0) r$p_fmt[1] else NA_character_)
    }
  })
  bind_rows(hdr, rows)
}

sheet2 <- bind_rows(
  build_or_group("Sex",         c("Male","Female"),              "Male"),
  build_or_group("Age group",   age_labels,                      "18-34"),
  build_or_group("Race",
    c("White","Black","Hispanic",
      "Asian or Pacific Islander or Native American or Other"),  "White"),
  build_or_group("Median household income of patients zipcode by quartile",
    c("0-25th","26th to 50th","51st to 75th","76th to 100th"), "76th to 100th"),
  build_or_group("Insurance type",
    c("Medicare","Medicaid","Private insurance","Self-pay","Other"),
    "Private insurance"),
  build_or_group("Hospital region",
    c("Northeast","Midwest","South","West"),                     "South"),
  build_or_group("Hospital location", c("Rural","Urban"),        "Urban"),
  build_or_group("Hospital teach",    c("Nonteaching","Teaching"),"Nonteaching")
)

# Build the dynamic "significant predictors" string for footnotes
sig_predictors <- or_df %>%
  filter(`Pr(>|t|)` < 0.05, term %in% names(term_map)) %>%
  mutate(label = term_map[term]) %>%
  pull(label)
sig_pred_text <- if (length(sig_predictors) > 0) {
  paste(sig_predictors, collapse=", ")
} else {
  "none at p<0.05"
}

sheet2 %>%
  kable(caption="Multivariable Logistic Regression for Odds of Facial Nerve Repair.") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  row_spec(which(is.na(sheet2$`Adjusted OR (95% CI)`) |
                   sheet2$`Adjusted OR (95% CI)`==""), bold=TRUE)
```

# Sheet 4 — Repair Rates by Year

```{r sheet4}
#| cache: false
sheet4 <- enc_all %>%
  group_by(year) %>%
  summarise(
    `Traumatic facial nerve injury (unweighted N)` = n(),
    `Traumatic facial nerve injury (weighted N)`   = round(sum(discwt, na.rm=TRUE)),
    `Repair (unweighted N)`      = sum(had_repair),
    `Repair (weighted N)`        = round(sum(discwt[had_repair==1], na.rm=TRUE)),
    `Repair rate — weighted (%)` = round(
      100*sum(discwt[had_repair==1],na.rm=TRUE)/sum(discwt,na.rm=TRUE),1),
    .groups="drop"
  ) %>% rename(Year=year)

sheet4 %>%
  kable(caption="Repair Rates by Year. HCUP NASS 2016-2023.",
        format.args=list(big.mark=",")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)

# Survey-weighted chi-square test for repair rate difference across years
year_chisq <- svychisq(~had_repair + year,
                        design = svy,
                        statistic = "Chisq")
year_p <- fmt_p(year_chisq$p.value)

message("Chi-square test for repair rate across years: p = ", year_p)
```

# Sheet 5 — Charges Analysis

```{r sheet5}
#| cache: false
charges_data <- enc_all %>% filter(!is.na(totchg), totchg > 0)

overall_row <- tibble(
  Characteristic                                 = "Total charges of encounter",
  `No Repair Median (IQR)`                       = wt_median_iqr(charges_data, 0),
  `Repair Median (IQR)`                          = wt_median_iqr(charges_data, 1),
  `p value (no repair: within-group comparison)` = NA_character_,
  `p value (repair: within-group comparison)`    = NA_character_
)

ins_levels      <- c("Medicare","Medicaid","Private insurance","Self-pay","Other")
ins_anova_norep <- fmt_p(svy_anova_charges_p(0, "insurance"))
ins_anova_rep   <- fmt_p(svy_anova_charges_p(1, "insurance"))

ins_hdr <- tibble(
  Characteristic                                 = "Insurance type",
  `No Repair Median (IQR)`                       = NA_character_,
  `Repair Median (IQR)`                          = NA_character_,
  `p value (no repair: within-group comparison)` = ins_anova_norep,
  `p value (repair: within-group comparison)`    = ins_anova_rep
)

ins_rows <- map_dfr(ins_levels, function(lev) {
  d <- charges_data %>% filter(!is.na(insurance), insurance==lev)
  tibble(
    Characteristic                                 = paste0("   ", lev),
    `No Repair Median (IQR)`                       = wt_median_iqr(d, 0),
    `Repair Median (IQR)`                          = wt_median_iqr(d, 1),
    `p value (no repair: within-group comparison)` = NA_character_,
    `p value (repair: within-group comparison)`    = NA_character_
  )
})

teach_ttest_norep <- fmt_p(svy_ttest_charges_p(0, "hosp_teach"))
teach_ttest_rep   <- fmt_p(svy_ttest_charges_p(1, "hosp_teach"))

teach_hdr <- tibble(
  Characteristic                                 = "Hospital teaching status",
  `No Repair Median (IQR)`                       = NA_character_,
  `Repair Median (IQR)`                          = NA_character_,
  `p value (no repair: within-group comparison)` = teach_ttest_norep,
  `p value (repair: within-group comparison)`    = teach_ttest_rep
)

teach_rows <- map_dfr(c("Nonteaching","Teaching"), function(lev) {
  d <- charges_data %>% filter(!is.na(hosp_teach), hosp_teach==lev)
  tibble(
    Characteristic                                 = paste0("   ", lev),
    `No Repair Median (IQR)`                       = wt_median_iqr(d, 0),
    `Repair Median (IQR)`                          = wt_median_iqr(d, 1),
    `p value (no repair: within-group comparison)` = NA_character_,
    `p value (repair: within-group comparison)`    = NA_character_
  )
})

sheet5 <- bind_rows(overall_row, ins_hdr, ins_rows, teach_hdr, teach_rows)

sheet5 %>%
  kable(caption=paste0(
    "Charge Analysis: Repair vs No Repair. Median (IQR) weighted national estimates. ",
    "Insurance type p-values: survey-weighted ANOVA on log-charges within each group. ",
    "Teaching status p-values: survey-weighted t-test on log-charges within each group.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE) %>%
  row_spec(which(is.na(sheet5$`No Repair Median (IQR)`)), bold=TRUE)
```

# NIS 2023 — Facial Nerve Repair (Inpatient)

```{r nis-data}
#| cache: false
# Source: completed three-phase NIS analysis
#   v1.0 nis_facial_nerve_supplement_query    (frequency, all PCS codes)
#   v1.1 nis_facial_nerve_miscoding_confirmation (diagnosis profile)
#   v1.2 nis_facial_nerve_trauma_classification  (trauma classification)
#
# Survey design throughout: NIS_STRATUM strata, HOSP_NIS cluster, DISCWT
# weight, nest=TRUE, options(survey.lonely.psu="adjust"). DISCWT read
# directly without scaling. Column positions verified against HCUP's
# StataLoad_NIS_2023_Core.Do (12/10/2025).
#
# Per HCUP DUA, individual unweighted counts <11 are suppressed (NA) in
# this published table; the per-criterion subtotal rows below the union
# are shown as suppressed where the underlying cell <11.
nis_sheet <- tibble(
  Section = c(
    "ALL FACIAL NERVE REPAIR",
    "ALL FACIAL NERVE REPAIR",
    "ALL FACIAL NERVE REPAIR",
    "TRAUMATIC SUBSET",
    "TRAUMATIC SUBSET",
    "TRAUMATIC SUBSET",
    "TRAUMATIC SUBSET",
    "TRAUMATIC SUBSET",
    "TRAUMATIC SUBSET",
    "NON-TRAUMATIC SUBSET"
  ),
  Category = c(
    "Total inpatient facial nerve repair (any PCS code)",
    "   PNS section, correctly coded (01QM0/3/4ZZ)",
    "   CNS section, miscoded (00QM0/3/4ZZ)",
    "Traumatic union (any of S04.5, S01.4/S01.8, V01\u2013Y99)",
    "   Per-criterion: facial nerve injury dx (S04.5*)",
    "   Per-criterion: facial laceration dx (S01.4*/S01.8*)",
    "   Per-criterion: external cause code (V01\u2013Y99)",
    "   Of which: CNS-miscoded (00QM)",
    "   Of which: PNS-correctly-coded (01QM)",
    "Non-traumatic remainder (elective / oncologic / iatrogenic)"
  ),
  `Unweighted N` = c(
    57L, NA_integer_, 54L,           # All repair (PNS=3 suppressed per <11 rule)
    26L, 14L, 14L, 24L,              # Traumatic union and per-criterion
    25L, NA_integer_,                # CNS-miscoded traumatic / PNS traumatic (=1, suppressed)
    31L                              # Non-traumatic
  ),
  `Weighted N (national estimate)` = c(
    285L, 15L, 270L,
    130L, 70L, 70L, 120L,
    125L, 5L,
    155L
  )
)

# Inline metrics for the footnote and downstream rendering
nis_total_w        <- 285
nis_cns_w          <- 270
nis_cns_share      <- round(100 * nis_cns_w / nis_total_w, 1)         # 94.7%
nis_trauma_w       <- 130
nis_trauma_share   <- round(100 * nis_trauma_w / nis_total_w, 1)      # 45.6%
nis_trauma_cns     <- 125
nis_trauma_miscode <- round(100 * nis_trauma_cns / nis_trauma_w, 1)   # 96.2%
nis_miscode_confirmed_pct <- 75.9     # diagnosis profile match rate (v1.1)

message("NIS 2023 \u2014 Total inpatient facial nerve repair: ",
        nis_total_w, " weighted (", nis_cns_share,
        "% CNS-miscoded)")
message("NIS 2023 \u2014 Traumatic subset: ",
        nis_trauma_w, " weighted (", nis_trauma_share,
        "% of all repair)")
message("NIS 2023 \u2014 Of traumatic cases, ", nis_trauma_miscode,
        "% are CNS-miscoded")

nis_sheet %>%
  kable(caption=paste0(
    "NIS 2023 Facial Nerve Repair \u2014 Inpatient Database Results, with ",
    "Traumatic Classification. Total inpatient burden of any facial nerve repair = ",
    nis_total_w, " weighted (", nis_cns_share,
    "% miscoded to CNS section). The TRAUMATIC subset = ",
    nis_trauma_w, " weighted (", nis_trauma_share,
    "% of all facial nerve repair); ", nis_trauma_miscode,
    "% of traumatic cases are CNS-miscoded. ",
    "PNS = Peripheral Nervous System (section 01, correct under ICD-10-PCS); ",
    "CNS = Central Nervous System (section 00). The miscoding hypothesis ",
    "was confirmed in v1.1 by diagnosis-profile analysis (",
    nis_miscode_confirmed_pct, "% of CNS-miscoded cases carry a ",
    "facial-nerve-consistent diagnosis). NA = suppressed per HCUP DUA ",
    "(unweighted N < 11). HCUP NIS 2023."),
    format.args=list(big.mark=",")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  pack_rows("All facial nerve repair",            1, 3) %>%
  pack_rows("Traumatic subset (45.6% of total)",  4, 9) %>%
  pack_rows("Non-traumatic subset",              10, 10) %>%
  row_spec(1, bold=TRUE, background="#FCF3CF") %>%        # all-repair total
  row_spec(4, bold=TRUE, background="#D5F5E3") %>%        # traumatic union
  row_spec(10, bold=TRUE, background="#E8DAEF") %>%       # non-traumatic
  row_spec(c(3,8), color="#6C3483", italic=TRUE)          # CNS-miscoded rows
```

# Publication Figures

## Figure 1 — Study Population Flow Diagram

```{r fig-flow}
g1_n <- nrow(enc_all)
g1_w <- round(total_g1_w)
g2_n <- sum(enc_all$had_repair==1)
g2_w <- round(total_repair_w)
g3_n <- sum(enc_all$had_repair==0)
g3_w <- round(total_norepair_w)

flow <- grViz(paste0('
digraph flow {
  graph [layout=dot, rankdir=TB, fontname="Arial", nodesep=0.7, ranksep=0.9]
  node [shape=rectangle, style="filled,rounded", fontname="Arial",
        fontsize=11, margin="0.25,0.15", penwidth=1.5]

  A [label="HCUP NASS 2016-2023\\nFull Ambulatory Surgery Sample\\n(all encounter years combined)",
     fillcolor="#1A5276", fontcolor=white, color="#1A5276"]

  B [label="Group 1: Traumatic Facial Nerve Laceration\\nUnweighted N = ', format(g1_n, big.mark=","), '\\nWeighted national estimate = ', format(g1_w, big.mark=","), '\\n\\nInclusion criteria (either A or B):\\nA. S04.5* (initial or subsequent encounter,\\n    7th character A or D) co-occurring with\\n    S01.* or S11.* (initial or subsequent)\\nB. CPT 64864, 64865, 64885, or 64886\\n    co-occurring with S01.* or S11.* (A or D)",
     fillcolor="#2E86C1", fontcolor=white, color="#2175A9"]

  C [label="Group 2: Traumatic Facial Nerve Laceration WITH Repair\\nUnweighted N = ', format(g2_n, big.mark=","), '  |  Weighted N = ', format(g2_w, big.mark=","), '\\n\\nInclusion criteria:\\nAny encounter from Group 1\\nco-occurring with CPT for facial nerve repair or graft\\n(CPT 64864, 64865, 64885, or 64886)",
     fillcolor="#EBF5FB", fontcolor="#1A5276", color="#2E86C1"]

  D [label="Group 3: Traumatic Facial Nerve Laceration WITHOUT Repair\\nUnweighted N = ', format(g3_n, big.mark=","), '  |  Weighted N = ', format(g3_w, big.mark=","), '\\n\\nInclusion criteria:\\nGroup 1 minus Group 2\\n(no qualifying CPT code for facial nerve suture present)",
     fillcolor="#FEF9E7", fontcolor="#7D6608", color="#D4AC0D"]

  A -> B [label="  Injury identification", fontname="Arial", fontsize=10,
          color="#566573", style=bold]
  B -> C [label="  CPT repair\\n  code present", fontname="Arial",
          fontsize=10, color="#2E86C1", style=bold, fontcolor="#2E86C1"]
  B -> D [label="  No CPT repair\\n  code present", fontname="Arial",
          fontsize=10, color="#D4AC0D", style=bold, fontcolor="#D4AC0D"]
}'))

flow
```

```{r save-flow}
#| cache: false
tryCatch({
  svg_txt  <- export_svg(flow)
  svg_path <- tempfile(fileext=".svg")
  writeLines(svg_txt, svg_path)
  rsvg_png(svg_path,
           file  = file.path(fig_dir, "fig1_flow_diagram.png"),
           width = 1200)
  rsvg_pdf(svg_path,
           file  = file.path(fig_dir, "fig1_flow_diagram.pdf"))
  message("Flow diagram saved as PNG and PDF.")
}, error = function(e) {
  message("Flow diagram save error: ", e$message)
})
```

## Figure 2 — Repair Trend by Year

```{r fig-trend}
trend_data <- sheet4 %>%
  mutate(
    Year     = as.integer(Year),
    repair_w = `Repair (weighted N)`,
    injury_w = `Traumatic facial nerve injury (weighted N)`,
    pct      = `Repair rate — weighted (%)`
  )

max_w <- max(trend_data$injury_w, na.rm=TRUE)

fig_trend <- ggplot(trend_data, aes(x=Year)) +
  geom_col(aes(y=injury_w, fill="All traumatic facial nerve laceration"),
           width=0.6, alpha=0.5) +
  geom_col(aes(y=repair_w, fill="With repair"), width=0.6) +
  geom_line(aes(y=pct*max_w/100), color="#C00000",
            linewidth=1.2, linetype="dashed") +
  geom_point(aes(y=pct*max_w/100), color="#C00000", size=3) +
  geom_text(aes(y=pct*max_w/100, label=paste0(pct,"%")),
            vjust=-0.9, size=3.5, color="#C00000", fontface="bold") +
  scale_y_continuous(
    name=  "Weighted encounters (N)", labels=label_comma(),
    sec.axis=sec_axis(~.*100/max_w, name="Repair rate (%)",
                      labels=function(x) paste0(x,"%"))) +
  scale_x_continuous(breaks=2016:2023) +
  scale_fill_manual(values=c("All traumatic facial nerve laceration"="#4E79A7",
                              "With repair"="#F28E2B")) +
  labs(title="Traumatic Facial Nerve Laceration Encounters and Repair Rate by Year",
       subtitle="HCUP NASS 2016\u20132023 | Weighted national estimates",
       x="Year", fill=NULL,
       caption="Dashed red line = weighted repair rate (right axis)") +
  theme_classic(base_size=12) +
  theme(legend.position="bottom", plot.title=element_text(face="bold"),
        axis.title=element_text(face="bold"))

fig_trend
ggsave(file.path(fig_dir,"fig2_repair_trend.pdf"),
       fig_trend, width=10, height=6)
ggsave(file.path(fig_dir,"fig2_repair_trend.png"),
       fig_trend, width=10, height=6, dpi=300)
```

## Figure 3 — Forest Plot

```{r fig-forest}
#| cache: false
forest_data <- or_df %>%
  mutate(
    label = term_map[term],
    group = case_when(
      str_starts(term,"sex")          ~ "Sex",
      str_starts(term,"age_group")    ~ "Age group",
      str_starts(term,"race")         ~ "Race",
      str_starts(term,"income_qrtl")  ~ "Income quartile",
      str_starts(term,"insurance")    ~ "Insurance type",
      str_starts(term,"hosp_region")  ~ "Hospital region",
      TRUE ~ "Other")
  ) %>%
  filter(!is.na(label),
         !str_starts(term, "hosp_location"),
         !str_starts(term, "hosp_teach")) %>%
  mutate(label=factor(label, levels=rev(unique(label))),
         sig=`Pr(>|t|)` < 0.05)

fig_forest <- ggplot(forest_data,
                     aes(x=OR, y=label, xmin=lower, xmax=upper, color=sig)) +
  geom_vline(xintercept=1, linetype="dashed", color="grey50", linewidth=0.8) +
  geom_errorbarh(height=0.3, linewidth=0.7) +
  geom_point(size=3) +
  geom_text(aes(x=upper,
                label=paste0(sprintf("%.2f",OR)," (",
                             sprintf("%.2f",lower),"\u2013",
                             sprintf("%.2f",upper),")")),
            hjust=-0.1, size=2.8, color="black") +
  scale_color_manual(values=c("FALSE"="grey50","TRUE"="#C00000"),
                     labels=c("FALSE"="p\u22650.05","TRUE"="p<0.05"),name="") +
  scale_x_log10(limits=c(0.05,25),
                breaks=c(0.1,0.25,0.5,1,2,5,10),
                labels=c("0.1","0.25","0.5","1","2","5","10")) +
  facet_grid(group~., scales="free_y", space="free_y") +
  labs(title="Multivariable Logistic Regression for Odds of Facial Nerve Repair",
       subtitle="Adjusted odds ratios | HCUP NASS 2016\u20132023",
       x="Adjusted Odds Ratio (log scale)", y=NULL,
       caption=paste0("Reference: Male, Age 18-34, White, 76th-100th income quartile,\n",
                      "Private insurance, South region. Error bars = 95% CI.\n",
                      "Hospital location and teaching status included in model but",
                      " not shown.")) +
  theme_classic(base_size=11) +
  theme(strip.text=element_text(face="bold", size=9),
        strip.background=element_rect(fill="#EBF4FB", color=NA),
        plot.title=element_text(face="bold"), legend.position="bottom")

fig_forest
ggsave(file.path(fig_dir,"fig3_forest_plot.pdf"),
       fig_forest, width=10, height=10)
ggsave(file.path(fig_dir,"fig3_forest_plot.png"),
       fig_forest, width=10, height=10, dpi=300)
```

## Figure 4 — Geographic Heat Map

```{r fig-heatmap}
region_data <- enc_all %>%
  filter(!is.na(hosp_region)) %>%
  group_by(hosp_region) %>%
  summarise(repair_rate_w=round(
    100*sum(discwt[had_repair==1],na.rm=TRUE)/sum(discwt,na.rm=TRUE),2),
    .groups="drop")

state_region <- tibble(
  state=c("connecticut","maine","massachusetts","new hampshire","rhode island",
          "vermont","new jersey","new york","pennsylvania",
          "illinois","indiana","michigan","ohio","wisconsin","iowa","kansas",
          "minnesota","missouri","nebraska","north dakota","south dakota",
          "delaware","florida","georgia","maryland","north carolina",
          "south carolina","virginia","west virginia","alabama","kentucky",
          "mississippi","tennessee","arkansas","louisiana","oklahoma","texas",
          "district of columbia","arizona","colorado","idaho","montana",
          "nevada","new mexico","utah","wyoming","alaska","california",
          "hawaii","oregon","washington"),
  hosp_region=c(rep("Northeast",9),rep("Midwest",12),
                rep("South",17),   rep("West",13))
)

us_map <- map_data("state") %>%
  left_join(state_region, by=c("region"="state")) %>%
  left_join(region_data, by="hosp_region")

region_centers <- us_map %>%
  group_by(hosp_region, repair_rate_w) %>%
  summarise(long=mean(long,na.rm=TRUE), lat=mean(lat,na.rm=TRUE),
            .groups="drop") %>%
  filter(!is.na(hosp_region))

fig_map <- ggplot(us_map, aes(x=long, y=lat, group=group, fill=repair_rate_w)) +
  geom_polygon(color="white", linewidth=0.3) +
  geom_text(data=region_centers,
            aes(x=long, y=lat,
                label=paste0(hosp_region,"\n",repair_rate_w,"%"), group=NULL),
            size=3.5, fontface="bold", color="white") +
  scale_fill_gradient(low="#D6EAF8", high="#1A5276",
                      name="Repair rate\n(weighted %)", na.value="grey85") +
  coord_map("albers", lat0=39, lat1=45) +
  labs(title="Traumatic Facial Nerve Repair Rate by U.S. Census Region",
       subtitle="HCUP NASS 2016\u20132023 | Weighted national estimates",
       caption=paste0("State shading reflects the repair rate for its Census region.\n",
                      "State-level estimates suppressed per HCUP cell size restrictions.\n",
                      "NOTE: 2023 HOSP_REGION not available in NASS Hospital file; ",
                      "regional estimates reflect 2016\u20132022 data only."),
       x=NULL, y=NULL) +
  theme_void(base_size=12) +
  theme(plot.title=element_text(face="bold", hjust=0.5),
        plot.subtitle=element_text(hjust=0.5, color="grey40"),
        plot.caption=element_text(hjust=0.5, size=9, color="grey50"),
        legend.position="right")

fig_map
ggsave(file.path(fig_dir,"fig4_heatmap_region.pdf"),
       fig_map, width=12, height=7)
ggsave(file.path(fig_dir,"fig4_heatmap_region.png"),
       fig_map, width=12, height=7, dpi=300)
```

## Figure 5 — IPTW: Effect of Repair on Total Charges

```{r fig-iptw}
#| cache: true
iptw_data <- enc_all %>%
  filter(!is.na(totchg), totchg>0,
         !is.na(sex), !is.na(age_group), !is.na(race),
         !is.na(income_qrtl), !is.na(insurance),
         !is.na(hosp_region), !is.na(hosp_location), !is.na(hosp_teach)) %>%
  mutate(
    had_repair    = as.integer(had_repair),
    sex           = factor(sex),
    age_group     = factor(age_group, levels=age_labels),
    race          = factor(race),
    income_qrtl   = factor(income_qrtl),
    insurance     = factor(insurance),
    hosp_region   = factor(hosp_region),
    hosp_location = factor(hosp_location),
    hosp_teach    = factor(hosp_teach)
  )

set.seed(42)
w_out <- weightit(
  had_repair ~ sex + age_group + race + income_qrtl + insurance +
               hosp_region + hosp_location + hosp_teach,
  data=iptw_data, method="ps", estimand="ATE"
)

iptw_data$iptw_weight <- w_out$weights * iptw_data$discwt

svy_iptw   <- svydesign(ids=~hosp_id, weights=~iptw_weight,
                         data=iptw_data, nest=TRUE)
iptw_model <- svyglm(log(totchg) ~ had_repair,
                     design=svy_iptw, family=gaussian())

iptw_coef <- summary(iptw_model)$coefficients
eff_pct   <- round((exp(iptw_coef["had_repair","Estimate"])-1)*100, 1)
ci_lo     <- round((exp(iptw_coef["had_repair","Estimate"] -
                        1.96*iptw_coef["had_repair","Std. Error"])-1)*100,1)
ci_hi     <- round((exp(iptw_coef["had_repair","Estimate"] +
                        1.96*iptw_coef["had_repair","Std. Error"])-1)*100,1)
iptw_p    <- fmt_p(iptw_coef["had_repair","Pr(>|t|)"])

iptw_excel <- tibble(
  Metric = c(
    "Method",
    "Outcome variable",
    "Treatment variable",
    "Covariates balanced",
    "IPTW estimand",
    "No repair \u2014 median total charges (unweighted)",
    "With repair \u2014 median total charges (unweighted)",
    "Effect of repair on charges (IPTW-adjusted)",
    "95% Confidence interval",
    "p value",
    "Interpretation"
  ),
  Value = c(
    "Inverse Probability of Treatment Weighting (IPTW) using propensity scores (WeightIt package)",
    "Log total charges",
    "Had repair (CPT 64864, 64865, 64885, or 64886) vs no repair",
    "Sex, age group, race, income quartile, insurance type, hospital region, location, teaching status",
    "Average Treatment Effect (ATE)",
    paste0("$", format(round(median(iptw_data$totchg[iptw_data$had_repair==0],
                                    na.rm=TRUE)), big.mark=",")),
    paste0("$", format(round(median(iptw_data$totchg[iptw_data$had_repair==1],
                                    na.rm=TRUE)), big.mark=",")),
    paste0(eff_pct, "% higher charges associated with repair"),
    paste0(ci_lo, "% to ", ci_hi, "%"),
    iptw_p,
    paste0("After balancing patient and hospital characteristics via IPTW, ",
           "facial nerve repair was associated with a ", eff_pct,
           "% increase in total charges (95% CI: ", ci_lo, "% to ", ci_hi,
           "%, p=", iptw_p, ").")
  )
)

fig_iptw <- ggplot(iptw_data, aes(x=log(totchg), fill=factor(had_repair))) +
  geom_density(alpha=0.6) +
  scale_fill_manual(values=c("0"="#4E79A7","1"="#F28E2B"),
                    labels=c("0"="No repair","1"="With repair"), name="") +
  labs(title="Distribution of Log Total Charges by Repair Status",
       subtitle=paste0("IPTW-adjusted. Repair associated with ",
                       eff_pct,"% change in charges (p=",iptw_p,")"),
       x="Log total charges", y="Density") +
  theme_classic(base_size=12) +
  theme(plot.title=element_text(face="bold"), legend.position="bottom")

fig_iptw
ggsave(file.path(fig_dir,"fig5_iptw_charges.pdf"),
       fig_iptw, width=9, height=6)
ggsave(file.path(fig_dir,"fig5_iptw_charges.png"),
       fig_iptw, width=9, height=6, dpi=300)
```

# Sensitivity Analyses

## Sensitivity Analysis 1 — Criterion A Only (no external cause, no Criterion B)

```{r sensitivity-1}
#| cache: false
# Criterion A restricted to S04.5 + (S01.* OR S11.*), excluding the external
# cause path AND excluding Criterion B (procedural pathway).
sens1_data <- enc_all %>%
  filter(flag_s045 & (flag_s01 | flag_s11))

sens1_g1_unw <- nrow(sens1_data)
sens1_g1_w   <- round(sum(sens1_data$discwt, na.rm=TRUE))
sens1_g2_unw <- sum(sens1_data$had_repair == 1)
sens1_g2_w   <- round(sum(sens1_data$discwt[sens1_data$had_repair == 1], na.rm=TRUE))
sens1_g3_unw <- sens1_g1_unw - sens1_g2_unw
sens1_g3_w   <- sens1_g1_w - sens1_g2_w
sens1_rate   <- if (sens1_g1_w > 0) round(100*sens1_g2_w/sens1_g1_w, 1) else NA_real_

sens1_tbl <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate (%)"
  ),
  Value = c(
    format(sens1_g1_unw, big.mark=","),
    format(sens1_g1_w,   big.mark=","),
    format(sens1_g2_unw, big.mark=","),
    format(sens1_g2_w,   big.mark=","),
    format(sens1_g3_unw, big.mark=","),
    format(sens1_g3_w,   big.mark=","),
    paste0(sens1_rate, "%")
  )
)

sens1_tbl %>%
  kable(caption=paste0(
    "Sensitivity Analysis 1 \u2014 Criterion A only. ",
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, ",
    "excluding the external-cause-only path and excluding the Criterion B ",
    "procedural pathway. HCUP NASS 2016\u20132023.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

## Sensitivity Analysis 2 — Highest Mechanistic Specificity

```{r sensitivity-2}
#| cache: false
# Highest mechanistic specificity: S04.5 + (S01.* OR S11.*) AND external cause
sens2_data <- enc_all %>%
  filter(flag_s045 & (flag_s01 | flag_s11) & flag_extcause)

sens2_g1_unw <- nrow(sens2_data)
sens2_g1_w   <- round(sum(sens2_data$discwt, na.rm=TRUE))
sens2_g2_unw <- sum(sens2_data$had_repair == 1)
sens2_g2_w   <- round(sum(sens2_data$discwt[sens2_data$had_repair == 1], na.rm=TRUE))
sens2_g3_unw <- sens2_g1_unw - sens2_g2_unw
sens2_g3_w   <- sens2_g1_w - sens2_g2_w
sens2_rate   <- if (sens2_g1_w > 0) round(100*sens2_g2_w/sens2_g1_w, 1) else NA_real_

sens2_tbl <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate (%)"
  ),
  Value = c(
    format(sens2_g1_unw, big.mark=","),
    format(sens2_g1_w,   big.mark=","),
    format(sens2_g2_unw, big.mark=","),
    format(sens2_g2_w,   big.mark=","),
    format(sens2_g3_unw, big.mark=","),
    format(sens2_g3_w,   big.mark=","),
    paste0(sens2_rate, "%")
  )
)

sens2_tbl %>%
  kable(caption=paste0(
    "Sensitivity Analysis 2 \u2014 Highest mechanistic specificity. ",
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
    "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "HCUP NASS 2016\u20132023.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

## Sensitivity Analysis 3 — Maximum Specificity (Laterality + Anatomic Wound + External Cause)

```{r sensitivity-3}
#| cache: false
# Laterality-specified facial nerve injury (S04.51* right, S04.52* left)
# PLUS anatomically concordant facial wound:
#   S01.1* = eyelid/periocular, S01.3* = ear, S01.4* = cheek/TMJ,
#   S01.5* = lip/oral cavity,   S01.81* = other part of face
# PLUS external cause sharp object or transport mechanism
#   (W25-W27, W45, X99, Y28, V01-V99 — already captured as flag_extcause)
# Group 2 = Group 1 + CPT 64864/64865/64885/64886 (= had_repair == 1)

sa3_g1 <- enc_all$flag_sa3_nerve & enc_all$flag_sa3_wound & enc_all$flag_extcause

sa3_g1_unw <- sum(sa3_g1)
sa3_g1_w   <- round(sum(enc_all$discwt[sa3_g1], na.rm=TRUE))
sa3_g2_unw <- sum(sa3_g1 & enc_all$had_repair == 1)
sa3_g2_w   <- round(sum(enc_all$discwt[sa3_g1 & enc_all$had_repair == 1], na.rm=TRUE))
sa3_g3_unw <- sa3_g1_unw - sa3_g2_unw
sa3_g3_w   <- sa3_g1_w - sa3_g2_w
sa3_rate   <- if (sa3_g1_w > 0) round(100 * sa3_g2_w / sa3_g1_w, 1) else NA_real_

message("=== Sensitivity Analysis 3 ===")
message("Group 1 unweighted: ", format(sa3_g1_unw, big.mark=","))
message("Group 1 weighted:   ", format(sa3_g1_w,   big.mark=","))
message("Repair rate:        ", sa3_rate, "%")

# Note if flag_extcause finds no rows — may indicate external cause codes
# were not present in the data for the qualifying nerve/wound population
if (sa3_g1_unw == 0) {
  warning("Sensitivity Analysis 3: zero qualifying encounters. ",
          "External cause codes (W25-W27, W45, X99, Y28, V*) may not be ",
          "available in NASS for this population. Verify flag_extcause.")
}

sens3_tbl <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate (%)"
  ),
  Value = c(
    format(sa3_g1_unw, big.mark=","),
    format(sa3_g1_w,   big.mark=","),
    format(sa3_g2_unw, big.mark=","),
    format(sa3_g2_w,   big.mark=","),
    format(sa3_g3_unw, big.mark=","),
    format(sa3_g3_w,   big.mark=","),
    paste0(sa3_rate, "%")
  )
)

sens3_tbl %>%
  kable(caption=paste0(
    "Sensitivity Analysis 3 \u2014 Maximum Specificity. ",
    "Requires laterality-specified facial nerve injury (S04.51* or S04.52*) ",
    "co-occurring with anatomically concordant facial wound ",
    "(S01.1*, S01.3*, S01.4*, S01.5*, or S01.81*) AND external cause code ",
    "(W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "Group 2 = Group 1 + CPT 64864, 64865, 64885, or 64886. ",
    "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

# Co-occurrence Analysis — Group 2 Encounters (Exploratory)

```{r cooccur-explore}
#| cache: false
# ── What other ICD-10-CM and CPT codes appear alongside the Group 2 repair codes?
# This is exploratory / internal only. Produces inline tables and an Excel sheet.

g2_rows <- enc_all %>% filter(had_repair == 1)

# Helper: unpack pipe-joined code strings into a long tidy frame
unpack_codes <- function(df, col, repair_val = 1) {
  df %>%
    filter(had_repair == repair_val, nzchar(.data[[col]])) %>%
    select(discwt, code_str = all_of(col)) %>%
    mutate(code = strsplit(code_str, "\\|")) %>%
    tidyr::unnest(code) %>%
    filter(nzchar(code)) %>%
    select(discwt, code)
}

# ── Diagnosis code co-occurrence ──────────────────────────────────────────────
dx_long <- unpack_codes(enc_all, "dx_raw", repair_val = 1)

# ICD-10 chapter labels for broad grouping (first character)
icd_chapter <- function(code) {
  ch <- substr(code, 1, 1)
  case_when(
    ch == "S" ~ "S — Injury, head/neck/trunk/limb",
    ch == "T" ~ "T — Injury, poisoning, external causes",
    ch == "V" ~ "V — Transport accidents",
    ch == "W" ~ "W — External causes: falls/drowning/contact",
    ch == "X" ~ "X — External causes: other",
    ch == "Y" ~ "Y — External causes: undetermined/sequelae",
    ch == "Z" ~ "Z — Factors influencing health status",
    ch == "G" ~ "G — Nervous system diseases",
    ch == "H" ~ "H — Eye/ear diseases",
    ch == "I" ~ "I — Circulatory diseases",
    ch == "J" ~ "J — Respiratory diseases",
    ch == "K" ~ "K — Digestive diseases",
    ch == "L" ~ "L — Skin diseases",
    ch == "M" ~ "M — Musculoskeletal diseases",
    ch == "N" ~ "N — Genitourinary diseases",
    ch == "O" ~ "O — Pregnancy/childbirth",
    ch == "Q" ~ "Q — Congenital malformations",
    ch == "R" ~ "R — Symptoms/signs NEC",
    ch == "E" ~ "E — Endocrine/nutritional diseases",
    ch == "F" ~ "F — Mental/behavioral disorders",
    ch == "C" ~ "C/D — Neoplasms",
    ch == "D" ~ "C/D — Neoplasms/blood diseases",
    ch == "A" ~ "A/B — Infectious diseases",
    ch == "B" ~ "A/B — Infectious diseases",
    TRUE ~ paste0(ch, " — Other")
  )
}

# Codes that define Group 1/2 — exclude from co-occurrence display
# (they're present by definition in most Group 2 rows)
defining_stems <- c("S045", "S01", "S11",
                    "W25","W26","W27","W45","X99","Y28")
is_defining <- function(code) {
  any(sapply(defining_stems, function(s) str_starts(code, s)))
}

dx_cooccur <- dx_long %>%
  filter(!sapply(code, is_defining)) %>%
  mutate(
    icd3    = substr(code, 1, 3),
    chapter = icd_chapter(code)
  ) %>%
  group_by(code, icd3, chapter) %>%
  summarise(
    n_unweighted = n(),
    n_weighted   = round(sum(discwt, na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  arrange(desc(n_unweighted)) %>%
  slice_head(n = 40) %>%
  mutate(
    pct_of_g2_unw = round(100 * n_unweighted / nrow(g2_rows), 1),
    pct_of_g2_w   = round(100 * n_weighted   /
                            round(total_repair_w), 1)
  ) %>%
  select(
    `ICD-10-CM Code`  = code,
    `3-Char Category` = icd3,
    Chapter           = chapter,
    `Unweighted N`    = n_unweighted,
    `Weighted N`      = n_weighted,
    `% of Group 2 (unweighted)` = pct_of_g2_unw,
    `% of Group 2 (weighted)`   = pct_of_g2_w
  )

message("Top co-occurring DX codes in Group 2 — total Group 2 rows: ",
        nrow(g2_rows))

dx_cooccur %>%
  kable(caption = paste0(
    "Top 40 co-occurring ICD-10-CM diagnosis codes in Group 2 encounters ",
    "(facial nerve repair, CPT 64864/64865/64885/64886). ",
    "Excludes Group 1/2 defining codes (S04.5*, S01.*, S11.*, ",
    "W25-W27, W45, X99, Y28). ",
    "Percentages are of total Group 2 encounters (unweighted N = ",
    format(nrow(g2_rows), big.mark=","), "). ",
    "HCUP NASS 2016\u20132023. Exploratory — not for manuscript."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE)

# ── CPT code co-occurrence ────────────────────────────────────────────────────
cpt_long <- unpack_codes(enc_all, "cpt_raw", repair_val = 1)

# Exclude the Group 2 repair codes themselves
repair_cpts <- cpt_codes   # c("64864","64865","64885","64886")

cpt_cooccur <- cpt_long %>%
  filter(!code %in% repair_cpts, nzchar(code)) %>%
  group_by(`CPT Code` = code) %>%
  summarise(
    `Unweighted N` = n(),
    `Weighted N`   = round(sum(discwt, na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    `% of Group 2 (unweighted)` = round(100 * `Unweighted N` / nrow(g2_rows), 1),
    `% of Group 2 (weighted)`   = round(100 * `Weighted N`   /
                                          round(total_repair_w), 1)
  ) %>%
  arrange(desc(`Unweighted N`)) %>%
  slice_head(n = 30)

cpt_cooccur %>%
  kable(caption = paste0(
    "Top 30 co-occurring CPT codes in Group 2 encounters. ",
    "Excludes defining repair codes (64864, 64865, 64885, 64886). ",
    "HCUP NASS 2016\u20132023. Exploratory — not for manuscript."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE)

# ── S04.5 subcode breakdown ───────────────────────────────────────────────────
s045_long <- unpack_codes(enc_all, "dx_raw", repair_val = 1) %>%
  filter(str_starts(code, "S045")) %>%
  group_by(`S04.5 Subcode` = code) %>%
  summarise(
    `Unweighted N` = n(),
    `Weighted N`   = round(sum(discwt, na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    `% of Group 2 (unweighted)` = round(100 * `Unweighted N` / nrow(g2_rows), 1)
  ) %>%
  arrange(desc(`Unweighted N`))

s045_long %>%
  kable(caption = paste0(
    "S04.5* subcode breakdown in Group 2 encounters. ",
    "Shows which specific facial nerve injury codes are most common ",
    "among repaired cases. Note: one encounter may carry multiple subcodes. ",
    "Sequela codes (7th char S) excluded from Group 1 as of v1.4. ",
    "HCUP NASS 2016\u20132023. Exploratory."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)

# ── 7th character distribution of S04.5 codes in Group 1 ─────────────────────
# Confirms sequela exclusion is working and shows encounter-type distribution.
# ICD-10-CM 7th characters for trauma codes:
#   A = initial encounter (active/definitive treatment)
#   D = subsequent encounter (routine care after active treatment)
#   S = sequela (late effect — excluded from Group 1 in v1.4)
s045_7th <- enc_all %>%
  select(discwt, dx_raw) %>%
  mutate(code = strsplit(dx_raw, "\\|")) %>%
  tidyr::unnest(code) %>%
  filter(str_starts(code, "S045"), nzchar(code)) %>%
  mutate(
    `7th Character` = substr(code, nchar(code), nchar(code)),
    `Encounter Type` = case_when(
      `7th Character` == "A" ~ "A — Initial encounter (active treatment)",
      `7th Character` == "D" ~ "D — Subsequent encounter (routine care)",
      `7th Character` == "S" ~ "S — Sequela (late effect) *** SHOULD BE ZERO ***",
      TRUE ~ paste0(`7th Character`, " — Other/unknown")
    )
  ) %>%
  group_by(`7th Character`, `Encounter Type`) %>%
  summarise(
    `Unweighted N (code occurrences)` = n(),
    `Weighted N` = round(sum(discwt, na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  arrange(`7th Character`)

s045_7th %>%
  kable(caption = paste0(
    "7th character distribution of S04.5* codes in Group 1 encounters ",
    "(after v1.4 sequela exclusion). Sequela (S) should be zero — if not, ",
    "the exclusion filter is not working. ",
    "HCUP NASS 2016\u20132023. Exploratory / quality check.")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(which(s045_7th$`7th Character` == "S"),
           bold = TRUE, color = "red",
           background = "#FFF3F3")

# ── Open wound subcode breakdown ──────────────────────────────────────────────
wound_long <- unpack_codes(enc_all, "dx_raw", repair_val = 1) %>%
  filter(str_starts(code, "S01") | str_starts(code, "S11")) %>%
  group_by(`Open Wound Code` = code) %>%
  summarise(
    `Unweighted N` = n(),
    `Weighted N`   = round(sum(discwt, na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    `% of Group 2 (unweighted)` = round(100 * `Unweighted N` / nrow(g2_rows), 1),
    `Body region` = case_when(
      str_starts(`Open Wound Code`, "S010") ~ "S01.0 — scalp",
      str_starts(`Open Wound Code`, "S011") ~ "S01.1 — eyelid/periocular",
      str_starts(`Open Wound Code`, "S012") ~ "S01.2 — nose",
      str_starts(`Open Wound Code`, "S013") ~ "S01.3 — ear",
      str_starts(`Open Wound Code`, "S014") ~ "S01.4 — cheek/TMJ area",
      str_starts(`Open Wound Code`, "S015") ~ "S01.5 — lip/oral cavity",
      str_starts(`Open Wound Code`, "S018") ~ "S01.8 — other/unspecified head",
      str_starts(`Open Wound Code`, "S019") ~ "S01.9 — head, unspecified",
      str_starts(`Open Wound Code`, "S11")  ~ "S11.* — neck",
      TRUE ~ "Other"
    )
  ) %>%
  arrange(desc(`Unweighted N`)) %>%
  slice_head(n = 30) %>%
  select(`Open Wound Code`, `Body region`, `Unweighted N`,
         `Weighted N`, `% of Group 2 (unweighted)`)

wound_long %>%
  kable(caption = paste0(
    "Open wound subcode (S01.*, S11.*) breakdown in Group 2 encounters. ",
    "Shows anatomic distribution of co-occurring wound codes in repaired cases. ",
    "HCUP NASS 2016\u20132023. Exploratory."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE)
```

# Mechanism of Injury (Group 1)

```{r cooccur-mechanism}
#| cache: false
# Classify each Group 1 encounter by mechanism using V/W/X/Y external cause
# codes carried in dx_raw. Encounters can carry multiple external cause codes
# but each encounter is assigned to its most specific mechanism category
# using a hierarchy: assault > self-harm > transport > cutting > fall >
# animal > undetermined > other > none.

mechanism_categories <- c(
  "Assault by sharp object (X99)",
  "Assault other (X9*, Y0*)",
  "Self-harm (X71\u2013X83)",
  "Transport-related (V01\u2013V99)",
  "Cutting/piercing accident (W25\u2013W27, W45)",
  "Fall (W00\u2013W19)",
  "Animal-related (W50\u2013W64)",
  "Undetermined intent (Y10\u2013Y34)",
  "Other external cause",
  "No external cause coded"
)

classify_mechanism <- function(codes) {
  if (length(codes) == 0) return("No external cause coded")
  # Hierarchy: most specific / clinically meaningful first
  if (any(str_starts(codes, "X99")))
    return("Assault by sharp object (X99)")
  if (any(str_detect(codes, "^(X9[0-8]|Y0[0-9])")))
    return("Assault other (X9*, Y0*)")
  if (any(str_detect(codes, "^(X7[1-9]|X8[0-3])")))
    return("Self-harm (X71\u2013X83)")
  if (any(str_detect(codes, "^V[0-9]{2}")))
    return("Transport-related (V01\u2013V99)")
  if (any(str_detect(codes, "^(W2[5-7]|W45)")))
    return("Cutting/piercing accident (W25\u2013W27, W45)")
  if (any(str_detect(codes, "^W[01][0-9]")))
    return("Fall (W00\u2013W19)")
  if (any(str_detect(codes, "^W[56][0-9]")))
    return("Animal-related (W50\u2013W64)")
  if (any(str_detect(codes, "^Y[12][0-9]|^Y3[0-4]")))
    return("Undetermined intent (Y10\u2013Y34)")
  # Any other external cause code (W*, X*, Y*)
  if (any(str_detect(codes, "^[WXY][0-9]")))
    return("Other external cause")
  return("No external cause coded")
}

mechanism_assignments <- enc_all %>%
  mutate(
    code_list = strsplit(dx_raw, "\\|"),
    mechanism = vapply(code_list, classify_mechanism, character(1))
  ) %>%
  select(year, had_repair, discwt, mechanism)

mechanism_summary <- mechanism_assignments %>%
  group_by(mechanism) %>%
  summarise(
    `Unweighted N`            = n(),
    `Weighted N`              = round(sum(discwt, na.rm=TRUE)),
    `Repair (unweighted N)`   = sum(had_repair == 1),
    `Repair (weighted N)`     = round(sum(discwt[had_repair == 1], na.rm=TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    `% of Group 1 (unweighted)` = round(100 * `Unweighted N` / nrow(enc_all), 1),
    `% of Group 1 (weighted)`   = round(100 * `Weighted N` /
                                          round(total_g1_w), 1),
    `Repair rate (%)`           = round(100 * `Repair (weighted N)` /
                                          `Weighted N`, 1)
  ) %>%
  mutate(mechanism = factor(mechanism, levels = mechanism_categories)) %>%
  arrange(mechanism) %>%
  mutate(mechanism = as.character(mechanism))

# Apply HCUP cell suppression for any unweighted < 11
mechanism_display <- mechanism_summary %>%
  mutate(across(c(`Unweighted N`, `Weighted N`,
                   `Repair (unweighted N)`, `Repair (weighted N)`,
                   `% of Group 1 (unweighted)`, `% of Group 1 (weighted)`,
                   `Repair rate (%)`),
                ~ ifelse(`Unweighted N` < 11, NA, .))) %>%
  rename(`Mechanism of injury` = mechanism)

mechanism_display %>%
  kable(caption = paste0(
    "Mechanism of injury distribution in Group 1 encounters, classified using ",
    "ICD-10-CM external cause codes carried on the encounter. Each encounter ",
    "is assigned to its most specific mechanism category using a clinical ",
    "hierarchy (assault > self-harm > transport > cutting > fall > animal > ",
    "undetermined > other > none). HCUP NASS 2016\u20132023. Per HCUP DUA, ",
    "categories with unweighted N < 11 are suppressed (NA)."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE)

# Inline summary metrics for the Discussion paragraph
n_with_mechanism <- sum(mechanism_assignments$mechanism != "No external cause coded")
pct_with_mechanism <- round(100 * n_with_mechanism / nrow(enc_all), 1)
message("Encounters with any external cause coded: ",
        n_with_mechanism, " (", pct_with_mechanism, "%)")
```

# Phase 1 vs Phase 2 Comparison (Internal Quality Check)

```{r phase1-vs-phase2}
#| cache: false
# Internal investigator quality check only — does NOT appear in manuscript.
# Confirms Phase 2 produces a more complete cohort and that direction of
# significant predictors is consistent across definitions.

# Helper: pull a Phase 2 OR row by its display label
get_or_row <- function(label) {
  key <- names(term_map)[term_map == label]
  r   <- or_df %>% filter(term %in% key)
  if (nrow(r) == 0) return("not estimated")
  paste0(sprintf("%.2f",r$OR[1])," (95% CI ",
         sprintf("%.2f",r$lower[1]),"\u2013",
         sprintf("%.2f",r$upper[1]),"), p=", r$p_fmt[1])
}

phase2_overall_rate <- round(100*total_repair_w/total_g1_w, 1)
phase2_iptw_text    <- paste0(eff_pct, "%, p=", iptw_p)
phase2_teach_text   <- paste0("p=", teach_ttest_norep)

phase_compare <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate",
    "Age 55\u201365 OR",
    "Age 65+ OR",
    "Midwest OR",
    "West OR",
    "IPTW charge increase with repair",
    "Teaching vs nonteaching charge difference (no repair)"
  ),
  `Phase 1 Value` = c(
    "1,656",
    "2,276",
    "488",
    "664",
    "1,168",
    "1,613",
    "29.2%",
    "0.32 (95% CI 0.18\u20130.57), p<0.001",
    "0.41 (95% CI 0.20\u20130.85), p=0.018",
    "1.63 (95% CI 1.10\u20132.41), p=0.015",
    "1.81 (95% CI 1.05\u20133.10), p=0.033",
    "34.9%, p<0.001",
    "Significant, p<0.001"
  ),
  `Phase 2 Value` = c(
    format(nrow(enc_all), big.mark=","),
    format(round(total_g1_w), big.mark=","),
    format(sum(enc_all$had_repair==1), big.mark=","),
    format(round(total_repair_w), big.mark=","),
    format(sum(enc_all$had_repair==0), big.mark=","),
    format(round(total_norepair_w), big.mark=","),
    paste0(phase2_overall_rate, "%"),
    get_or_row("55-65"),
    get_or_row("65+"),
    get_or_row("Midwest"),
    get_or_row("West"),
    phase2_iptw_text,
    phase2_teach_text
  )
)

phase_compare %>%
  kable(caption=paste0(
    "Phase 1 vs Phase 2 Comparison \u2014 Internal Quality Check. ",
    "Investigator use only; does not appear in the Phase 2 manuscript.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE)

# Sanity check on the bias-correction direction
if (phase2_overall_rate > 29.2) {
  warning("Phase 2 overall repair rate (", phase2_overall_rate,
          "%) is HIGHER than Phase 1 (29.2%). Per Phase 2 design rationale, ",
          "this should not happen — review extraction logic.")
} else {
  message("Phase 2 overall repair rate = ", phase2_overall_rate,
          "% (Phase 1 = 29.2%); bias-correction direction is as expected.")
}
```

# Excel Output

```{r excel-output}
#| cache: false
wb <- createWorkbook()

title_s  <- createStyle(fontName="Arial",fontSize=12,textDecoration="bold")
header_s <- createStyle(fontName="Arial",fontSize=11,textDecoration="bold",
                        border="TopBottom",borderStyle="medium",
                        halign="center",wrapText=TRUE)
group_s  <- createStyle(fontName="Arial",fontSize=11,textDecoration="bold")
data_s   <- createStyle(fontName="Arial",fontSize=11,halign="left")
center_s <- createStyle(fontName="Arial",fontSize=11,halign="center")
ref_s    <- createStyle(fontName="Arial",fontSize=11,
                        fontColour="#808080",halign="center")
bot_s    <- createStyle(fontName="Arial",fontSize=11,
                        border="Bottom",borderStyle="medium")
note_s   <- createStyle(fontName="Arial",fontSize=9,
                        wrapText=TRUE,valign="top")

write_table_sheet <- function(wb, sname, title, tbl,
                              grp_id_col, footnote, ref_col=NULL) {
  n <- ncol(tbl)
  addWorksheet(wb, sname)
  writeData(wb, sname, title, startRow=1, startCol=1)
  addStyle(wb, sname, title_s, rows=1, cols=1)
  writeData(wb, sname, as.data.frame(t(names(tbl))),
            startRow=2, startCol=1, colNames=FALSE)
  addStyle(wb, sname, header_s, rows=2, cols=1:n, gridExpand=TRUE)
  grp_labels <- tbl[[1]][is.na(tbl[[grp_id_col]]) &
                          !str_starts(tbl[[1]],"   ")]
  for (i in seq_len(nrow(tbl))) {
    re     <- i + 2
    cv     <- tbl[[1]][i]
    is_grp <- cv %in% grp_labels
    is_ref <- !is.null(ref_col) && !is.na(tbl[[ref_col]][i]) &&
               tbl[[ref_col]][i] == "Reference"
    writeData(wb, sname, tbl[i,], startRow=re, startCol=1, colNames=FALSE)
    if (is_grp) {
      addStyle(wb, sname, group_s, rows=re, cols=1:n, gridExpand=TRUE)
    } else {
      addStyle(wb, sname, data_s, rows=re, cols=1)
      addStyle(wb, sname, if(is_ref) ref_s else center_s,
               rows=re, cols=2:n, gridExpand=TRUE)
    }
  }
  lr <- nrow(tbl) + 2
  addStyle(wb, sname, bot_s, rows=lr, cols=1:n, gridExpand=TRUE)
  setColWidths(wb, sname, cols=1, widths=38)
  setColWidths(wb, sname, cols=2:n, widths=20)
  fn <- lr + 2
  writeData(wb, sname, footnote, startRow=fn, startCol=1)
  mergeCells(wb, sname, cols=1:n, rows=fn)
  addStyle(wb, sname, note_s, rows=fn, cols=1)
  setRowHeights(wb, sname, rows=fn, heights=70)
}

# ── Sheet 1 ──────────────────────────────────────────────────────────────────
write_table_sheet(wb, "Sheet 1",
  paste0("National Estimates of Patient and Hospital Characteristics by ",
         "Facial Nerve Repair Status, Including Repair Rates"),
  sheet1, grp_id_col="No Repair (n, %)",
  footnote=paste0(
    "HCUP NASS 2016\u20132023. Total = All patients with traumatic facial nerve ",
    "laceration nationally. Group 1 (Phase 2) = encounters meeting Criterion A ",
    "(S04.5* co-occurring with S01.* or S11.*, all initial or subsequent ",
    "encounters / 7th character A or D) OR Criterion B (CPT 64864/64865/64885/64886 ",
    "co-occurring with S01.* or S11.*, A or D). Sequela codes (7th char S) excluded ",
    "from Group 1. External cause codes (W25\u2013W27, W45, X99, Y28, V01\u2013V99) ",
    "are not part of primary inclusion; see sensitivity analyses for higher-specificity ",
    "subsets. Group 2 (Repair) = CPT 64864, 64865, 64885, or 64886 within Group 1. ",
    "Group 3 (No Repair) = Group 1 minus Group 2. All values are weighted ",
    "national estimates using DISCWT discharge weight. Repair Rate = Weighted repair ",
    "encounters / Weighted Group 1 encounters x 100. Age mean \u00b1 SD estimated ",
    "using age group midpoints with discharge weights. NOTE: HOSP_REGION was not ",
    "available in the 2023 NASS Hospital file; regional estimates reflect ",
    "2016\u20132022 data only."))

# ── Sheet 2 ──────────────────────────────────────────────────────────────────
write_table_sheet(wb, "Sheet 2",
  "Multivariable Logistic Regression for Odds of Facial Nerve Repair",
  sheet2, grp_id_col="Adjusted OR (95% CI)",
  ref_col="Adjusted OR (95% CI)",
  footnote=paste0(
    "Survey-weighted multivariable logistic regression (svyglm, quasibinomial). ",
    "HCUP NASS 2016\u20132023. Reference categories: Male, Age 18-34, White, ",
    "76th-100th income quartile, Private insurance, South region, Urban, ",
    "Nonteaching hospital. OR = odds ratio; CI = 95% confidence interval. ",
    "Significant predictors (p<0.05): ", sig_pred_text, "."))

# ── Sheet 4 ──────────────────────────────────────────────────────────────────
addWorksheet(wb, "Sheet 4")
writeData(wb, "Sheet 4", "Repair Rates by Year", startRow=1, startCol=1)
addStyle(wb, "Sheet 4", title_s, rows=1, cols=1)
writeData(wb, "Sheet 4", sheet4, startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sheet 4", header_s, rows=3, cols=1:6, gridExpand=TRUE)
addStyle(wb, "Sheet 4", data_s, rows=4:(nrow(sheet4)+3), cols=1:6, gridExpand=TRUE)
setColWidths(wb, "Sheet 4", cols=1, widths=10)
setColWidths(wb, "Sheet 4", cols=2:6, widths=28)
fn4 <- nrow(sheet4)+5
writeData(wb, "Sheet 4",
  paste0("HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight. ",
       "Repair rate = Weighted repair / Weighted total injury x 100. ",
       "Chi-square test for difference in repair rates across years: p = ", year_p, ". ",
       "The apparent decline in repair rate between 2016\u20132018 and 2019\u20132023 ",
       "likely reflects increasing completeness of ICD-10-CM injury coding ",
       "(particularly of non-operative encounters that contribute to the denominator) ",
       "rather than a true change in surgical practice. ICD-10-CM was implemented ",
       "October 2015, and HCUP documentation supports gradual improvement in ",
       "coding completeness over the early years of the new code set. ",
       "The 2019\u20132023 estimates provide the most reliable contemporary repair rate."),
  startRow=fn4, startCol=1)
mergeCells(wb, "Sheet 4", cols=1:6, rows=fn4)
addStyle(wb, "Sheet 4", note_s, rows=fn4, cols=1)
setRowHeights(wb, "Sheet 4", rows=fn4, heights=110)

# ── Sheet 5 ──────────────────────────────────────────────────────────────────
addWorksheet(wb, "Sheet 5")
writeData(wb, "Sheet 5", "Charge Analysis: Repair vs No Repair",
          startRow=1, startCol=1)
addStyle(wb, "Sheet 5", title_s, rows=1, cols=1)
writeData(wb, "Sheet 5", sheet5, startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sheet 5", header_s, rows=3, cols=1:5, gridExpand=TRUE)
grp5 <- sheet5$Characteristic[is.na(sheet5$`No Repair Median (IQR)`) &
                               !str_starts(sheet5$Characteristic,"   ")]
for (i in seq_len(nrow(sheet5))) {
  re <- i+3
  if (sheet5$Characteristic[i] %in% grp5) {
    addStyle(wb, "Sheet 5", group_s, rows=re, cols=1:5, gridExpand=TRUE)
  } else {
    addStyle(wb, "Sheet 5", data_s,   rows=re, cols=1)
    addStyle(wb, "Sheet 5", center_s, rows=re, cols=2:5, gridExpand=TRUE)
  }
}
addStyle(wb, "Sheet 5", bot_s, rows=nrow(sheet5)+3, cols=1:5, gridExpand=TRUE)
setColWidths(wb, "Sheet 5", cols=1, widths=35)
setColWidths(wb, "Sheet 5", cols=2:5, widths=28)
fn5 <- nrow(sheet5)+5
writeData(wb, "Sheet 5",
  paste0("Median (IQR) weighted national estimates. HCUP NASS 2016\u20132023. ",
         "Insurance type p-values: survey-weighted ANOVA on log-charges within ",
         "no-repair group and within repair group separately. ",
         "Teaching status p-values: survey-weighted t-test on log-charges ",
         "comparing Nonteaching vs Teaching within each repair group."),
  startRow=fn5, startCol=1)
mergeCells(wb, "Sheet 5", cols=1:5, rows=fn5)
addStyle(wb, "Sheet 5", note_s, rows=fn5, cols=1)
setRowHeights(wb, "Sheet 5", rows=fn5, heights=55)

# ── NIS 2023 Sheet ────────────────────────────────────────────────────────────
addWorksheet(wb, "NIS 2023 Inpatient Results")
writeData(wb, "NIS 2023 Inpatient Results",
          "NIS 2023 Facial Nerve Repair \u2014 Inpatient Database, Traumatic Classification",
          startRow=1, startCol=1)
addStyle(wb, "NIS 2023 Inpatient Results", title_s, rows=1, cols=1)
mergeCells(wb, "NIS 2023 Inpatient Results", cols=1:4, rows=1)
writeData(wb, "NIS 2023 Inpatient Results", nis_sheet,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "NIS 2023 Inpatient Results", header_s, rows=3, cols=1:4,
         gridExpand=TRUE)

# Per-row styling: row index in nis_sheet drives the visual treatment.
# Row 1  = grand total (all repair) — yellow, bold
# Row 3  = CNS-miscoded all-repair — italic purple
# Row 4  = traumatic union — green, bold
# Row 8  = CNS-miscoded traumatic — italic purple
# Row 10 = non-traumatic — light purple, bold
for (i in seq_len(nrow(nis_sheet))) {
  re <- i + 3
  if (i == 1) {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fgFill="#FCF3CF", textDecoration="bold")
  } else if (i == 4) {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fgFill="#D5F5E3", textDecoration="bold")
  } else if (i == 10) {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fgFill="#E8DAEF", textDecoration="bold")
  } else if (i %in% c(3, 8)) {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fontColour="#6C3483", textDecoration="italic")
  } else {
    sty <- createStyle(fontName="Arial", fontSize=11)
  }
  addStyle(wb, "NIS 2023 Inpatient Results", sty,
           rows=re, cols=1:4, gridExpand=TRUE)
}
addStyle(wb, "NIS 2023 Inpatient Results", bot_s,
         rows=nrow(nis_sheet)+3, cols=1:4, gridExpand=TRUE)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=1, widths=22)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=2, widths=58)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=3:4, widths=22)
fn_nis <- nrow(nis_sheet)+5
writeData(wb, "NIS 2023 Inpatient Results",
  paste0("Source: completed three-phase NIS 2023 analysis (v1.0 frequency, ",
         "v1.1 miscoding confirmation, v1.2 trauma classification). ",
         "Survey-weighted national estimates use NIS_STRATUM clustering, ",
         "HOSP_NIS sampling unit, and DISCWT discharge weight (nest=TRUE, ",
         "options(survey.lonely.psu=\"adjust\")). Total inpatient burden of ",
         "any facial nerve repair = ", nis_total_w, " weighted, of which ",
         nis_cns_w, " (", nis_cns_share, "%) were coded to the CNS section ",
         "(00QM) instead of the correct PNS section (01QM). The miscoding ",
         "hypothesis was confirmed in v1.1 by diagnosis-profile analysis: ",
         nis_miscode_confirmed_pct, "% of CNS-miscoded cases carry a ",
         "facial-nerve-consistent diagnosis (G51.x, S04.5x, C07/D11.0, D33.3, ",
         "S02.1x, S01.4/8x, G52.x), exceeding the 70% confirmation threshold. ",
         "The traumatic subset = ", nis_trauma_w, " weighted (", nis_trauma_share,
         "% of all repair), defined as the deduplicated union of S04.5, ",
         "S01.4/S01.8, and external cause codes V01\u2013Y99. External cause ",
         "codes captured ~12 cases not identified by diagnosis codes alone. ",
         "Of traumatic cases, ", nis_trauma_miscode, "% are CNS-miscoded \u2014 ",
         "a query restricted to anatomically correct PCS codes would recover ",
         "only ~5 of 130 weighted traumatic encounters. Per HCUP DUA, individual ",
         "unweighted counts <11 are suppressed (NA) in this published table. ",
         "Italicized purple rows: CNS-miscoded entries. Yellow bold: all-repair ",
         "grand total. Green bold: traumatic union. Light purple bold: ",
         "non-traumatic remainder. The 94.7% CNS/PNS miscoding pattern is a ",
         "novel, confirmed administrative-data-quality finding to be reported ",
         "in the manuscript Discussion. NIS findings are descriptive only; ",
         "NASS remains the primary database for this analysis."),
  startRow=fn_nis, startCol=1)
mergeCells(wb, "NIS 2023 Inpatient Results", cols=1:4, rows=fn_nis)
addStyle(wb, "NIS 2023 Inpatient Results", note_s, rows=fn_nis, cols=1)
setRowHeights(wb, "NIS 2023 Inpatient Results", rows=fn_nis, heights=210)

# ── IPTW Sheet ────────────────────────────────────────────────────────────────
addWorksheet(wb, "IPTW Charges Analysis")
writeData(wb, "IPTW Charges Analysis",
          "IPTW Analysis: Effect of Repair on Total Charges",
          startRow=1, startCol=1)
addStyle(wb, "IPTW Charges Analysis", title_s, rows=1, cols=1)
writeData(wb, "IPTW Charges Analysis", iptw_excel,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "IPTW Charges Analysis", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "IPTW Charges Analysis", data_s,
         rows=4:(nrow(iptw_excel)+3), cols=1:2, gridExpand=TRUE)
setColWidths(wb, "IPTW Charges Analysis", cols=1, widths=40)
setColWidths(wb, "IPTW Charges Analysis", cols=2, widths=60)
fn_iptw <- nrow(iptw_excel)+5
writeData(wb, "IPTW Charges Analysis",
  paste0("IPTW estimated using propensity score weighting (WeightIt, ATE estimand). ",
         "Outcome modeled as log(total charges) via survey-weighted linear regression. ",
         "Effect reflects percentage change in charges associated with repair after ",
         "balancing all measured confounders. HCUP NASS 2016\u20132023."),
  startRow=fn_iptw, startCol=1)
mergeCells(wb, "IPTW Charges Analysis", cols=1:2, rows=fn_iptw)
addStyle(wb, "IPTW Charges Analysis", note_s, rows=fn_iptw, cols=1)
setRowHeights(wb, "IPTW Charges Analysis", rows=fn_iptw, heights=55)

# ── Phase 1 vs Phase 2 Comparison (Excel only — internal QC) ─────────────────
addWorksheet(wb, "Phase 1 vs Phase 2 Comparison")
writeData(wb, "Phase 1 vs Phase 2 Comparison",
          "Phase 1 vs Phase 2 Comparison \u2014 Internal Quality Check",
          startRow=1, startCol=1)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", title_s, rows=1, cols=1)
mergeCells(wb, "Phase 1 vs Phase 2 Comparison", cols=1:3, rows=1)
writeData(wb, "Phase 1 vs Phase 2 Comparison", phase_compare,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", header_s, rows=3, cols=1:3,
         gridExpand=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", data_s,
         rows=4:(nrow(phase_compare)+3), cols=1:3, gridExpand=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", bot_s,
         rows=nrow(phase_compare)+3, cols=1:3, gridExpand=TRUE)
setColWidths(wb, "Phase 1 vs Phase 2 Comparison", cols=1, widths=42)
setColWidths(wb, "Phase 1 vs Phase 2 Comparison", cols=2:3, widths=38)
fn_pc <- nrow(phase_compare)+5
writeData(wb, "Phase 1 vs Phase 2 Comparison",
  paste0("Investigator quality check only \u2014 does NOT appear in the Phase 2 ",
         "manuscript. Confirms Phase 2 produces a more complete cohort and that ",
         "the direction of significant predictors is consistent across definitions. ",
         "Phase 2 overall repair rate is expected to be lower than 29.2%; if higher, ",
         "review extraction logic. NOTE: HOSP_REGION not available in the 2023 NASS ",
         "Hospital file; regional ORs reflect 2016\u20132022 data only."),
  startRow=fn_pc, startCol=1)
mergeCells(wb, "Phase 1 vs Phase 2 Comparison", cols=1:3, rows=fn_pc)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", note_s, rows=fn_pc, cols=1)
setRowHeights(wb, "Phase 1 vs Phase 2 Comparison", rows=fn_pc, heights=70)

# ── Sensitivity Analysis 1 ───────────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 1")
writeData(wb, "Sensitivity Analysis 1",
          "Sensitivity Analysis 1 \u2014 Criterion A Only",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 1", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 1", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 1", sens1_tbl,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 1", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 1", data_s,
         rows=4:(nrow(sens1_tbl)+3), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 1", bot_s,
         rows=nrow(sens1_tbl)+3, cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 1", cols=1, widths=32)
setColWidths(wb, "Sensitivity Analysis 1", cols=2, widths=22)
fn_s1 <- nrow(sens1_tbl)+5
writeData(wb, "Sensitivity Analysis 1",
  paste0("Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, ",
         "excluding the external-cause-only path AND excluding the Criterion B ",
         "procedural pathway. HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."),
  startRow=fn_s1, startCol=1)
mergeCells(wb, "Sensitivity Analysis 1", cols=1:2, rows=fn_s1)
addStyle(wb, "Sensitivity Analysis 1", note_s, rows=fn_s1, cols=1)
setRowHeights(wb, "Sensitivity Analysis 1", rows=fn_s1, heights=55)

# ── Sensitivity Analysis 2 ───────────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 2")
writeData(wb, "Sensitivity Analysis 2",
          "Sensitivity Analysis 2 \u2014 Highest Mechanistic Specificity",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 2", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 2", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 2", sens2_tbl,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 2", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 2", data_s,
         rows=4:(nrow(sens2_tbl)+3), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 2", bot_s,
         rows=nrow(sens2_tbl)+3, cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 2", cols=1, widths=32)
setColWidths(wb, "Sensitivity Analysis 2", cols=2, widths=22)
fn_s2 <- nrow(sens2_tbl)+5
writeData(wb, "Sensitivity Analysis 2",
  paste0("Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
         "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
         "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight."),
  startRow=fn_s2, startCol=1)
mergeCells(wb, "Sensitivity Analysis 2", cols=1:2, rows=fn_s2)
addStyle(wb, "Sensitivity Analysis 2", note_s, rows=fn_s2, cols=1)
setRowHeights(wb, "Sensitivity Analysis 2", rows=fn_s2, heights=55)

# ── Sensitivity Analysis 3 ───────────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 3")
writeData(wb, "Sensitivity Analysis 3",
          "Sensitivity Analysis 3 \u2014 Maximum Specificity",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 3", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 3", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 3", sens3_tbl,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 3", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 3", data_s,
         rows=4:(nrow(sens3_tbl)+3), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 3", bot_s,
         rows=nrow(sens3_tbl)+3, cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 3", cols=1, widths=32)
setColWidths(wb, "Sensitivity Analysis 3", cols=2, widths=22)
fn_s3 <- nrow(sens3_tbl)+5
writeData(wb, "Sensitivity Analysis 3",
  paste0("Laterality-specified facial nerve injury (S04.51* or S04.52*) ",
         "co-occurring with anatomically concordant facial wound (S01.1*, S01.3*, ",
         "S01.4*, S01.5*, or S01.81*) AND external cause code (W25\u2013W27, W45, ",
         "X99, Y28, V01\u2013V99). This is the maximum specificity definition ",
         "feasible in ICD-10-CM administrative data. Note: external cause code ",
         "completeness varies by institution and year; cohort size reflects ",
         "availability of all three code types simultaneously. ",
         "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."),
  startRow=fn_s3, startCol=1)
mergeCells(wb, "Sensitivity Analysis 3", cols=1:2, rows=fn_s3)
addStyle(wb, "Sensitivity Analysis 3", note_s, rows=fn_s3, cols=1)
setRowHeights(wb, "Sensitivity Analysis 3", rows=fn_s3, heights=80)

# ── Co-occurrence Sheet (Exploratory) ────────────────────────────────────────
addWorksheet(wb, "Co-occurrence (Exploratory)")
writeData(wb, "Co-occurrence (Exploratory)",
          "Co-occurring ICD-10-CM Codes in Group 2 — Exploratory Analysis",
          startRow=1, startCol=1)
addStyle(wb, "Co-occurrence (Exploratory)", title_s, rows=1, cols=1)
mergeCells(wb, "Co-occurrence (Exploratory)", cols=1:7, rows=1)

# Section A: top DX codes
writeData(wb, "Co-occurrence (Exploratory)",
          "A. Top 40 Co-occurring Diagnosis Codes (excluding defining codes)",
          startRow=3, startCol=1)
addStyle(wb, "Co-occurrence (Exploratory)", group_s, rows=3, cols=1)
writeData(wb, "Co-occurrence (Exploratory)", dx_cooccur,
          startRow=4, startCol=1, colNames=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", header_s,
         rows=4, cols=1:7, gridExpand=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", data_s,
         rows=5:(nrow(dx_cooccur)+4), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", center_s,
         rows=5:(nrow(dx_cooccur)+4), cols=3:7, gridExpand=TRUE)
setColWidths(wb, "Co-occurrence (Exploratory)",
             cols=1:7, widths=c(18,14,36,14,14,22,22))

# Section B: CPT codes
cpt_start <- nrow(dx_cooccur) + 7
writeData(wb, "Co-occurrence (Exploratory)",
          "B. Top 30 Co-occurring CPT Codes (excluding repair codes 64864/64865/64885/64886)",
          startRow=cpt_start, startCol=1)
addStyle(wb, "Co-occurrence (Exploratory)", group_s, rows=cpt_start, cols=1)
writeData(wb, "Co-occurrence (Exploratory)", cpt_cooccur,
          startRow=cpt_start+1, startCol=1, colNames=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", header_s,
         rows=cpt_start+1, cols=1:5, gridExpand=TRUE)
if (nrow(cpt_cooccur) > 0) {
  addStyle(wb, "Co-occurrence (Exploratory)", data_s,
           rows=(cpt_start+2):(cpt_start+1+nrow(cpt_cooccur)), cols=1,
           gridExpand=TRUE)
  addStyle(wb, "Co-occurrence (Exploratory)", center_s,
           rows=(cpt_start+2):(cpt_start+1+nrow(cpt_cooccur)), cols=2:5,
           gridExpand=TRUE)
}

# Section C: S04.5 subcode breakdown
s045_start <- cpt_start + nrow(cpt_cooccur) + 4
writeData(wb, "Co-occurrence (Exploratory)",
          "C. S04.5* Subcode Breakdown in Group 2",
          startRow=s045_start, startCol=1)
addStyle(wb, "Co-occurrence (Exploratory)", group_s, rows=s045_start, cols=1)
writeData(wb, "Co-occurrence (Exploratory)", s045_long,
          startRow=s045_start+1, startCol=1, colNames=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", header_s,
         rows=s045_start+1, cols=1:4, gridExpand=TRUE)
if (nrow(s045_long) > 0) {
  addStyle(wb, "Co-occurrence (Exploratory)", data_s,
           rows=(s045_start+2):(s045_start+1+nrow(s045_long)), cols=1,
           gridExpand=TRUE)
  addStyle(wb, "Co-occurrence (Exploratory)", center_s,
           rows=(s045_start+2):(s045_start+1+nrow(s045_long)), cols=2:4,
           gridExpand=TRUE)
}

# Section D: Open wound subcode breakdown
wound_start <- s045_start + nrow(s045_long) + 4
writeData(wb, "Co-occurrence (Exploratory)",
          "D. Open Wound Subcode (S01.*/S11.*) Breakdown in Group 2",
          startRow=wound_start, startCol=1)
addStyle(wb, "Co-occurrence (Exploratory)", group_s, rows=wound_start, cols=1)
writeData(wb, "Co-occurrence (Exploratory)", wound_long,
          startRow=wound_start+1, startCol=1, colNames=TRUE)
addStyle(wb, "Co-occurrence (Exploratory)", header_s,
         rows=wound_start+1, cols=1:5, gridExpand=TRUE)
if (nrow(wound_long) > 0) {
  addStyle(wb, "Co-occurrence (Exploratory)", data_s,
           rows=(wound_start+2):(wound_start+1+nrow(wound_long)), cols=1:2,
           gridExpand=TRUE)
  addStyle(wb, "Co-occurrence (Exploratory)", center_s,
           rows=(wound_start+2):(wound_start+1+nrow(wound_long)), cols=3:5,
           gridExpand=TRUE)
}

fn_co <- wound_start + nrow(wound_long) + 3
writeData(wb, "Co-occurrence (Exploratory)",
  paste0("Exploratory analysis — for investigator use only, not for manuscript. ",
         "Co-occurrence defined as any code appearing in the same encounter as ",
         "a Group 2 repair code (CPT 64864, 64865, 64885, or 64886). ",
         "One encounter may contribute multiple codes. ",
         "Weighted N uses DISCWT discharge weight. HCUP NASS 2016\u20132023."),
  startRow=fn_co, startCol=1)
mergeCells(wb, "Co-occurrence (Exploratory)", cols=1:7, rows=fn_co)
addStyle(wb, "Co-occurrence (Exploratory)", note_s, rows=fn_co, cols=1)
setRowHeights(wb, "Co-occurrence (Exploratory)", rows=fn_co, heights=55)

# ── Mechanism of Injury Sheet ────────────────────────────────────────────────
addWorksheet(wb, "Mechanism of Injury")
writeData(wb, "Mechanism of Injury",
          "Mechanism of Injury Distribution \u2014 Group 1 Encounters",
          startRow=1, startCol=1)
addStyle(wb, "Mechanism of Injury", title_s, rows=1, cols=1)
mergeCells(wb, "Mechanism of Injury", cols=1:8, rows=1)
writeData(wb, "Mechanism of Injury", mechanism_display,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Mechanism of Injury", header_s, rows=3, cols=1:8,
         gridExpand=TRUE)
addStyle(wb, "Mechanism of Injury", data_s,
         rows=4:(nrow(mechanism_display)+3), cols=1, gridExpand=TRUE)
addStyle(wb, "Mechanism of Injury", center_s,
         rows=4:(nrow(mechanism_display)+3), cols=2:8, gridExpand=TRUE)
addStyle(wb, "Mechanism of Injury", bot_s,
         rows=nrow(mechanism_display)+3, cols=1:8, gridExpand=TRUE)
setColWidths(wb, "Mechanism of Injury", cols=1, widths=42)
setColWidths(wb, "Mechanism of Injury", cols=2:8, widths=18)
fn_m <- nrow(mechanism_display)+5
writeData(wb, "Mechanism of Injury",
  paste0("Each Group 1 encounter is classified by mechanism using ICD-10-CM ",
         "external cause codes (V01\u2013Y99) carried on the encounter. ",
         "Hierarchy (most specific first): assault by sharp object > assault other > ",
         "self-harm > transport > cutting/piercing accident > fall > animal-related > ",
         "undetermined intent > other external cause > no external cause coded. ",
         "Each encounter contributes to exactly one mechanism category. ",
         "Per HCUP DUA, categories with unweighted N < 11 are suppressed (blank). ",
         pct_with_mechanism, "% of Group 1 encounters carried at least one external ",
         "cause code. HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."),
  startRow=fn_m, startCol=1)
mergeCells(wb, "Mechanism of Injury", cols=1:8, rows=fn_m)
addStyle(wb, "Mechanism of Injury", note_s, rows=fn_m, cols=1)
setRowHeights(wb, "Mechanism of Injury", rows=fn_m, heights=110)

# ── Save Excel ────────────────────────────────────────────────────────────────
xlsx_path <- file.path(xlsx_dir, "NASS Laceration Phase 2 Output 2.0.xlsx")
saveWorkbook(wb, file=xlsx_path, overwrite=TRUE)
message("Excel saved: ", xlsx_path)
```

# GT Tables for Manuscript

```{r gt-export}
#| cache: false

save_gt_to_word <- function(gt_tbl, filename) {
  filepath <- file.path(word_dir, filename)
  gt::gtsave(gt_tbl, filename=filepath)
  message("Saved: ", filename)
}

# ── Sheet 1 gt ───────────────────────────────────────────────────────────────
gt_sheet1 <- sheet1 %>%
  setNames(c("Characteristic","No Repair","Repair","Total","Repair Rate")) %>%
  mutate(across(c(`No Repair`,`Repair`,`Total`,`Repair Rate`),
                ~ifelse(is.na(.), "", .))) %>%
  gt() %>%
  tab_header(title=md("**National Estimates of Patient and Hospital Characteristics by Facial Nerve Repair Status, Including Repair Rates**")) %>%
  tab_footnote(footnote=md(paste0(
    "HCUP NASS 2016\u20132023. Total = All patients with traumatic facial nerve ",
    "laceration nationally. All values are weighted national estimates (DISCWT). ",
    "Repair Rate = Weighted repair / Weighted Group 1 x 100. Age mean \u00b1 SD ",
    "estimated using age group midpoints. NOTE: HOSP_REGION not available in 2023 ",
    "NASS Hospital file; regional estimates reflect 2016\u20132022 only."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_column_labels()) %>%
  tab_style(
    style=cell_text(weight="bold"),
    locations=cells_body(
      rows=(`No Repair`=="" & !str_starts(Characteristic,"   ")) |
            Characteristic=="Total cases" |
            Characteristic=="Age, mean \u00b1 SD"
    )
  ) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic="Characteristic",
             `No Repair`="No Repair (n, %)",
             Repair="Repair (n, %)",
             Total="Total (n)",
             `Repair Rate`="Repair Rate (%)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 2 gt ───────────────────────────────────────────────────────────────
gt_sheet2 <- sheet2 %>%
  setNames(c("Characteristic","Adjusted OR","p value")) %>%
  gt() %>%
  tab_header(title=md("**Multivariable Logistic Regression for Odds of Facial Nerve Repair**")) %>%
  tab_footnote(footnote=md(paste0(
    "Survey-weighted multivariable logistic regression (svyglm, quasibinomial). ",
    "HCUP NASS 2016\u20132023. Reference: Male, Age 18\u201334, White, ",
    "76th\u2013100th income quartile, Private insurance, South, Urban, Nonteaching. ",
    "OR = odds ratio; CI = 95% confidence interval. Significant predictors (p<0.05): ",
    sig_pred_text, "."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_body(
              rows=is.na(`Adjusted OR`) | `Adjusted OR`=="" )) %>%
  tab_style(style=cell_text(color="grey", style="italic"),
            locations=cells_body(
              rows=!is.na(`Adjusted OR`) & `Adjusted OR`=="Reference")) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic="Characteristic",
             `Adjusted OR`="Adjusted OR (95% CI)",
             `p value`="p value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 4 gt ───────────────────────────────────────────────────────────────
gt_sheet4 <- sheet4 %>%
  setNames(c("Year","Injury Unweighted N","Injury Weighted N",
             "Repair Unweighted N","Repair Weighted N","Repair Rate (%)")) %>%
  gt() %>%
  tab_header(title=md("**Repair Rates by Year**")) %>%
  tab_footnote(footnote=md(paste0("HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight. ",
       "Repair rate = Weighted repair / Weighted total injury x 100. ",
       "Chi-square test for difference in repair rates across years: p = ", year_p, ". ",
       "The apparent decline in repair rate between 2016\u20132018 and 2019\u20132023 ",
       "likely reflects increasing completeness of ICD-10-CM injury coding rather than ",
       "a true change in surgical practice. ICD-10-CM was implemented October 2015; the ",
       "2019\u20132023 estimates provide the most reliable contemporary repair rate."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  fmt_number(columns=c(`Injury Weighted N`,`Repair Weighted N`),
             decimals=0, use_seps=TRUE) %>%
  cols_label(`Injury Unweighted N`="Traumatic facial nerve laceration (unweighted N)",
             `Injury Weighted N`  ="Traumatic facial nerve laceration (weighted N)",
             `Repair Unweighted N`="Repair (unweighted N)",
             `Repair Weighted N`  ="Repair (weighted N)",
             `Repair Rate (%)`    ="Repair rate \u2014 weighted (%)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 5 gt ───────────────────────────────────────────────────────────────
gt_sheet5 <- sheet5 %>%
  setNames(c("Characteristic","No Repair Median","Repair Median",
             "p value No Repair","p value Repair")) %>%
  gt() %>%
  tab_header(title=md("**Charge Analysis: Repair vs No Repair**")) %>%
  tab_footnote(footnote=md(paste0(
    "Median (IQR) weighted national estimates. HCUP NASS 2016\u20132023. ",
    "Insurance type p-values: survey-weighted ANOVA on log-charges within ",
    "no-repair group and within repair group separately. Teaching status p-values: ",
    "survey-weighted t-test comparing Nonteaching vs Teaching within each group."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_body(
              rows=is.na(`No Repair Median`) &
                   !str_starts(Characteristic,"   "))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic     ="Characteristic",
             `No Repair Median` ="No Repair Median (IQR)",
             `Repair Median`    ="Repair Median (IQR)",
             `p value No Repair`="p value (no repair: within-group)",
             `p value Repair`   ="p value (repair: within-group)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── NIS 2023 gt ──────────────────────────────────────────────────────────────
gt_nis <- nis_sheet %>%
  setNames(c("Section","Category","Unweighted N","Weighted N")) %>%
  gt(groupname_col="Section") %>%
  tab_header(
    title=md("**NIS 2023 Facial Nerve Repair \u2014 Inpatient Database, Traumatic Classification**"),
    subtitle=md("*HCUP National Inpatient Sample, 2023 \u2014 three-phase analysis*")) %>%
  tab_footnote(footnote=md(paste0(
    "Survey-weighted national estimates (DISCWT, NIS_STRATUM, HOSP_NIS, ",
    "nest=TRUE). Total inpatient burden of any facial nerve repair = ",
    nis_total_w, " weighted, of which ", nis_cns_w, " (", nis_cns_share,
    "%) were coded to the CNS section (00QM) instead of the correct PNS ",
    "section (01QM) under ICD-10-PCS. The miscoding hypothesis was confirmed ",
    "in v1.1 by diagnosis-profile analysis (",
    nis_miscode_confirmed_pct, "% of CNS cases carry a facial-nerve-consistent ",
    "diagnosis, exceeding 70% threshold). Traumatic subset = ",
    nis_trauma_w, " weighted (", nis_trauma_share,
    "% of all repair) defined as the deduplicated union of S04.5, ",
    "S01.4/S01.8, and external cause codes V01\u2013Y99. Of traumatic cases, ",
    nis_trauma_miscode, "% are CNS-miscoded. Per HCUP DUA, unweighted counts ",
    "<11 are suppressed (NA). The 94.7% miscoding pattern is a novel, ",
    "confirmed data-quality finding for the manuscript Discussion. ",
    "NASS remains the primary database; NIS is reported as a descriptive ",
    "comparator only."))) %>%
  tab_style(style=list(cell_text(weight="bold"),
                       cell_fill(color="#FCF3CF")),
            locations=cells_body(rows=1)) %>%        # all-repair total
  tab_style(style=list(cell_text(weight="bold"),
                       cell_fill(color="#D5F5E3")),
            locations=cells_body(rows=4)) %>%        # traumatic union
  tab_style(style=list(cell_text(weight="bold"),
                       cell_fill(color="#E8DAEF")),
            locations=cells_body(rows=10)) %>%       # non-traumatic
  tab_style(style=list(cell_text(color="#6C3483", style="italic")),
            locations=cells_body(rows=c(3, 8))) %>%  # CNS-miscoded rows
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  sub_missing(missing_text="\u2014") %>%
  cols_label(Category="Category",
             `Unweighted N`="Unweighted N",
             `Weighted N`="Weighted N (national estimate)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              row_group.font.weight="bold",
              row_group.background.color="#F5F8FB",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sensitivity Analysis 1 gt ────────────────────────────────────────────────
gt_sens1 <- sens1_tbl %>%
  setNames(c("Metric","Value")) %>%
  gt() %>%
  tab_header(title=md("**Sensitivity Analysis 1 \u2014 Criterion A Only**")) %>%
  tab_footnote(footnote=md(paste0(
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, excluding ",
    "external-cause-only entries and excluding the Criterion B procedural ",
    "pathway. HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Metric="Metric", Value="Value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sensitivity Analysis 2 gt ────────────────────────────────────────────────
gt_sens2 <- sens2_tbl %>%
  setNames(c("Metric","Value")) %>%
  gt() %>%
  tab_header(title=md("**Sensitivity Analysis 2 \u2014 Highest Mechanistic Specificity**")) %>%
  tab_footnote(footnote=md(paste0(
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
    "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Metric="Metric", Value="Value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sensitivity Analysis 3 gt ────────────────────────────────────────────────
gt_sens3 <- sens3_tbl %>%
  setNames(c("Metric","Value")) %>%
  gt() %>%
  tab_header(title=md("**Sensitivity Analysis 3 \u2014 Maximum Specificity**"),
             subtitle=md("*Laterality-specified nerve + anatomic wound + external cause*")) %>%
  tab_footnote(footnote=md(paste0(
    "Laterality-specified facial nerve injury (S04.51* or S04.52*) co-occurring ",
    "with anatomically concordant wound (S01.1*, S01.3*, S01.4*, S01.5*, S01.81*) ",
    "AND external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "Group 2 = Group 1 + CPT 64864, 64865, 64885, or 64886. ",
    "External cause code completeness varies by institution and year. ",
    "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Metric="Metric", Value="Value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Mechanism of Injury gt ───────────────────────────────────────────────────
gt_mechanism <- mechanism_display %>%
  gt() %>%
  tab_header(
    title=md("**Mechanism of Injury Distribution \u2014 Group 1 Encounters**"),
    subtitle=md("*HCUP NASS 2016\u20132023*")) %>%
  tab_footnote(footnote=md(paste0(
    "Each Group 1 encounter classified by its most specific external cause code ",
    "(V01\u2013Y99) using a clinical hierarchy. ", pct_with_mechanism,
    "% of Group 1 encounters carried at least one external cause code. ",
    "Per HCUP DUA, categories with unweighted N <11 are suppressed. ",
    "Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  sub_missing(missing_text="\u2014") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Save all Word tables ──────────────────────────────────────────────────────
save_gt_to_word(gt_sheet1,    "Table1_Patient_Characteristics.docx")
save_gt_to_word(gt_sheet2,    "Table2_Logistic_Regression.docx")
save_gt_to_word(gt_sheet4,    "Table4_Repair_Rates_by_Year.docx")
save_gt_to_word(gt_sheet5,    "Table5_Charges_Analysis.docx")
save_gt_to_word(gt_nis,       "Table_NIS2023_Inpatient_Results.docx")
save_gt_to_word(gt_sens1,     "Table_Sensitivity_Analysis_1.docx")
save_gt_to_word(gt_sens2,     "Table_Sensitivity_Analysis_2.docx")
save_gt_to_word(gt_sens3,     "Table_Sensitivity_Analysis_3.docx")
save_gt_to_word(gt_mechanism, "Table_Mechanism_of_Injury.docx")

message("\nAll Word tables saved to: ", word_dir)
```

# NIS Exploratory Parallel Cohort (2023)

```{r nis-parallel-cohort}
#| cache: false
# Apply the v1.7 NASS inclusion logic to NIS 2023 inpatient data, adapted
# for ICD-10-PCS procedure codes. Includes BOTH PNS-correctly-coded (01QM*)
# AND CNS-miscoded (00QM*) facial nerve repair codes — the v1.1 confirmation
# analysis showed both represent real facial nerve surgery.
#
# This chunk computes the inpatient parallel to NASS Group 1 / Group 2 /
# Group 3 to enable direct comparison of national volume estimates and
# repair rates between the ambulatory and inpatient settings.

library(readr)

# v1.9: prefer local Desktop alias; fall back to OneDrive if not present
nis_subpath <- file.path("HCUP NIS", "NIS_2023", "NIS_2023_Core.ASC")
nis_local    <- file.path(local_data_root,    nis_subpath)
nis_onedrive <- file.path(onedrive_data_root, nis_subpath)

nis_core_asc <- if (file.exists(nis_local)) {
  message("NIS 2023 ASC located locally: ", nis_local)
  nis_local
} else if (file.exists(nis_onedrive)) {
  message("NIS 2023 ASC located in OneDrive (fallback): ", nis_onedrive)
  nis_onedrive
} else {
  ""
}

nis_available <- nzchar(nis_core_asc) && file.exists(nis_core_asc)
if (!nis_available) {
  message("NIS 2023 ASC file not found at either location.")
  message("Skipping NIS parallel cohort analysis. Update paths if needed.")
  nis_parallel_results <- tibble(
    Metric = c("Status"),
    Value  = c("NIS 2023 file unavailable \u2014 chunk skipped")
  )
} else {
  message("Reading NIS 2023 Core ASC (this may take 1\u20132 minutes)...")
  # Column positions per HCUP StataLoad_NIS_2023_Core.Do (12/10/2025):
  #   HOSP_NIS    1-5, NIS_STRATUM 16-19, DISCWT 31-41
  #   I10_DX1-40  starting at col 67, 7 chars each
  #   I10_PR1-25  starting at col 355, 7 chars each
  dx_starts <- seq(67,  by=7, length.out=40)
  pr_starts <- seq(355, by=7, length.out=25)

  col_starts <- c(1,  16,  31,  dx_starts, pr_starts)
  col_ends   <- c(5,  19,  41,  dx_starts+6, pr_starts+6)
  col_names  <- c("HOSP_NIS","NIS_STRATUM","DISCWT",
                  paste0("I10_DX", 1:40),
                  paste0("I10_PR", 1:25))

  col_spec <- fwf_positions(start = col_starts,
                             end   = col_ends,
                             col_names = col_names)

  nis <- read_fwf(nis_core_asc, col_positions = col_spec,
                   col_types = cols(.default = col_character()),
                   show_col_types = FALSE)
  nis$DISCWT <- as.numeric(nis$DISCWT)
  message("NIS 2023 read: ", format(nrow(nis), big.mark=","), " rows")

  # ── Build flags using research-question-aligned definition ──────────────
  # Group 1 = (Criterion A) OR (Criterion B), where:
  #   Criterion A: S04.5* (7th char A or D) co-occurring with S01.*/S11.*
  #   Criterion B: facial nerve repair PCS code co-occurring with S01.*/S11.*
  # Group 2 = Group 1 AND any repair PCS code present
  # Group 3 = Group 1 AND NOT Group 2
  nis_dx_cols <- paste0("I10_DX", 1:40)
  nis_pr_cols <- paste0("I10_PR", 1:25)

  # Facial nerve repair PCS codes — repair, supplement (graft), and reposition.
  # The 00QM codes (CNS section) are miscoded but represent real facial nerve
  # repairs per the v1.1 confirmation analysis (75.9% diagnosis-profile match,
  # exceeding the 70% threshold).
  # The 01UM codes (Supplement / nerve graft) are included for forward
  # compatibility with future NIS years; in 2023 they return zero hits.
  facial_nerve_repair_pcs <- c(
    "01QM0ZZ","01QM3ZZ","01QM4ZZ",   # PNS Repair (anatomically correct)
    "00QM0ZZ","00QM3ZZ","00QM4ZZ",   # CNS Repair (miscoded but confirmed real)
    "01UM07Z","01UM0KZ","01UM0JZ"    # PNS Supplement (autologous / nonauto / synthetic graft)
  )

  # Helper: vectorized check across a set of columns
  any_match_start <- function(df, cols, pattern) {
    rowSums(sapply(cols, function(c) {
      cd <- trimws(as.character(df[[c]]))
      !is.na(cd) & nzchar(cd) & str_starts(cd, pattern) & !str_ends(cd, "S")
    }), na.rm=TRUE) > 0
  }
  any_match_exact <- function(df, cols, codes) {
    rowSums(sapply(cols, function(c) {
      cd <- trimws(as.character(df[[c]]))
      !is.na(cd) & cd %in% codes
    }), na.rm=TRUE) > 0
  }

  message("Computing NIS flags...")

  nis_dx_s045 <- any_match_start(nis, nis_dx_cols, "S045")
  nis_s01     <- any_match_start(nis, nis_dx_cols, "S01")
  nis_s11     <- any_match_start(nis, nis_dx_cols, "S11")
  nis_pcs     <- any_match_exact(nis, nis_pr_cols, facial_nerve_repair_pcs)

  # Neoplasm/iatrogenic exclusion (v1.8)
  nis_excl <- rowSums(sapply(nis_dx_cols, function(c) {
    cd <- trimws(as.character(nis[[c]]))
    str_detect(cd, "^(C07|D110|C490|D333|C724|T854|Y83)") & nzchar(cd)
  })) > 0

  # External cause flag (informational, not used for inclusion)
  nis_ext <- rowSums(sapply(nis_dx_cols, function(c) {
    cd <- trimws(as.character(nis[[c]]))
    str_detect(cd, "^(W2[5-7]|W45|X99|Y28|V[0-9]{2})")
  })) > 0

  # Phase 2 v1.7 inclusion logic adapted for ICD-10-PCS:
  # Criterion A: S04.5 + (S01 OR S11)
  # Criterion B: any facial nerve repair PCS + (S01 OR S11)
  nis_crit_a <- nis_dx_s045 & (nis_s01 | nis_s11)
  nis_crit_b <- nis_pcs     & (nis_s01 | nis_s11)
  nis_g1     <- (nis_crit_a | nis_crit_b) & !nis_excl
  nis_g2     <- nis_pcs & nis_g1
  nis_g3     <- nis_g1 & !nis_g2

  # ── Add flags as columns so svydesign can access them ────────────────────
  nis$flag_g1 <- as.integer(nis_g1)
  nis$flag_g2 <- as.integer(nis_g2)
  nis$flag_g3 <- as.integer(nis_g3)

  # ── Compute weighted estimates ───────────────────────────────────────────
  # Direct sum-of-weights for point estimates (mathematically identical to
  # svytotal for a logical flag; survey design retained for any future
  # variance estimation).
  options(survey.lonely.psu = "adjust")
  nis_design <- svydesign(
    ids     = ~HOSP_NIS,
    strata  = ~NIS_STRATUM,
    weights = ~DISCWT,
    data    = nis,
    nest    = TRUE
  )

  nis_total_unw  <- sum(nis_g1, na.rm=TRUE)
  nis_g2_unw     <- sum(nis_g2, na.rm=TRUE)
  nis_g3_unw     <- sum(nis_g3, na.rm=TRUE)

  nis_total_w    <- round(sum(nis$DISCWT[nis_g1], na.rm=TRUE))
  nis_g2_w       <- round(sum(nis$DISCWT[nis_g2], na.rm=TRUE))
  nis_g3_w       <- round(sum(nis$DISCWT[nis_g3], na.rm=TRUE))
  nis_rate_w     <- if (nis_total_w > 0)
                      round(100 * nis_g2_w / nis_total_w, 1) else NA_real_

  # Criterion contributions
  nis_a_only <- sum( nis_crit_a & !nis_crit_b & !nis_excl)
  nis_b_only <- sum(!nis_crit_a &  nis_crit_b & !nis_excl)
  nis_both   <- sum( nis_crit_a &  nis_crit_b & !nis_excl)
  nis_dropped_excl <- sum((nis_crit_a | nis_crit_b) & nis_excl)

  # ── Build nis_g1_data with dx_raw for downstream mechanism analysis ─────
  # Only Group 1 encounters; carries all 40 dx columns concatenated for
  # mechanism-of-injury classification in the next chunk.
  if (nis_total_unw > 0) {
    clean_codes <- function(mat) {
      apply(mat, 1, function(row) {
        codes <- trimws(as.character(row))
        codes <- codes[!is.na(codes) & nzchar(codes) & codes != "."]
        paste(codes, collapse="|")
      })
    }
    nis_g1_data <- tibble(
      had_repair = as.integer(nis_g2[nis_g1]),
      discwt     = nis$DISCWT[nis_g1],
      dx_raw     = clean_codes(nis[nis_g1, nis_dx_cols, drop=FALSE])
    )
  } else {
    nis_g1_data <- tibble(had_repair=integer(0), discwt=numeric(0), dx_raw=character(0))
  }

  message("\n=== NIS 2023 Parallel Cohort Results ===")
  message("Group 1: ", nis_total_unw, " unweighted / ",
          format(nis_total_w, big.mark=","), " weighted")
  message("Group 2: ", nis_g2_unw,    " unweighted / ",
          format(nis_g2_w,    big.mark=","), " weighted")
  message("Group 3: ", nis_g3_unw,    " unweighted / ",
          format(nis_g3_w,    big.mark=","), " weighted")
  message("Repair rate: ", nis_rate_w, "%")
  message("Criterion A only / B only / both: ",
          nis_a_only, " / ", nis_b_only, " / ", nis_both)
  message("Dropped by neoplasm/iatrogenic exclusion: ", nis_dropped_excl)

  # ── Sample-size guards for HCUP DUA cell suppression ─────────────────────
  fmt_or_suppress <- function(unw, w) {
    if (is.na(unw) || unw < 11) return("\u2014 (suppressed, N<11)")
    paste0(format(w, big.mark=","), " (N=", unw, ")")
  }

  # NASS comparison numbers (using full 2016\u20132023 NASS cohort + 2023 only)
  nass_2023 <- enc_all %>% filter(year == "2023")
  nass_2023_g1_unw <- nrow(nass_2023)
  nass_2023_g1_w   <- round(sum(nass_2023$discwt, na.rm=TRUE))
  nass_2023_g2_unw <- sum(nass_2023$had_repair == 1)
  nass_2023_g2_w   <- round(sum(nass_2023$discwt[nass_2023$had_repair==1],
                                 na.rm=TRUE))
  nass_2023_rate   <- if (nass_2023_g1_w > 0)
    round(100 * nass_2023_g2_w / nass_2023_g1_w, 1) else NA_real_

  nis_parallel_results <- tibble(
    Metric = c(
      "Group 1 unweighted N",
      "Group 1 weighted N (national estimate)",
      "Group 2 unweighted N",
      "Group 2 weighted N",
      "Group 3 unweighted N",
      "Group 3 weighted N",
      "Repair rate (%, weighted)",
      "Criterion A only (unweighted)",
      "Criterion B only (unweighted)",
      "Both criteria (unweighted)",
      "Excluded by neoplasm/iatrogenic codes"
    ),
    `NASS 2023 (ambulatory)` = c(
      format(nass_2023_g1_unw, big.mark=","),
      format(nass_2023_g1_w,   big.mark=","),
      format(nass_2023_g2_unw, big.mark=","),
      format(nass_2023_g2_w,   big.mark=","),
      format(nass_2023_g1_unw - nass_2023_g2_unw, big.mark=","),
      format(nass_2023_g1_w   - nass_2023_g2_w,   big.mark=","),
      paste0(nass_2023_rate, "%"),
      "(see main analysis)",
      "(see main analysis)",
      "(see main analysis)",
      "(see main analysis)"
    ),
    `NIS 2023 (inpatient)` = c(
      format(nis_total_unw, big.mark=","),
      format(nis_total_w,   big.mark=","),
      format(nis_g2_unw,    big.mark=","),
      format(nis_g2_w,      big.mark=","),
      format(nis_g3_unw,    big.mark=","),
      format(nis_g3_w,      big.mark=","),
      paste0(nis_rate_w, "%"),
      format(nis_a_only, big.mark=","),
      format(nis_b_only, big.mark=","),
      format(nis_both,   big.mark=","),
      format(nis_dropped_excl, big.mark=",")
    )
  )
}

nis_parallel_results %>%
  kable(caption = paste0(
    "NIS 2023 vs NASS 2023 \u2014 parallel cohort comparison applying ",
    "the v1.8 inclusion logic to both databases. NIS extraction includes ",
    "both PNS-correctly-coded (01QM) and CNS-miscoded (00QM) facial nerve ",
    "repair PCS codes per the v1.1 NIS confirmation analysis. ",
    "External cause codes are not used for inclusion in either database. ",
    "Per HCUP DUA, unweighted N <11 are flagged as suppressed for any ",
    "external publication."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(7, bold = TRUE, background = "#EBF4FB")
```

# NIS Mechanism of Injury (2023)

```{r nis-mechanism}
#| cache: false
# Apply the same hierarchical mechanism-of-injury classification used in the
# NASS chunk to the NIS Group 1 cohort. Allows direct ambulatory-vs-inpatient
# comparison of the mechanism distribution.
#
# Hierarchy (most specific first): assault by sharp object > other assault >
# self-harm > transport > cutting/piercing > fall > animal > undetermined >
# other > none.

nis_mechanism_categories <- c(
  "Assault by sharp object (X99)",
  "Assault other (X9*, Y0*)",
  "Self-harm (X71\u2013X83)",
  "Transport-related (V01\u2013V99)",
  "Cutting/piercing accident (W25\u2013W27, W45)",
  "Fall (W00\u2013W19)",
  "Animal-related (W50\u2013W64)",
  "Undetermined intent (Y10\u2013Y34)",
  "Other external cause",
  "No external cause coded"
)

classify_mechanism_nis <- function(codes) {
  if (length(codes) == 0) return("No external cause coded")
  if (any(str_starts(codes, "X99")))
    return("Assault by sharp object (X99)")
  if (any(str_detect(codes, "^(X9[0-8]|Y0[0-9])")))
    return("Assault other (X9*, Y0*)")
  if (any(str_detect(codes, "^(X7[1-9]|X8[0-3])")))
    return("Self-harm (X71\u2013X83)")
  if (any(str_detect(codes, "^V[0-9]{2}")))
    return("Transport-related (V01\u2013V99)")
  if (any(str_detect(codes, "^(W2[5-7]|W45)")))
    return("Cutting/piercing accident (W25\u2013W27, W45)")
  if (any(str_detect(codes, "^W[01][0-9]")))
    return("Fall (W00\u2013W19)")
  if (any(str_detect(codes, "^W[56][0-9]")))
    return("Animal-related (W50\u2013W64)")
  if (any(str_detect(codes, "^Y[12][0-9]|^Y3[0-4]")))
    return("Undetermined intent (Y10\u2013Y34)")
  if (any(str_detect(codes, "^[WXY][0-9]")))
    return("Other external cause")
  return("No external cause coded")
}

if (!exists("nis_g1_data") || nrow(nis_g1_data) == 0) {
  message("NIS Group 1 cohort empty or unavailable; skipping NIS mechanism analysis.")
  nis_mechanism_display <- tibble(
    `Mechanism of injury` = "Status",
    `Unweighted N` = NA_integer_,
    `Weighted N`   = NA_integer_
  )
  nis_pct_with_mechanism <- NA_real_
} else {
  nis_mech_assignments <- nis_g1_data %>%
    mutate(
      code_list = strsplit(dx_raw, "\\|"),
      mechanism = vapply(code_list, classify_mechanism_nis, character(1))
    ) %>%
    select(had_repair, discwt, mechanism)

  nis_mechanism_summary <- nis_mech_assignments %>%
    group_by(mechanism) %>%
    summarise(
      `Unweighted N`            = n(),
      `Weighted N`              = round(sum(discwt, na.rm=TRUE)),
      `Repair (unweighted N)`   = sum(had_repair == 1),
      `Repair (weighted N)`     = round(sum(discwt[had_repair == 1], na.rm=TRUE)),
      .groups = "drop"
    ) %>%
    mutate(
      `% of Group 1 (unweighted)` = round(100 * `Unweighted N` /
                                            nrow(nis_g1_data), 1),
      `% of Group 1 (weighted)`   = round(100 * `Weighted N` /
                                            max(round(sum(nis_g1_data$discwt,
                                                          na.rm=TRUE)), 1), 1),
      `Repair rate (%)`           = ifelse(`Weighted N` > 0,
                                            round(100 * `Repair (weighted N)` /
                                                  `Weighted N`, 1),
                                            NA_real_)
    ) %>%
    mutate(mechanism = factor(mechanism, levels = nis_mechanism_categories)) %>%
    arrange(mechanism) %>%
    mutate(mechanism = as.character(mechanism))

  # HCUP DUA cell suppression for any unweighted < 11
  nis_mechanism_display <- nis_mechanism_summary %>%
    mutate(across(c(`Unweighted N`, `Weighted N`,
                     `Repair (unweighted N)`, `Repair (weighted N)`,
                     `% of Group 1 (unweighted)`, `% of Group 1 (weighted)`,
                     `Repair rate (%)`),
                  ~ ifelse(`Unweighted N` < 11, NA, .))) %>%
    rename(`Mechanism of injury` = mechanism)

  n_with_mech_nis <- sum(nis_mech_assignments$mechanism != "No external cause coded")
  nis_pct_with_mechanism <- round(100 * n_with_mech_nis /
                                    nrow(nis_g1_data), 1)
  message("NIS Group 1 encounters with any external cause coded: ",
          n_with_mech_nis, " (", nis_pct_with_mechanism, "%)")
}

nis_mechanism_display %>%
  kable(caption = paste0(
    "NIS 2023 Mechanism of injury distribution in Group 1 encounters, classified ",
    "using ICD-10-CM external cause codes carried on the encounter. Each ",
    "encounter is assigned to its most specific mechanism category using the same ",
    "clinical hierarchy applied to the NASS cohort. ",
    if (!is.na(nis_pct_with_mechanism))
      paste0(nis_pct_with_mechanism, "% of NIS Group 1 encounters carry at least ",
             "one external cause code. ") else "",
    "Per HCUP DUA, categories with unweighted N < 11 are suppressed (NA). ",
    "HCUP NIS 2023."),
    format.args = list(big.mark=",")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE)
```

# Draft Methods Paragraph

```{r draft-methods}
#| cache: false
g1_unw_fmt   <- format(nrow(enc_all),                  big.mark=",")
g1_w_fmt     <- format(round(total_g1_w),              big.mark=",")
g2_w_fmt     <- format(round(total_repair_w),          big.mark=",")
g3_w_fmt     <- format(round(total_norepair_w),        big.mark=",")
rate_pct     <- round(100*total_repair_w/total_g1_w,   1)
norate_pct   <- round(100*total_norepair_w/total_g1_w, 1)

draft_methods <- paste0(
  "We analyzed the HCUP National Ambulatory Surgery Sample (NASS) for 2016 ",
  "through 2023. A total of ", g1_w_fmt, " encounters with traumatic facial ",
  "nerve laceration were identified nationally (unweighted N = ", g1_unw_fmt,
  "), of whom ", g2_w_fmt, " (", rate_pct, "%) underwent surgical repair ",
  "(CPT 64864, 64865, 64885, or 64886). ", g3_w_fmt, " patients (", norate_pct,
  "%) were managed without operative repair."
)

cat(draft_methods)
```

# Copy Outputs to OneDrive

```{r copy-to-drive}
#| cache: false
# Copy the rendered HTML and the .qmd source into the Quarto Versions subfolder.
qmd_basename  <- "nass_phase2_analysis_2_0.qmd"
html_basename <- "nass_phase2_analysis_2_0.html"

# HTML: search common locations for the rendered output
html_candidates <- c(
  file.path(getwd(),     html_basename),
  file.path("~/Downloads", html_basename),
  file.path("~/Documents", html_basename)
)
html_src <- html_candidates[file.exists(path.expand(html_candidates))][1]
if (!is.na(html_src)) {
  file.copy(path.expand(html_src),
            file.path(qmd_dir, html_basename),
            overwrite=TRUE)
  message("HTML copied from: ", html_src)
} else {
  message("HTML not yet available; render the document and re-run this chunk.")
}

# QMD source: only copy if source and destination are different paths
qmd_src <- file.path(getwd(), qmd_basename)
qmd_dst <- file.path(qmd_dir, qmd_basename)
if (file.exists(qmd_src) &&
    normalizePath(qmd_src) != normalizePath(qmd_dst)) {
  file.copy(qmd_src, qmd_dst, overwrite=TRUE)
  message("QMD copied from: ", qmd_src)
} else {
  message("QMD already in Quarto Versions — skipping self-copy.")
}

message("\nAll outputs saved under: ", output_dir)
message("  Figures      : ", fig_dir)
message("  Excel        : ", xlsx_dir)
message("  Word Tables  : ", word_dir)
message("  Quarto       : ", qmd_dir)
```

# Session Info

```{r session}
sessionInfo()
```
