---
title: "Traumatic Facial Nerve Laceration: HCUP NASS 2016–2023 (Phase 2)"
author: "Solomon Johnson"
date: "`r Sys.Date()`"
output-file: "nass_phase2_analysis_1_0.html"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
  freeze: auto
---

<!--
============================================================
CHANGELOG
v1.0 — 2026-04-27 — Initial Phase 2 analysis. New Group 1
       definition (Criterion A: S04.5 + open wound or external
       cause; Criterion B: CPT + open wound). Adds two
       sensitivity analyses, Phase 1 vs Phase 2 comparison
       sheet, and an automated draft methods paragraph.
       Output redirected from Desktop to OneDrive Phase 2
       folder with subfolder routing (Figures, Excel, Word
       Tables, Quarto Versions).
============================================================
-->

```{r setup}
#| cache: false
pkgs <- c("haven","dplyr","purrr","furrr","future","readr","tidyr",
          "stringr","survey","openxlsx","ggplot2","scales","patchwork",
          "ggrepel","DiagrammeR","DiagrammeRsvg","rsvg",
          "WeightIt","cobalt","knitr","kableExtra","maps","mapproj",
          "gt","officer","flextable")
for (p in pkgs) if (!requireNamespace(p, quietly=TRUE)) install.packages(p)

library(haven);     library(dplyr);     library(purrr);    library(furrr)
library(future);    library(readr);     library(tidyr);    library(stringr)
library(survey);    library(openxlsx);  library(ggplot2);  library(scales)
library(patchwork); library(ggrepel);   library(WeightIt); library(cobalt)
library(knitr);     library(kableExtra);library(maps);     library(mapproj)
library(DiagrammeR);library(DiagrammeRsvg);library(rsvg)
library(gt);

# Phase 2 output directory — OneDrive only, no Desktop output
output_dir <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Analysis/Phase 2"

if (!dir.exists(path.expand(output_dir))) {
  stop("Phase 2 output directory not found: ", output_dir)
}

fig_dir   <- file.path(path.expand(output_dir), "Figures")
xlsx_dir  <- file.path(path.expand(output_dir), "Excel")
word_dir  <- file.path(path.expand(output_dir), "Word Tables")
qmd_dir   <- file.path(path.expand(output_dir), "Quarto Versions")

for (d in c(fig_dir, xlsx_dir, word_dir, qmd_dir)) {
  if (!dir.exists(d)) {
    stop("Required Phase 2 subfolder not found: ", d)
  }
}

message("Phase 2 output_dir confirmed: ", output_dir)
```

# Configuration

```{r config}
d16_23 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/NASS_HCUP Database/RAW Data Files HNASS"
d21 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2021"
d22 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2022"
d23 <- "/Users/solomon/Library/CloudStorage/OneDrive-UW/Research projects/NASS traumatic facial nerve repair project/Data/HCUP Original Data Files 21 through 23/HCUP NASS 21 through 23/NASS_2023"

enc_files <- list(
  "2016" = file.path(d16_23, "NASS_2016_Encounter-002.dta"),
  "2017" = file.path(d16_23, "NASS_2017_Encounter-003.dta"),
  "2018" = file.path(d16_23, "NASS_2018_Encounter-001.dta"),
  "2019" = file.path(d16_23, "NASS_2019_Encounter-005.dta"),
  "2020" = file.path(d16_23, "NASS_2020_Encounter.dta"),
  "2021" = file.path(d21,    "NASS_2021_Encounter.dta"),
  "2022" = file.path(d22,    "NASS_2022_Encounter.dta"),
  "2023" = file.path(d23,    "NASS_2023_Encounter.dta")
)

hosp_files <- list(
  "2016" = file.path(d16_23, "NASS_2016_Hospital.dta"),
  "2017" = file.path(d16_23, "NASS_2017_Hospital.dta"),
  "2018" = file.path(d16_23, "NASS_2018_Hospital.dta"),
  "2019" = file.path(d16_23, "NASS_2019_Hospital.dta"),
  "2020" = file.path(d16_23, "NASS_2020_Hospital.dta"),
  "2021" = file.path(d21,    "NASS_2021_Hospital.dta"),
  "2022" = file.path(d22,    "NASS_2022_Hospital.dta"),
  "2023" = file.path(d23,    "NASS_2023_Hospital.dta")
)

cpt_codes  <- c("64864","64865")
dx_stem    <- "S045"

# Phase 2 inclusion logic — broader open wound matching
s01_stem   <- "S01"   # open wound of head, any subcode
s11_stem   <- "S11"   # open wound of neck, any subcode

# External cause codes: W25-W27, W45, X99, Y28, V01-V99
# Matched against the leading 3 characters of any dx code.
ext_cause_pattern <- "^(W2[5-7]|W45|X99|Y28|V(0[1-9]|[1-9][0-9]))"

age_breaks <- c(18, 35, 45, 55, 65, Inf)
age_labels <- c("18-34","35-44","45-54","55-65","65+")

assign_age_group <- function(x) {
  cut(as.numeric(x), breaks=age_breaks, labels=age_labels,
      right=FALSE, include.lowest=TRUE)
}

# HCUP NASS variable coding (verified per official NASS Notes documentation)
# HOSP_LOCATION:  0=Rural, 1=Urban
# HOSP_TEACH:     0=Nonteaching, 1=Teaching
# HOSP_LOCTEACH:  1=Rural nonteaching, 2=Urban nonteaching, 3=Urban teaching
# HOSP_REGION:    1=Northeast, 2=Midwest, 3=South, 4=West
# FEMALE:         0=Male, 1=Female
# PAY1:           1=Medicare,2=Medicaid,3=Private,4=Self-pay,5=No charge,6=Other
# RACE:           1=White,2=Black,3=Hispanic,4=Asian/PI,5=Native American,6=Other
# ZIPINC_QRTL:    1=0-25th,2=26-50th,3=51-75th,4=76-100th
```

# Data Extraction

```{r extract-function}
extract_year <- function(year, enc_fp, hosp_fp) {
  message("  Loading: ", year)
  if (!file.exists(enc_fp))  { warning("Missing: ", enc_fp);  return(NULL) }
  if (!file.exists(hosp_fp)) { warning("Missing: ", hosp_fp); return(NULL) }

  enc  <- read_dta(enc_fp)
  hosp <- read_dta(hosp_fp)
  ec   <- names(enc); hc <- names(hosp)

  get_col <- function(pat, cols) grep(pat, cols, value=TRUE, ignore.case=TRUE)[1]

  dx_cols  <- grep("^i10_dx",   ec, value=TRUE, ignore.case=TRUE)
  cpt_cols <- grep("^cpt[0-9]", ec, value=TRUE, ignore.case=TRUE)

  age_col    <- get_col("^age$",          ec)
  female_col <- get_col("^female$",       ec)
  race_col   <- get_col("^race$",         ec)
  pay1_col   <- get_col("^pay1$",         ec)
  zip_col    <- get_col("^zipinc_qrtl$",  ec)
  totchg_col <- get_col("^totchg",        ec)
  discwt_col <- get_col("^discwt$",       ec)
  hosp_id_e  <- get_col("^hosp_nass$",    ec)
  hosp_id_h  <- get_col("^hosp_nass$",    hc)
  region_col <- get_col("^hosp_region$",  hc)
  loc_col    <- get_col("^hosp_location$",hc)
  teach_col  <- get_col("^hosp_teach$|^hosp_locteach$", hc)

  # Phase 2 flag construction
  dx_s045 <- rowSums(sapply(dx_cols, function(col)
    str_starts(trimws(as.character(enc[[col]])), dx_stem))) > 0
  s01_flag <- rowSums(sapply(dx_cols, function(col)
    str_starts(trimws(as.character(enc[[col]])), s01_stem))) > 0
  s11_flag <- rowSums(sapply(dx_cols, function(col)
    str_starts(trimws(as.character(enc[[col]])), s11_stem))) > 0
  ext_cause_flag <- rowSums(sapply(dx_cols, function(col)
    str_detect(trimws(as.character(enc[[col]])), ext_cause_pattern))) > 0
  cpt_flag <- rowSums(sapply(cpt_cols, function(col)
    trimws(as.character(enc[[col]])) %in% cpt_codes)) > 0

  # Phase 2 Group 1 = Criterion A OR Criterion B
  #   Criterion A: S04.5 co-occurring with S01.* OR S11.* OR external cause
  #   Criterion B: CPT 64864/64865 co-occurring with S01.* OR S11.*
  crit_a <- dx_s045 & (s01_flag | s11_flag | ext_cause_flag)
  crit_b <- cpt_flag & (s01_flag | s11_flag)
  g1     <- crit_a | crit_b
  g2     <- cpt_flag & g1
  keep   <- g1
  if (sum(keep) == 0) return(NULL)

  cpt_assigned <- rep(NA_character_, nrow(enc))
  for (code in cpt_codes) {
    hit <- rowSums(sapply(cpt_cols, function(col)
      trimws(as.character(enc[[col]])) == code)) > 0
    cpt_assigned[hit & is.na(cpt_assigned)] <- code
  }

  out <- tibble(
    year       = year,
    had_repair = as.integer(g2[keep]),
    cpt_code   = cpt_assigned[keep],
    discwt     = if (!is.na(discwt_col))
                   as.numeric(enc[[discwt_col]][keep]) else rep(1, sum(keep)),
    hosp_id    = if (!is.na(hosp_id_e))
                   as.character(enc[[hosp_id_e]][keep])
                 else as.character(seq_len(sum(keep))),
    totchg     = if (!is.na(totchg_col))
                   as.numeric(enc[[totchg_col]][keep]) else NA_real_,
    # Phase 2 — preserve per-row flags so sensitivity analyses can subset
    # enc_all directly without a re-extraction pass
    flag_s045      = dx_s045[keep],
    flag_s01       = s01_flag[keep],
    flag_s11       = s11_flag[keep],
    flag_extcause  = ext_cause_flag[keep],
    flag_cpt       = cpt_flag[keep],
    flag_crit_a    = crit_a[keep],
    flag_crit_b    = crit_b[keep]
  )

  out$age_group <- if (!is.na(age_col))
    as.character(assign_age_group(enc[[age_col]][keep]))
  else NA_character_

  if (!is.na(female_col)) {
    f <- as.numeric(enc[[female_col]][keep])
    out$sex <- case_when(f==1~"Female", f==0~"Male", TRUE~NA_character_)
  } else out$sex <- NA_character_

  if (!is.na(race_col)) {
    r <- as.character(as.numeric(enc[[race_col]][keep]))
    out$race <- case_when(
      r=="1" ~ "White", r=="2" ~ "Black", r=="3" ~ "Hispanic",
      r %in% c("4","5","6","7") ~
        "Asian or Pacific Islander or Native American or Other",
      TRUE ~ NA_character_)
  } else out$race <- NA_character_

  if (!is.na(pay1_col)) {
    p <- as.character(as.numeric(enc[[pay1_col]][keep]))
    out$insurance <- case_when(
      p=="1"~"Medicare", p=="2"~"Medicaid", p=="3"~"Private insurance",
      p=="4"~"Self-pay", p %in% c("5","6")~"Other", TRUE~NA_character_)
  } else out$insurance <- NA_character_

  if (!is.na(zip_col)) {
    z <- as.character(as.numeric(enc[[zip_col]][keep]))
    out$income_qrtl <- case_when(
      z=="1"~"0-25th", z=="2"~"26th to 50th",
      z=="3"~"51st to 75th", z=="4"~"76th to 100th", TRUE~NA_character_)
  } else out$income_qrtl <- NA_character_

  if (!is.na(hosp_id_h)) {
    hosp_slim <- hosp %>%
      mutate(hosp_id = as.character(.data[[hosp_id_h]]))

    hosp_slim$hosp_region <- if (!is.na(region_col))
      case_when(as.numeric(hosp[[region_col]])==1~"Northeast",
                as.numeric(hosp[[region_col]])==2~"Midwest",
                as.numeric(hosp[[region_col]])==3~"South",
                as.numeric(hosp[[region_col]])==4~"West",
                TRUE~NA_character_)
    else NA_character_

    hosp_slim$hosp_location <- if (!is.na(loc_col))
      case_when(as.numeric(hosp[[loc_col]])==0~"Rural",
                as.numeric(hosp[[loc_col]])==1~"Urban",
                TRUE~NA_character_)
    else NA_character_

    if (!is.na(teach_col)) {
      t_raw <- as.numeric(hosp[[teach_col]])
      hosp_slim$hosp_teach <- if (grepl("locteach", tolower(teach_col)))
        case_when(t_raw==3~"Teaching", t_raw %in% c(1,2)~"Nonteaching",
                  TRUE~NA_character_)
      else
        case_when(t_raw==1~"Teaching", t_raw==0~"Nonteaching",
                  TRUE~NA_character_)
    } else hosp_slim$hosp_teach <- NA_character_

    out <- left_join(out,
                     hosp_slim %>% select(hosp_id, hosp_region,
                                          hosp_location, hosp_teach),
                     by="hosp_id")
  } else {
    out$hosp_region   <- NA_character_
    out$hosp_location <- NA_character_
    out$hosp_teach    <- NA_character_
  }
  out
}
```

```{r run-extraction}
#| cache: true
message("Starting parallel extraction 2016-2023 (Phase 2 criteria)...")
plan(multisession, workers = max(1, availableCores()-1))
enc_all <- bind_rows(Filter(Negate(is.null),
  future_map2(names(enc_files), enc_files,
    ~ extract_year(.x, .y, hosp_files[[.x]]),
    .options = furrr_options(seed=TRUE))))
plan(sequential)

enc_all <- enc_all %>%
  mutate(discwt = as.numeric(discwt), had_repair = as.integer(had_repair))

message("Group 1 (unweighted): ", format(nrow(enc_all), big.mark=","))
message("Group 2 repair:       ", format(sum(enc_all$had_repair), big.mark=","))
```

# Survey Design

```{r survey-design}
svy <- svydesign(ids=~hosp_id, weights=~discwt, data=enc_all, nest=TRUE)

total_repair_w   <- sum(enc_all$discwt[enc_all$had_repair==1], na.rm=TRUE)
total_norepair_w <- sum(enc_all$discwt[enc_all$had_repair==0], na.rm=TRUE)
total_g1_w       <- sum(enc_all$discwt, na.rm=TRUE)

message("Weighted Group 1: ", format(round(total_g1_w),       big.mark=","))
message("Weighted Group 2: ", format(round(total_repair_w),   big.mark=","))
message("Weighted Group 3: ", format(round(total_norepair_w), big.mark=","))
```

# Helper Functions

```{r helpers}
fmt_p <- function(p) {
  sapply(p, function(x) {
    if (is.na(x)) return(NA_character_)
    if (x < 0.001) return("<0.001")
    sprintf("%.3f", round(x, 3))
  })
}

wt_cell <- function(var, level, repair_val) {
  idx   <- enc_all$had_repair==repair_val &
           !is.na(enc_all[[var]]) & enc_all[[var]]==level
  n_w   <- sum(enc_all$discwt[idx], na.rm=TRUE)
  denom <- if (repair_val==1) total_repair_w else total_norepair_w
  paste0(format(round(n_w), big.mark=","),
         " (", round(100*n_w/denom, 1), "%)")
}

wt_cell_g1 <- function(var, level) {
  idx <- !is.na(enc_all[[var]]) & enc_all[[var]]==level
  n_w <- sum(enc_all$discwt[idx], na.rm=TRUE)
  paste0(format(round(n_w), big.mark=","),
         " (", round(100*n_w/total_g1_w, 1), "%)")
}

repair_rate <- function(var, level) {
  idx_g1     <- !is.na(enc_all[[var]]) & enc_all[[var]]==level
  idx_repair <- idx_g1 & enc_all$had_repair==1
  n_g1  <- sum(enc_all$discwt[idx_g1],    na.rm=TRUE)
  n_rep <- sum(enc_all$discwt[idx_repair], na.rm=TRUE)
  if (n_g1 == 0) return(NA_character_)
  paste0(round(100*n_rep/n_g1, 1), "%")
}

svy_chisq_p <- function(var) {
  d <- enc_all %>%
    filter(!is.na(.data[[var]]),
           !.data[[var]] %in% c("Unknown","Missing/Unknown"))
  if (length(unique(d[[var]])) < 2) return(NA_real_)
  s <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  r <- tryCatch(
    svychisq(as.formula(paste0("~had_repair+",var)), design=s, statistic="Chisq"),
    error=function(e) NULL)
  if (is.null(r)) return(NA_real_)
  r$p.value
}

# Survey-weighted ANOVA on log-charges within a repair group
# Uses grp column to avoid .data[[]] formula issues in svyglm
svy_anova_charges_p <- function(repair_val, group_var) {
  d <- enc_all %>%
    filter(had_repair==repair_val, !is.na(totchg), totchg>0,
           !is.na(.data[[group_var]]),
           !.data[[group_var]] %in% c("Unknown","Missing/Unknown")) %>%
    mutate(grp = factor(.data[[group_var]]))
  if (length(unique(d$grp)) < 2 || nrow(d) < 10) return(NA_real_)
  s   <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  fit <- tryCatch(svyglm(log(totchg) ~ grp, design=s, family=gaussian()),
                  error=function(e) NULL)
  if (is.null(fit)) return(NA_real_)
  an  <- tryCatch(regTermTest(fit, "grp"), error=function(e) NULL)
  if (is.null(an)) return(NA_real_)
  an$p
}

# Survey-weighted t-test on log-charges between two levels
svy_ttest_charges_p <- function(repair_val, group_var) {
  d <- enc_all %>%
    filter(had_repair==repair_val, !is.na(totchg), totchg>0,
           !is.na(.data[[group_var]]),
           !.data[[group_var]] %in% c("Unknown","Missing/Unknown")) %>%
    mutate(grp = factor(.data[[group_var]]))
  if (length(unique(d$grp)) != 2 || nrow(d) < 10) return(NA_real_)
  s   <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  fit <- tryCatch(svyglm(log(totchg) ~ grp, design=s, family=gaussian()),
                  error=function(e) NULL)
  if (is.null(fit)) return(NA_real_)
  coefs <- summary(fit)$coefficients
  if (nrow(coefs) < 2) return(NA_real_)
  coefs[2, "Pr(>|t|)"]
}

wt_median_iqr <- function(data, repair_val) {
  d <- data %>% filter(had_repair==repair_val)
  if (nrow(d)==0) return(NA_character_)
  s <- svydesign(ids=~hosp_id, weights=~discwt, data=d, nest=TRUE)
  q <- tryCatch(
    svyquantile(~totchg, s, quantiles=c(0.25,0.5,0.75), na.rm=TRUE),
    error=function(e) NULL)
  if (is.null(q)) return(NA_character_)
  v <- as.numeric(q$totchg)
  paste0("$",format(round(v[2]),big.mark=","),
         " ($",format(round(v[1]),big.mark=","),
         "\u2013$",format(round(v[3]),big.mark=","),")")
}

# Weighted mean ± SD using age group midpoints
mean_sd_cell <- function(repair_val) {
  d   <- enc_all[which(enc_all$had_repair==as.integer(repair_val) &
                       !is.na(enc_all$age_group)), ]
  mid <- c("18-34"=26, "35-44"=39.5, "45-54"=49.5, "55-65"=60, "65+"=72)
  ages <- mid[d$age_group]
  wts  <- d$discwt
  m    <- weighted.mean(ages, wts, na.rm=TRUE)
  v    <- sum(wts*(ages-m)^2, na.rm=TRUE)/sum(wts, na.rm=TRUE)
  paste0(round(m,1)," \u00b1 ",round(sqrt(v),1))
}

mean_sd_cell_g1 <- function() {
  d   <- enc_all[which(!is.na(enc_all$age_group)), ]
  mid <- c("18-34"=26, "35-44"=39.5, "45-54"=49.5, "55-65"=60, "65+"=72)
  ages <- mid[d$age_group]
  wts  <- d$discwt
  m    <- weighted.mean(ages, wts, na.rm=TRUE)
  v    <- sum(wts*(ages-m)^2, na.rm=TRUE)/sum(wts, na.rm=TRUE)
  paste0(round(m,1)," \u00b1 ",round(sqrt(v),1))
}

make_rows <- function(var, levels_ordered, label) {
  hdr <- tibble(
    Characteristic    = label,
    `No Repair (n, %)` = NA_character_,
    `Repair (n, %)`    = NA_character_,
    `Total (n)`        = NA_character_,
    `Repair Rate (%)`  = NA_character_
  )
  sub <- map_dfr(levels_ordered, function(lev) {
    tibble(
      Characteristic     = paste0("   ", lev),
      `No Repair (n, %)` = wt_cell(var, lev, 0),
      `Repair (n, %)`    = wt_cell(var, lev, 1),
      `Total (n)`        = wt_cell_g1(var, lev),
      `Repair Rate (%)`  = repair_rate(var, lev)
    )
  })
  bind_rows(hdr, sub)
}
```

# Sheet 1 — National Estimates Table

```{r sheet1}
#| cache: false
total_rate <- paste0(round(100*total_repair_w/total_g1_w, 1), "%")

year_rows <- bind_rows(
  tibble(
    Characteristic     = "Total cases",
    `No Repair (n, %)` = paste0(format(round(total_norepair_w),big.mark=",")," (100%)"),
    `Repair (n, %)`    = paste0(format(round(total_repair_w),  big.mark=",")," (100%)"),
    `Total (n)`        = format(round(total_g1_w), big.mark=","),
    `Repair Rate (%)`  = total_rate
  ),
  map_dfr(as.character(2016:2023), function(yr) {
    idx_yr  <- enc_all$year==yr
    idx_rep <- idx_yr & enc_all$had_repair==1
    n_yr    <- sum(enc_all$discwt[idx_yr],  na.rm=TRUE)
    n_rep   <- sum(enc_all$discwt[idx_rep], na.rm=TRUE)
    tibble(
      Characteristic     = paste0("   ", yr),
      `No Repair (n, %)` = wt_cell("year", yr, 0),
      `Repair (n, %)`    = wt_cell("year", yr, 1),
      `Total (n)`        = format(round(n_yr), big.mark=","),
      `Repair Rate (%)`  = paste0(round(100*n_rep/n_yr, 1), "%")
    )
  })
)

sheet1 <- bind_rows(
  year_rows,
  tibble(
    Characteristic     = "Age, mean \u00b1 SD",
    `No Repair (n, %)` = mean_sd_cell(0),
    `Repair (n, %)`    = mean_sd_cell(1),
    `Total (n)`        = mean_sd_cell_g1(),
    `Repair Rate (%)`  = NA_character_
  ),
  make_rows("sex",         c("Male","Female"),                "Sex"),
  make_rows("age_group",   age_labels,                        "Age group"),
  make_rows("race",
    c("White","Black","Hispanic",
      "Asian or Pacific Islander or Native American or Other"), "Race"),
  make_rows("income_qrtl",
    c("0-25th","26th to 50th","51st to 75th","76th to 100th"),
    "Median household income of patients zipcode by quartile"),
  make_rows("insurance",
    c("Medicare","Medicaid","Private insurance","Self-pay","Other"),
    "Insurance type"),
  make_rows("hosp_region", c("Northeast","Midwest","South","West"),
            "Hospital region"),
  make_rows("hosp_location",c("Rural","Urban"),               "Hospital location"),
  make_rows("hosp_teach",  c("Nonteaching","Teaching"),       "Hospital teach")
)

sheet1 %>%
  kable(caption=paste0(
    "National Estimates of Patient and Hospital Characteristics by ",
    "Facial Nerve Repair Status, Including Repair Rates. ",
    "HCUP NASS 2016\u20132023. Weighted national estimates (DISCWT). ",
    "Total = All patients with traumatic facial nerve laceration nationally.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  row_spec(which(is.na(sheet1$`No Repair (n, %)`)), bold=TRUE)
```

# Sheet 2 — Multivariable Logistic Regression

```{r sheet2}
#| cache: true
reg_data <- enc_all %>%
  filter(!is.na(sex), !is.na(age_group), !is.na(race),
         !is.na(income_qrtl), !is.na(insurance),
         !is.na(hosp_region), !is.na(hosp_location), !is.na(hosp_teach)) %>%
  mutate(
    had_repair    = as.integer(had_repair),
    discwt        = as.numeric(discwt),
    sex           = relevel(factor(sex),           ref="Male"),
    age_group     = relevel(factor(age_group, levels=age_labels), ref="18-34"),
    race          = relevel(factor(race),           ref="White"),
    income_qrtl   = relevel(factor(income_qrtl,
                      levels=c("0-25th","26th to 50th",
                               "51st to 75th","76th to 100th")),
                      ref="76th to 100th"),
    insurance     = relevel(factor(insurance),      ref="Private insurance"),
    hosp_region   = relevel(factor(hosp_region),   ref="South"),
    hosp_location = relevel(factor(hosp_location), ref="Urban"),
    hosp_teach    = relevel(factor(hosp_teach),     ref="Nonteaching")
  )

svy_reg <- svydesign(ids=~hosp_id, weights=~discwt, data=reg_data, nest=TRUE)

model <- svyglm(
  had_repair ~ sex + age_group + race + income_qrtl + insurance +
               hosp_region + hosp_location + hosp_teach,
  design=svy_reg, family=quasibinomial(link="logit")
)

coef_df      <- as.data.frame(summary(model)$coefficients)
coef_df$term <- rownames(coef_df)

or_df <- coef_df %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR    = exp(Estimate),
    lower = exp(Estimate - 1.96 * `Std. Error`),
    upper = exp(Estimate + 1.96 * `Std. Error`),
    OR_CI = paste0(sprintf("%.2f",OR)," (",
                   sprintf("%.2f",lower),"\u2013",
                   sprintf("%.2f",upper),")"),
    p_fmt = fmt_p(`Pr(>|t|)`)
  )

term_map <- c(
  "sexFemale"="Female",
  "age_group35-44"="35-44","age_group45-54"="45-54",
  "age_group55-65"="55-65","age_group65+"="65+",
  "raceBlack"="Black","raceHispanic"="Hispanic",
  "raceAsian or Pacific Islander or Native American or Other"=
    "Asian or Pacific Islander or Native American or Other",
  "income_qrtl0-25th"="0-25th",
  "income_qrtl26th to 50th"="26th to 50th",
  "income_qrtl51st to 75th"="51st to 75th",
  "insuranceMedicare"="Medicare","insuranceMedicaid"="Medicaid",
  "insuranceSelf-pay"="Self-pay","insuranceOther"="Other",
  "hosp_regionNortheast"="Northeast","hosp_regionMidwest"="Midwest",
  "hosp_regionWest"="West",
  "hosp_locationRural"="Rural",
  "hosp_teachTeaching"="Teaching"
)

build_or_group <- function(label, levels_ord, ref) {
  hdr  <- tibble(Characteristic=label,
                 `Adjusted OR (95% CI)`=NA_character_,
                 `p value`=NA_character_)
  rows <- map_dfr(levels_ord, function(lev) {
    if (lev==ref) {
      tibble(Characteristic=paste0("   ",lev),
             `Adjusted OR (95% CI)`="Reference", `p value`=NA_character_)
    } else {
      key <- names(term_map)[term_map==lev]
      r   <- or_df %>% filter(term %in% key)
      tibble(Characteristic=paste0("   ",lev),
             `Adjusted OR (95% CI)`=if(nrow(r)>0) r$OR_CI[1] else NA_character_,
             `p value`=if(nrow(r)>0) r$p_fmt[1] else NA_character_)
    }
  })
  bind_rows(hdr, rows)
}

sheet2 <- bind_rows(
  build_or_group("Sex",         c("Male","Female"),              "Male"),
  build_or_group("Age group",   age_labels,                      "18-34"),
  build_or_group("Race",
    c("White","Black","Hispanic",
      "Asian or Pacific Islander or Native American or Other"),  "White"),
  build_or_group("Median household income of patients zipcode by quartile",
    c("0-25th","26th to 50th","51st to 75th","76th to 100th"), "76th to 100th"),
  build_or_group("Insurance type",
    c("Medicare","Medicaid","Private insurance","Self-pay","Other"),
    "Private insurance"),
  build_or_group("Hospital region",
    c("Northeast","Midwest","South","West"),                     "South"),
  build_or_group("Hospital location", c("Rural","Urban"),        "Urban"),
  build_or_group("Hospital teach",    c("Nonteaching","Teaching"),"Nonteaching")
)

# Build the dynamic "significant predictors" string for footnotes
sig_predictors <- or_df %>%
  filter(`Pr(>|t|)` < 0.05, term %in% names(term_map)) %>%
  mutate(label = term_map[term]) %>%
  pull(label)
sig_pred_text <- if (length(sig_predictors) > 0) {
  paste(sig_predictors, collapse=", ")
} else {
  "none at p<0.05"
}

sheet2 %>%
  kable(caption="Multivariable Logistic Regression for Odds of Facial Nerve Repair.") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  row_spec(which(is.na(sheet2$`Adjusted OR (95% CI)`) |
                   sheet2$`Adjusted OR (95% CI)`==""), bold=TRUE)
```

# Sheet 4 — Repair Rates by Year

```{r sheet4}
#| cache: false
sheet4 <- enc_all %>%
  group_by(year) %>%
  summarise(
    `Traumatic facial nerve injury (unweighted N)` = n(),
    `Traumatic facial nerve injury (weighted N)`   = round(sum(discwt, na.rm=TRUE)),
    `Repair (unweighted N)`      = sum(had_repair),
    `Repair (weighted N)`        = round(sum(discwt[had_repair==1], na.rm=TRUE)),
    `Repair rate — weighted (%)` = round(
      100*sum(discwt[had_repair==1],na.rm=TRUE)/sum(discwt,na.rm=TRUE),1),
    .groups="drop"
  ) %>% rename(Year=year)

sheet4 %>%
  kable(caption="Repair Rates by Year. HCUP NASS 2016-2023.",
        format.args=list(big.mark=",")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)

# Survey-weighted chi-square test for repair rate difference across years
year_chisq <- svychisq(~had_repair + year,
                        design = svy,
                        statistic = "Chisq")
year_p <- fmt_p(year_chisq$p.value)

message("Chi-square test for repair rate across years: p = ", year_p)
```

# Sheet 5 — Charges Analysis

```{r sheet5}
#| cache: false
charges_data <- enc_all %>% filter(!is.na(totchg), totchg > 0)

overall_row <- tibble(
  Characteristic                                 = "Total charges of encounter",
  `No Repair Median (IQR)`                       = wt_median_iqr(charges_data, 0),
  `Repair Median (IQR)`                          = wt_median_iqr(charges_data, 1),
  `p value (no repair: within-group comparison)` = NA_character_,
  `p value (repair: within-group comparison)`    = NA_character_
)

ins_levels      <- c("Medicare","Medicaid","Private insurance","Self-pay","Other")
ins_anova_norep <- fmt_p(svy_anova_charges_p(0, "insurance"))
ins_anova_rep   <- fmt_p(svy_anova_charges_p(1, "insurance"))

ins_hdr <- tibble(
  Characteristic                                 = "Insurance type",
  `No Repair Median (IQR)`                       = NA_character_,
  `Repair Median (IQR)`                          = NA_character_,
  `p value (no repair: within-group comparison)` = ins_anova_norep,
  `p value (repair: within-group comparison)`    = ins_anova_rep
)

ins_rows <- map_dfr(ins_levels, function(lev) {
  d <- charges_data %>% filter(!is.na(insurance), insurance==lev)
  tibble(
    Characteristic                                 = paste0("   ", lev),
    `No Repair Median (IQR)`                       = wt_median_iqr(d, 0),
    `Repair Median (IQR)`                          = wt_median_iqr(d, 1),
    `p value (no repair: within-group comparison)` = NA_character_,
    `p value (repair: within-group comparison)`    = NA_character_
  )
})

teach_ttest_norep <- fmt_p(svy_ttest_charges_p(0, "hosp_teach"))
teach_ttest_rep   <- fmt_p(svy_ttest_charges_p(1, "hosp_teach"))

teach_hdr <- tibble(
  Characteristic                                 = "Hospital teaching status",
  `No Repair Median (IQR)`                       = NA_character_,
  `Repair Median (IQR)`                          = NA_character_,
  `p value (no repair: within-group comparison)` = teach_ttest_norep,
  `p value (repair: within-group comparison)`    = teach_ttest_rep
)

teach_rows <- map_dfr(c("Nonteaching","Teaching"), function(lev) {
  d <- charges_data %>% filter(!is.na(hosp_teach), hosp_teach==lev)
  tibble(
    Characteristic                                 = paste0("   ", lev),
    `No Repair Median (IQR)`                       = wt_median_iqr(d, 0),
    `Repair Median (IQR)`                          = wt_median_iqr(d, 1),
    `p value (no repair: within-group comparison)` = NA_character_,
    `p value (repair: within-group comparison)`    = NA_character_
  )
})

sheet5 <- bind_rows(overall_row, ins_hdr, ins_rows, teach_hdr, teach_rows)

sheet5 %>%
  kable(caption=paste0(
    "Charge Analysis: Repair vs No Repair. Median (IQR) weighted national estimates. ",
    "Insurance type p-values: survey-weighted ANOVA on log-charges within each group. ",
    "Teaching status p-values: survey-weighted t-test on log-charges within each group.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE) %>%
  row_spec(which(is.na(sheet5$`No Repair Median (IQR)`)), bold=TRUE)
```

# NIS 2023 — Facial Nerve Repair (Inpatient)

```{r nis-data}
#| cache: false
nis_sheet <- tibble(
  `ICD-10-PCS Code` = c(
    "00QM0ZZ",
    "00QM3ZZ",
    "00QM4ZZ",
    "01QM0ZZ",
    "01QM3ZZ",
    "01QM4ZZ",
    "TOTAL (PNS codes)"
  ),
  Description = c(
    "Repair Facial Nerve, Open Approach (CNS section \u2014 incorrect)",
    "Repair Facial Nerve, Percutaneous Approach (CNS section \u2014 incorrect)",
    "Repair Facial Nerve, Percutaneous Endoscopic Approach (CNS section \u2014 incorrect)",
    "Repair Facial Nerve, Open Approach (PNS section \u2014 correct)",
    "Repair Facial Nerve, Percutaneous Approach (PNS section \u2014 correct)",
    "Repair Facial Nerve, Percutaneous Endoscopic Approach (PNS section \u2014 correct)",
    "All three PNS facial nerve repair codes combined"
  ),
  `Unweighted N` = c(0L, 0L, 0L, NA_integer_, NA_integer_, NA_integer_, 3L),
  `Weighted N (national estimate)` = c(0L, 0L, 0L, NA_integer_, NA_integer_, NA_integer_, 15L)
)

nis_sheet %>%
  kable(caption=paste0(
    "NIS 2023 Facial Nerve Repair \u2014 Inpatient Database Results. ",
    "Source: HCUP National Inpatient Sample (NIS) 2023. ",
    "Carried forward from primary analysis. ",
    "Individual PNS code breakdown not reported per HCUP cell size restrictions (<11). ",
    "NA = suppressed. CNS = Central Nervous System (section 00); ",
    "PNS = Peripheral Nervous System (section 01).")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE) %>%
  row_spec(7, bold=TRUE, background="#EBF4FB") %>%
  row_spec(1:3, color="grey")
```

# Publication Figures

## Figure 1 — Study Population Flow Diagram

```{r fig-flow}
g1_n <- nrow(enc_all)
g1_w <- round(total_g1_w)
g2_n <- sum(enc_all$had_repair==1)
g2_w <- round(total_repair_w)
g3_n <- sum(enc_all$had_repair==0)
g3_w <- round(total_norepair_w)

flow <- grViz(paste0('
digraph flow {
  graph [layout=dot, rankdir=TB, fontname="Arial", nodesep=0.7, ranksep=0.9]
  node [shape=rectangle, style="filled,rounded", fontname="Arial",
        fontsize=11, margin="0.25,0.15", penwidth=1.5]

  A [label="HCUP NASS 2016-2023\\nFull Ambulatory Surgery Sample\\n(all encounter years combined)",
     fillcolor="#1A5276", fontcolor=white, color="#1A5276"]

  B [label="Group 1: Traumatic Facial Nerve Laceration\\nUnweighted N = ', format(g1_n, big.mark=","), '\\nWeighted national estimate = ', format(g1_w, big.mark=","), '\\n\\nInclusion criteria (either A or B):\\nA. S04.5 family code co-occurring with\\n    S01.* (head) or S11.* (neck) open wound\\n    OR external cause code (W25-W27, W45,\\n    X99, Y28, V01-V99)\\nB. CPT 64864 or 64865 co-occurring with\\n    S01.* or S11.* open wound code",
     fillcolor="#2E86C1", fontcolor=white, color="#2175A9"]

  C [label="Group 2: Traumatic Facial Nerve Laceration WITH Repair\\nUnweighted N = ', format(g2_n, big.mark=","), '  |  Weighted N = ', format(g2_w, big.mark=","), '\\n\\nInclusion criteria:\\nAny encounter from Group 1\\nco-occurring with CPT for facial nerve suture\\n(CPT 64864 or CPT 64865)",
     fillcolor="#EBF5FB", fontcolor="#1A5276", color="#2E86C1"]

  D [label="Group 3: Traumatic Facial Nerve Laceration WITHOUT Repair\\nUnweighted N = ', format(g3_n, big.mark=","), '  |  Weighted N = ', format(g3_w, big.mark=","), '\\n\\nInclusion criteria:\\nGroup 1 minus Group 2\\n(no qualifying CPT code for facial nerve suture present)",
     fillcolor="#FEF9E7", fontcolor="#7D6608", color="#D4AC0D"]

  A -> B [label="  Injury identification", fontname="Arial", fontsize=10,
          color="#566573", style=bold]
  B -> C [label="  CPT repair\\n  code present", fontname="Arial",
          fontsize=10, color="#2E86C1", style=bold, fontcolor="#2E86C1"]
  B -> D [label="  No CPT repair\\n  code present", fontname="Arial",
          fontsize=10, color="#D4AC0D", style=bold, fontcolor="#D4AC0D"]
}'))

flow
```

```{r save-flow}
#| cache: false
tryCatch({
  svg_txt  <- export_svg(flow)
  svg_path <- tempfile(fileext=".svg")
  writeLines(svg_txt, svg_path)
  rsvg_png(svg_path,
           file  = file.path(fig_dir, "fig1_flow_diagram.png"),
           width = 1200)
  rsvg_pdf(svg_path,
           file  = file.path(fig_dir, "fig1_flow_diagram.pdf"))
  message("Flow diagram saved as PNG and PDF.")
}, error = function(e) {
  message("Flow diagram save error: ", e$message)
})
```

## Figure 2 — Repair Trend by Year

```{r fig-trend}
trend_data <- sheet4 %>%
  mutate(
    Year     = as.integer(Year),
    repair_w = `Repair (weighted N)`,
    injury_w = `Traumatic facial nerve injury (weighted N)`,
    pct      = `Repair rate — weighted (%)`
  )

max_w <- max(trend_data$injury_w, na.rm=TRUE)

fig_trend <- ggplot(trend_data, aes(x=Year)) +
  geom_col(aes(y=injury_w, fill="All traumatic facial nerve laceration"),
           width=0.6, alpha=0.5) +
  geom_col(aes(y=repair_w, fill="With repair"), width=0.6) +
  geom_line(aes(y=pct*max_w/100), color="#C00000",
            linewidth=1.2, linetype="dashed") +
  geom_point(aes(y=pct*max_w/100), color="#C00000", size=3) +
  geom_text(aes(y=pct*max_w/100, label=paste0(pct,"%")),
            vjust=-0.9, size=3.5, color="#C00000", fontface="bold") +
  scale_y_continuous(
    name=  "Weighted encounters (N)", labels=label_comma(),
    sec.axis=sec_axis(~.*100/max_w, name="Repair rate (%)",
                      labels=function(x) paste0(x,"%"))) +
  scale_x_continuous(breaks=2016:2023) +
  scale_fill_manual(values=c("All traumatic facial nerve laceration"="#4E79A7",
                              "With repair"="#F28E2B")) +
  labs(title="Traumatic Facial Nerve Laceration Encounters and Repair Rate by Year",
       subtitle="HCUP NASS 2016\u20132023 | Weighted national estimates",
       x="Year", fill=NULL,
       caption="Dashed red line = weighted repair rate (right axis)") +
  theme_classic(base_size=12) +
  theme(legend.position="bottom", plot.title=element_text(face="bold"),
        axis.title=element_text(face="bold"))

fig_trend
ggsave(file.path(fig_dir,"fig2_repair_trend.pdf"),
       fig_trend, width=10, height=6)
ggsave(file.path(fig_dir,"fig2_repair_trend.png"),
       fig_trend, width=10, height=6, dpi=300)
```

## Figure 3 — Forest Plot

```{r fig-forest}
#| cache: false
forest_data <- or_df %>%
  mutate(
    label = term_map[term],
    group = case_when(
      str_starts(term,"sex")          ~ "Sex",
      str_starts(term,"age_group")    ~ "Age group",
      str_starts(term,"race")         ~ "Race",
      str_starts(term,"income_qrtl")  ~ "Income quartile",
      str_starts(term,"insurance")    ~ "Insurance type",
      str_starts(term,"hosp_region")  ~ "Hospital region",
      TRUE ~ "Other")
  ) %>%
  filter(!is.na(label),
         !str_starts(term, "hosp_location"),
         !str_starts(term, "hosp_teach")) %>%
  mutate(label=factor(label, levels=rev(unique(label))),
         sig=`Pr(>|t|)` < 0.05)

fig_forest <- ggplot(forest_data,
                     aes(x=OR, y=label, xmin=lower, xmax=upper, color=sig)) +
  geom_vline(xintercept=1, linetype="dashed", color="grey50", linewidth=0.8) +
  geom_errorbarh(height=0.3, linewidth=0.7) +
  geom_point(size=3) +
  geom_text(aes(x=upper,
                label=paste0(sprintf("%.2f",OR)," (",
                             sprintf("%.2f",lower),"\u2013",
                             sprintf("%.2f",upper),")")),
            hjust=-0.1, size=2.8, color="black") +
  scale_color_manual(values=c("FALSE"="grey50","TRUE"="#C00000"),
                     labels=c("FALSE"="p\u22650.05","TRUE"="p<0.05"),name="") +
  scale_x_log10(limits=c(0.05,25),
                breaks=c(0.1,0.25,0.5,1,2,5,10),
                labels=c("0.1","0.25","0.5","1","2","5","10")) +
  facet_grid(group~., scales="free_y", space="free_y") +
  labs(title="Multivariable Logistic Regression for Odds of Facial Nerve Repair",
       subtitle="Adjusted odds ratios | HCUP NASS 2016\u20132023",
       x="Adjusted Odds Ratio (log scale)", y=NULL,
       caption=paste0("Reference: Male, Age 18-34, White, 76th-100th income quartile,\n",
                      "Private insurance, South region. Error bars = 95% CI.\n",
                      "Hospital location and teaching status included in model but",
                      " not shown.")) +
  theme_classic(base_size=11) +
  theme(strip.text=element_text(face="bold", size=9),
        strip.background=element_rect(fill="#EBF4FB", color=NA),
        plot.title=element_text(face="bold"), legend.position="bottom")

fig_forest
ggsave(file.path(fig_dir,"fig3_forest_plot.pdf"),
       fig_forest, width=10, height=10)
ggsave(file.path(fig_dir,"fig3_forest_plot.png"),
       fig_forest, width=10, height=10, dpi=300)
```

## Figure 4 — Geographic Heat Map

```{r fig-heatmap}
region_data <- enc_all %>%
  filter(!is.na(hosp_region)) %>%
  group_by(hosp_region) %>%
  summarise(repair_rate_w=round(
    100*sum(discwt[had_repair==1],na.rm=TRUE)/sum(discwt,na.rm=TRUE),2),
    .groups="drop")

state_region <- tibble(
  state=c("connecticut","maine","massachusetts","new hampshire","rhode island",
          "vermont","new jersey","new york","pennsylvania",
          "illinois","indiana","michigan","ohio","wisconsin","iowa","kansas",
          "minnesota","missouri","nebraska","north dakota","south dakota",
          "delaware","florida","georgia","maryland","north carolina",
          "south carolina","virginia","west virginia","alabama","kentucky",
          "mississippi","tennessee","arkansas","louisiana","oklahoma","texas",
          "district of columbia","arizona","colorado","idaho","montana",
          "nevada","new mexico","utah","wyoming","alaska","california",
          "hawaii","oregon","washington"),
  hosp_region=c(rep("Northeast",9),rep("Midwest",12),
                rep("South",17),   rep("West",13))
)

us_map <- map_data("state") %>%
  left_join(state_region, by=c("region"="state")) %>%
  left_join(region_data, by="hosp_region")

region_centers <- us_map %>%
  group_by(hosp_region, repair_rate_w) %>%
  summarise(long=mean(long,na.rm=TRUE), lat=mean(lat,na.rm=TRUE),
            .groups="drop") %>%
  filter(!is.na(hosp_region))

fig_map <- ggplot(us_map, aes(x=long, y=lat, group=group, fill=repair_rate_w)) +
  geom_polygon(color="white", linewidth=0.3) +
  geom_text(data=region_centers,
            aes(x=long, y=lat,
                label=paste0(hosp_region,"\n",repair_rate_w,"%"), group=NULL),
            size=3.5, fontface="bold", color="white") +
  scale_fill_gradient(low="#D6EAF8", high="#1A5276",
                      name="Repair rate\n(weighted %)", na.value="grey85") +
  coord_map("albers", lat0=39, lat1=45) +
  labs(title="Traumatic Facial Nerve Repair Rate by U.S. Census Region",
       subtitle="HCUP NASS 2016\u20132023 | Weighted national estimates",
       caption=paste0("State shading reflects the repair rate for its Census region.\n",
                      "State-level estimates suppressed per HCUP cell size restrictions.\n",
                      "NOTE: 2023 HOSP_REGION not available in NASS Hospital file; ",
                      "regional estimates reflect 2016\u20132022 data only."),
       x=NULL, y=NULL) +
  theme_void(base_size=12) +
  theme(plot.title=element_text(face="bold", hjust=0.5),
        plot.subtitle=element_text(hjust=0.5, color="grey40"),
        plot.caption=element_text(hjust=0.5, size=9, color="grey50"),
        legend.position="right")

fig_map
ggsave(file.path(fig_dir,"fig4_heatmap_region.pdf"),
       fig_map, width=12, height=7)
ggsave(file.path(fig_dir,"fig4_heatmap_region.png"),
       fig_map, width=12, height=7, dpi=300)
```

## Figure 5 — IPTW: Effect of Repair on Total Charges

```{r fig-iptw}
#| cache: true
iptw_data <- enc_all %>%
  filter(!is.na(totchg), totchg>0,
         !is.na(sex), !is.na(age_group), !is.na(race),
         !is.na(income_qrtl), !is.na(insurance),
         !is.na(hosp_region), !is.na(hosp_location), !is.na(hosp_teach)) %>%
  mutate(
    had_repair    = as.integer(had_repair),
    sex           = factor(sex),
    age_group     = factor(age_group, levels=age_labels),
    race          = factor(race),
    income_qrtl   = factor(income_qrtl),
    insurance     = factor(insurance),
    hosp_region   = factor(hosp_region),
    hosp_location = factor(hosp_location),
    hosp_teach    = factor(hosp_teach)
  )

set.seed(42)
w_out <- weightit(
  had_repair ~ sex + age_group + race + income_qrtl + insurance +
               hosp_region + hosp_location + hosp_teach,
  data=iptw_data, method="ps", estimand="ATE"
)

iptw_data$iptw_weight <- w_out$weights * iptw_data$discwt

svy_iptw   <- svydesign(ids=~hosp_id, weights=~iptw_weight,
                         data=iptw_data, nest=TRUE)
iptw_model <- svyglm(log(totchg) ~ had_repair,
                     design=svy_iptw, family=gaussian())

iptw_coef <- summary(iptw_model)$coefficients
eff_pct   <- round((exp(iptw_coef["had_repair","Estimate"])-1)*100, 1)
ci_lo     <- round((exp(iptw_coef["had_repair","Estimate"] -
                        1.96*iptw_coef["had_repair","Std. Error"])-1)*100,1)
ci_hi     <- round((exp(iptw_coef["had_repair","Estimate"] +
                        1.96*iptw_coef["had_repair","Std. Error"])-1)*100,1)
iptw_p    <- fmt_p(iptw_coef["had_repair","Pr(>|t|)"])

iptw_excel <- tibble(
  Metric = c(
    "Method",
    "Outcome variable",
    "Treatment variable",
    "Covariates balanced",
    "IPTW estimand",
    "No repair \u2014 median total charges (unweighted)",
    "With repair \u2014 median total charges (unweighted)",
    "Effect of repair on charges (IPTW-adjusted)",
    "95% Confidence interval",
    "p value",
    "Interpretation"
  ),
  Value = c(
    "Inverse Probability of Treatment Weighting (IPTW) using propensity scores (WeightIt package)",
    "Log total charges",
    "Had repair (CPT 64864 or 64865) vs no repair",
    "Sex, age group, race, income quartile, insurance type, hospital region, location, teaching status",
    "Average Treatment Effect (ATE)",
    paste0("$", format(round(median(iptw_data$totchg[iptw_data$had_repair==0],
                                    na.rm=TRUE)), big.mark=",")),
    paste0("$", format(round(median(iptw_data$totchg[iptw_data$had_repair==1],
                                    na.rm=TRUE)), big.mark=",")),
    paste0(eff_pct, "% higher charges associated with repair"),
    paste0(ci_lo, "% to ", ci_hi, "%"),
    iptw_p,
    paste0("After balancing patient and hospital characteristics via IPTW, ",
           "facial nerve repair was associated with a ", eff_pct,
           "% increase in total charges (95% CI: ", ci_lo, "% to ", ci_hi,
           "%, p=", iptw_p, ").")
  )
)

fig_iptw <- ggplot(iptw_data, aes(x=log(totchg), fill=factor(had_repair))) +
  geom_density(alpha=0.6) +
  scale_fill_manual(values=c("0"="#4E79A7","1"="#F28E2B"),
                    labels=c("0"="No repair","1"="With repair"), name="") +
  labs(title="Distribution of Log Total Charges by Repair Status",
       subtitle=paste0("IPTW-adjusted. Repair associated with ",
                       eff_pct,"% change in charges (p=",iptw_p,")"),
       x="Log total charges", y="Density") +
  theme_classic(base_size=12) +
  theme(plot.title=element_text(face="bold"), legend.position="bottom")

fig_iptw
ggsave(file.path(fig_dir,"fig5_iptw_charges.pdf"),
       fig_iptw, width=9, height=6)
ggsave(file.path(fig_dir,"fig5_iptw_charges.png"),
       fig_iptw, width=9, height=6, dpi=300)
```

# Sensitivity Analyses

## Sensitivity Analysis 1 — Criterion A Only (no external cause, no Criterion B)

```{r sensitivity-1}
#| cache: false
# Criterion A restricted to S04.5 + (S01.* OR S11.*), excluding the external
# cause path AND excluding Criterion B (procedural pathway).
sens1_data <- enc_all %>%
  filter(flag_s045 & (flag_s01 | flag_s11))

sens1_g1_unw <- nrow(sens1_data)
sens1_g1_w   <- round(sum(sens1_data$discwt, na.rm=TRUE))
sens1_g2_unw <- sum(sens1_data$had_repair == 1)
sens1_g2_w   <- round(sum(sens1_data$discwt[sens1_data$had_repair == 1], na.rm=TRUE))
sens1_g3_unw <- sens1_g1_unw - sens1_g2_unw
sens1_g3_w   <- sens1_g1_w - sens1_g2_w
sens1_rate   <- if (sens1_g1_w > 0) round(100*sens1_g2_w/sens1_g1_w, 1) else NA_real_

sens1_tbl <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate (%)"
  ),
  Value = c(
    format(sens1_g1_unw, big.mark=","),
    format(sens1_g1_w,   big.mark=","),
    format(sens1_g2_unw, big.mark=","),
    format(sens1_g2_w,   big.mark=","),
    format(sens1_g3_unw, big.mark=","),
    format(sens1_g3_w,   big.mark=","),
    paste0(sens1_rate, "%")
  )
)

sens1_tbl %>%
  kable(caption=paste0(
    "Sensitivity Analysis 1 \u2014 Criterion A only. ",
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, ",
    "excluding the external-cause-only path and excluding the Criterion B ",
    "procedural pathway. HCUP NASS 2016\u20132023.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

## Sensitivity Analysis 2 — Highest Mechanistic Specificity

```{r sensitivity-2}
#| cache: false
# Highest mechanistic specificity: S04.5 + (S01.* OR S11.*) AND external cause
sens2_data <- enc_all %>%
  filter(flag_s045 & (flag_s01 | flag_s11) & flag_extcause)

sens2_g1_unw <- nrow(sens2_data)
sens2_g1_w   <- round(sum(sens2_data$discwt, na.rm=TRUE))
sens2_g2_unw <- sum(sens2_data$had_repair == 1)
sens2_g2_w   <- round(sum(sens2_data$discwt[sens2_data$had_repair == 1], na.rm=TRUE))
sens2_g3_unw <- sens2_g1_unw - sens2_g2_unw
sens2_g3_w   <- sens2_g1_w - sens2_g2_w
sens2_rate   <- if (sens2_g1_w > 0) round(100*sens2_g2_w/sens2_g1_w, 1) else NA_real_

sens2_tbl <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate (%)"
  ),
  Value = c(
    format(sens2_g1_unw, big.mark=","),
    format(sens2_g1_w,   big.mark=","),
    format(sens2_g2_unw, big.mark=","),
    format(sens2_g2_w,   big.mark=","),
    format(sens2_g3_unw, big.mark=","),
    format(sens2_g3_w,   big.mark=","),
    paste0(sens2_rate, "%")
  )
)

sens2_tbl %>%
  kable(caption=paste0(
    "Sensitivity Analysis 2 \u2014 Highest mechanistic specificity. ",
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
    "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "HCUP NASS 2016\u20132023.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

## Sensitivity Analysis 3: Maximum specificity

```{r sensitivity-3}
#| cache: false

# ── Sensitivity Analysis 3: Maximum specificity ───────────────────────────────
# Definition: S04.51* or S04.52* (laterality specified, right or left)
#             PLUS anatomically concordant facial nerve distribution wound codes
#             PLUS external cause sharp object or transport mechanism code
#
# S04.51* = injury of facial nerve, right side
# S04.52* = injury of facial nerve, left side
# Anatomic wound codes: S01.1* (eyelid/periocular), S01.3* (ear),
#                       S01.4* (cheek/TMJ), S01.5* (lip/oral cavity),
#                       S01.81* (other part of face)
# External cause: W25-W27 (sharp object), W45 (foreign body through skin),
#                 X99 (assault with sharp object), Y28 (sharp object
#                 undetermined intent), V01-V99 (transport accidents)

dx_cols  <- grep("^i10_dx",   names(enc_all), value=TRUE, ignore.case=TRUE)
cpt_cols <- grep("^cpt",      names(enc_all), value=TRUE, ignore.case=TRUE)
ec_cols  <- grep("^i10_ecause|^e_cause|^externalcause",
                 names(enc_all), value=TRUE, ignore.case=TRUE)

# Laterality-specified facial nerve injury codes
s3_nerve <- function(row) {
  any(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc_all[[col]][row]))
    str_starts(cd, "S0451") | str_starts(cd, "S0452")
  }))
}

# Anatomically relevant facial wound codes
s3_wound <- function(row) {
  any(sapply(dx_cols, function(col) {
    cd <- trimws(as.character(enc_all[[col]][row]))
    str_starts(cd, "S011") |   # eyelid/periocular
    str_starts(cd, "S013") |   # ear
    str_starts(cd, "S014") |   # cheek/TMJ
    str_starts(cd, "S015") |   # lip/oral cavity
    str_starts(cd, "S0181")    # other part of face
  }))
}

# External cause sharp object or transport codes
s3_cause <- function(row) {
  if (length(ec_cols) == 0) return(FALSE)
  any(sapply(ec_cols, function(col) {
    cd <- trimws(as.character(enc_all[[col]][row]))
    str_starts(cd, "W25") | str_starts(cd, "W26") |
    str_starts(cd, "W27") | str_starts(cd, "W45") |
    str_starts(cd, "X99") | str_starts(cd, "Y28") |
    str_starts(cd, "V")
  }))
}

# Apply vectorized across all rows
message("Running Sensitivity Analysis 3 extraction...")
nerve_flag <- rowSums(sapply(dx_cols, function(col) {
  cd <- trimws(as.character(enc_all[[col]]))
  str_starts(cd, "S0451") | str_starts(cd, "S0452")
})) > 0

wound_flag <- rowSums(sapply(dx_cols, function(col) {
  cd <- trimws(as.character(enc_all[[col]]))
  str_starts(cd, "S011") | str_starts(cd, "S013") |
  str_starts(cd, "S014") | str_starts(cd, "S015") |
  str_starts(cd, "S0181")
})) > 0

if (length(ec_cols) > 0) {
  cause_flag <- rowSums(sapply(ec_cols, function(col) {
    cd <- trimws(as.character(enc_all[[col]]))
    str_starts(cd, "W25") | str_starts(cd, "W26") |
    str_starts(cd, "W27") | str_starts(cd, "W45") |
    str_starts(cd, "X99") | str_starts(cd, "Y28") |
    str_starts(cd, "V")
  })) > 0
} else {
  message("No external cause columns found — cause_flag set to FALSE for all rows")
  cause_flag <- rep(FALSE, nrow(enc_all))
}

# Group 1 SA3: laterality-specified nerve + anatomic wound + external cause
sa3_g1 <- nerve_flag & wound_flag & cause_flag

# Group 2 SA3: Group 1 plus CPT repair code
cpt_flag_sa3 <- rowSums(sapply(cpt_cols, function(col) {
  trimws(as.character(enc_all[[col]])) %in% c("64864","64865")
})) > 0

sa3_g2 <- sa3_g1 & cpt_flag_sa3
sa3_g3 <- sa3_g1 & !cpt_flag_sa3

# Compute weighted estimates
sa3_data <- enc_all[sa3_g1, ]
sa3_g1_uw <- sum(sa3_g1)
sa3_g1_w  <- round(sum(enc_all$discwt[sa3_g1], na.rm=TRUE))
sa3_g2_uw <- sum(sa3_g2)
sa3_g2_w  <- round(sum(enc_all$discwt[sa3_g2], na.rm=TRUE))
sa3_g3_uw <- sum(sa3_g3)
sa3_g3_w  <- round(sum(enc_all$discwt[sa3_g3], na.rm=TRUE))
sa3_rate  <- round(100 * sa3_g2_w / sa3_g1_w, 1)

message("=== Sensitivity Analysis 3 Results ===")
message("Group 1 unweighted: ", format(sa3_g1_uw, big.mark=","))
message("Group 1 weighted:   ", format(sa3_g1_w,  big.mark=","))
message("Group 2 unweighted: ", format(sa3_g2_uw, big.mark=","))
message("Group 2 weighted:   ", format(sa3_g2_w,  big.mark=","))
message("Group 3 unweighted: ", format(sa3_g3_uw, big.mark=","))
message("Group 3 weighted:   ", format(sa3_g3_w,  big.mark=","))
message("Repair rate:        ", sa3_rate, "%")

# Build summary table
sa3_summary <- tibble(
  Definition = c(
    "Laterality-specified nerve injury (S04.51/52*)",
    "Plus anatomically concordant wound (S01.1/3/4/5/81*)",
    "Plus external cause code (W25-27, W45, X99, Y28, V*)",
    "Group 2: CPT 64864 or 64865 present"
  ),
  `Requirement` = c("Required","Required","Required","Required for Group 2")
)

sa3_results <- tibble(
  Metric = c(
    "Group 1 (all confirmed laceration) — unweighted N",
    "Group 1 (all confirmed laceration) — weighted N",
    "Group 2 (with repair) — unweighted N",
    "Group 2 (with repair) — weighted N",
    "Group 3 (without repair) — unweighted N",
    "Group 3 (without repair) — weighted N",
    "Overall repair rate (weighted)"
  ),
  Value = c(
    format(sa3_g1_uw, big.mark=","),
    format(sa3_g1_w,  big.mark=","),
    format(sa3_g2_uw, big.mark=","),
    format(sa3_g2_w,  big.mark=","),
    format(sa3_g3_uw, big.mark=","),
    format(sa3_g3_w,  big.mark=","),
    paste0(sa3_rate, "%")
  )
)

sa3_results %>%
  kable(caption=paste0(
    "Sensitivity Analysis 3: Maximum Specificity Definition. ",
    "Requires laterality-specified facial nerve injury code (S04.51* or S04.52*) ",
    "co-occurring with anatomically concordant facial wound code ",
    "(S01.1*, S01.3*, S01.4*, S01.5*, or S01.81*) ",
    "AND external cause sharp object or transport code. ",
    "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

Then add this to the `excel-output` chunk after the Sensitivity Analysis 2 sheet:

```r
# ── Sensitivity Analysis 3 Sheet ─────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 3")
writeData(wb, "Sensitivity Analysis 3",
          "Sensitivity Analysis 3: Maximum Specificity — Laterality + Anatomic Wound + External Cause",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 3", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 3", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 3", sa3_results,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 3", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 3", data_s,
         rows=4:(nrow(sa3_results)+3), cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 3", cols=1, widths=45)
setColWidths(wb, "Sensitivity Analysis 3", cols=2, widths=20)
fn_sa3 <- nrow(sa3_results)+5
writeData(wb, "Sensitivity Analysis 3",
  paste0("Sensitivity Analysis 3 definition: S04.51* or S04.52* (laterality-specified ",
         "facial nerve injury) co-occurring with S01.1*, S01.3*, S01.4*, S01.5*, or ",
         "S01.81* (anatomically relevant facial wound) AND external cause code from ",
         "W25\u2013W27, W45, X99, Y28, or V01\u2013V99 (sharp object or transport mechanism). ",
         "This represents the maximum specificity definition feasible in ICD-10-CM ",
         "administrative data for confirmed traumatic facial nerve laceration. ",
         "Note: external cause code completeness varies by institution and year; ",
         "cohort size reflects availability of all three code types simultaneously. ",
         "HCUP NASS 2016\u20132023."),
  startRow=fn_sa3, startCol=1)
mergeCells(wb, "Sensitivity Analysis 3", cols=1:2, rows=fn_sa3)
addStyle(wb, "Sensitivity Analysis 3", note_s, rows=fn_sa3, cols=1)
setRowHeights(wb, "Sensitivity Analysis 3", rows=fn_sa3, heights=80)
```

**One critical practical note before running**

The external cause code columns in NASS have variable names depending on the year — some years use `I10_ECAUSE1`, `I10_ECAUSE2` etc., others use different naming conventions. The code above searches for columns matching `i10_ecause`, `e_cause`, or `externalcause` patterns. Before running, check what external cause columns actually exist in your `enc_all` object by running this in the console:

```r
grep("cause|ecause|external", names(enc_all), 
     value=TRUE, ignore.case=TRUE)
```


# Phase 1 vs Phase 2 Comparison (Internal Quality Check)

```{r phase1-vs-phase2}
#| cache: false
# Internal investigator quality check only — does NOT appear in manuscript.
# Confirms Phase 2 produces a more complete cohort and that direction of
# significant predictors is consistent across definitions.

# Helper: pull a Phase 2 OR row by its display label
get_or_row <- function(label) {
  key <- names(term_map)[term_map == label]
  r   <- or_df %>% filter(term %in% key)
  if (nrow(r) == 0) return("not estimated")
  paste0(sprintf("%.2f",r$OR[1])," (95% CI ",
         sprintf("%.2f",r$lower[1]),"\u2013",
         sprintf("%.2f",r$upper[1]),"), p=", r$p_fmt[1])
}

phase2_overall_rate <- round(100*total_repair_w/total_g1_w, 1)
phase2_iptw_text    <- paste0(eff_pct, "%, p=", iptw_p)
phase2_teach_text   <- paste0("p=", teach_ttest_norep)

phase_compare <- tibble(
  Metric = c(
    "Group 1 unweighted N",
    "Group 1 weighted N",
    "Group 2 unweighted N",
    "Group 2 weighted N",
    "Group 3 unweighted N",
    "Group 3 weighted N",
    "Overall repair rate",
    "Age 55\u201365 OR",
    "Age 65+ OR",
    "Midwest OR",
    "West OR",
    "IPTW charge increase with repair",
    "Teaching vs nonteaching charge difference (no repair)"
  ),
  `Phase 1 Value` = c(
    "1,656",
    "2,276",
    "488",
    "664",
    "1,168",
    "1,613",
    "29.2%",
    "0.32 (95% CI 0.18\u20130.57), p<0.001",
    "0.41 (95% CI 0.20\u20130.85), p=0.018",
    "1.63 (95% CI 1.10\u20132.41), p=0.015",
    "1.81 (95% CI 1.05\u20133.10), p=0.033",
    "34.9%, p<0.001",
    "Significant, p<0.001"
  ),
  `Phase 2 Value` = c(
    format(nrow(enc_all), big.mark=","),
    format(round(total_g1_w), big.mark=","),
    format(sum(enc_all$had_repair==1), big.mark=","),
    format(round(total_repair_w), big.mark=","),
    format(sum(enc_all$had_repair==0), big.mark=","),
    format(round(total_norepair_w), big.mark=","),
    paste0(phase2_overall_rate, "%"),
    get_or_row("55-65"),
    get_or_row("65+"),
    get_or_row("Midwest"),
    get_or_row("West"),
    phase2_iptw_text,
    phase2_teach_text
  )
)

phase_compare %>%
  kable(caption=paste0(
    "Phase 1 vs Phase 2 Comparison \u2014 Internal Quality Check. ",
    "Investigator use only; does not appear in the Phase 2 manuscript.")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE)

# Sanity check on the bias-correction direction
if (phase2_overall_rate > 29.2) {
  warning("Phase 2 overall repair rate (", phase2_overall_rate,
          "%) is HIGHER than Phase 1 (29.2%). Per Phase 2 design rationale, ",
          "this should not happen — review extraction logic.")
} else {
  message("Phase 2 overall repair rate = ", phase2_overall_rate,
          "% (Phase 1 = 29.2%); bias-correction direction is as expected.")
}
```

# Excel Output

```{r excel-output}
#| cache: false
wb <- createWorkbook()

title_s  <- createStyle(fontName="Arial",fontSize=12,textDecoration="bold")
header_s <- createStyle(fontName="Arial",fontSize=11,textDecoration="bold",
                        border="TopBottom",borderStyle="medium",
                        halign="center",wrapText=TRUE)
group_s  <- createStyle(fontName="Arial",fontSize=11,textDecoration="bold")
data_s   <- createStyle(fontName="Arial",fontSize=11,halign="left")
center_s <- createStyle(fontName="Arial",fontSize=11,halign="center")
ref_s    <- createStyle(fontName="Arial",fontSize=11,
                        fontColour="#808080",halign="center")
bot_s    <- createStyle(fontName="Arial",fontSize=11,
                        border="Bottom",borderStyle="medium")
note_s   <- createStyle(fontName="Arial",fontSize=9,
                        wrapText=TRUE,valign="top")

write_table_sheet <- function(wb, sname, title, tbl,
                              grp_id_col, footnote, ref_col=NULL) {
  n <- ncol(tbl)
  addWorksheet(wb, sname)
  writeData(wb, sname, title, startRow=1, startCol=1)
  addStyle(wb, sname, title_s, rows=1, cols=1)
  writeData(wb, sname, as.data.frame(t(names(tbl))),
            startRow=2, startCol=1, colNames=FALSE)
  addStyle(wb, sname, header_s, rows=2, cols=1:n, gridExpand=TRUE)
  grp_labels <- tbl[[1]][is.na(tbl[[grp_id_col]]) &
                          !str_starts(tbl[[1]],"   ")]
  for (i in seq_len(nrow(tbl))) {
    re     <- i + 2
    cv     <- tbl[[1]][i]
    is_grp <- cv %in% grp_labels
    is_ref <- !is.null(ref_col) && !is.na(tbl[[ref_col]][i]) &&
               tbl[[ref_col]][i] == "Reference"
    writeData(wb, sname, tbl[i,], startRow=re, startCol=1, colNames=FALSE)
    if (is_grp) {
      addStyle(wb, sname, group_s, rows=re, cols=1:n, gridExpand=TRUE)
    } else {
      addStyle(wb, sname, data_s, rows=re, cols=1)
      addStyle(wb, sname, if(is_ref) ref_s else center_s,
               rows=re, cols=2:n, gridExpand=TRUE)
    }
  }
  lr <- nrow(tbl) + 2
  addStyle(wb, sname, bot_s, rows=lr, cols=1:n, gridExpand=TRUE)
  setColWidths(wb, sname, cols=1, widths=38)
  setColWidths(wb, sname, cols=2:n, widths=20)
  fn <- lr + 2
  writeData(wb, sname, footnote, startRow=fn, startCol=1)
  mergeCells(wb, sname, cols=1:n, rows=fn)
  addStyle(wb, sname, note_s, rows=fn, cols=1)
  setRowHeights(wb, sname, rows=fn, heights=70)
}

# ── Sheet 1 ──────────────────────────────────────────────────────────────────
write_table_sheet(wb, "Sheet 1",
  paste0("National Estimates of Patient and Hospital Characteristics by ",
         "Facial Nerve Repair Status, Including Repair Rates"),
  sheet1, grp_id_col="No Repair (n, %)",
  footnote=paste0(
    "HCUP NASS 2016\u20132023. Total = All patients with traumatic facial nerve ",
    "laceration nationally. Group 1 (Phase 2) = encounters meeting Criterion A ",
    "(S04.5 co-occurring with S01.*, S11.*, or external cause code) OR ",
    "Criterion B (CPT 64864/64865 co-occurring with S01.* or S11.*). ",
    "Group 2 (Repair) = CPT 64864 or 64865 within Group 1. ",
    "Group 3 (No Repair) = Group 1 minus Group 2. All values are weighted ",
    "national estimates using DISCWT discharge weight. Repair Rate = Weighted repair ",
    "encounters / Weighted Group 1 encounters x 100. Age mean \u00b1 SD estimated ",
    "using age group midpoints with discharge weights. NOTE: HOSP_REGION was not ",
    "available in the 2023 NASS Hospital file; regional estimates reflect ",
    "2016\u20132022 data only."))

# ── Sheet 2 ──────────────────────────────────────────────────────────────────
write_table_sheet(wb, "Sheet 2",
  "Multivariable Logistic Regression for Odds of Facial Nerve Repair",
  sheet2, grp_id_col="Adjusted OR (95% CI)",
  ref_col="Adjusted OR (95% CI)",
  footnote=paste0(
    "Survey-weighted multivariable logistic regression (svyglm, quasibinomial). ",
    "HCUP NASS 2016\u20132023. Reference categories: Male, Age 18-34, White, ",
    "76th-100th income quartile, Private insurance, South region, Urban, ",
    "Nonteaching hospital. OR = odds ratio; CI = 95% confidence interval. ",
    "Significant predictors (p<0.05): ", sig_pred_text, "."))

# ── Sheet 4 ──────────────────────────────────────────────────────────────────
addWorksheet(wb, "Sheet 4")
writeData(wb, "Sheet 4", "Repair Rates by Year", startRow=1, startCol=1)
addStyle(wb, "Sheet 4", title_s, rows=1, cols=1)
writeData(wb, "Sheet 4", sheet4, startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sheet 4", header_s, rows=3, cols=1:6, gridExpand=TRUE)
addStyle(wb, "Sheet 4", data_s, rows=4:(nrow(sheet4)+3), cols=1:6, gridExpand=TRUE)
setColWidths(wb, "Sheet 4", cols=1, widths=10)
setColWidths(wb, "Sheet 4", cols=2:6, widths=28)
fn4 <- nrow(sheet4)+5
writeData(wb, "Sheet 4",
  paste0("HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight. ",
       "Repair rate = Weighted repair / Weighted total injury x 100. ",
       "Chi-square test for difference in repair rates across years: p = ", year_p, "."),
  startRow=fn4, startCol=1)
mergeCells(wb, "Sheet 4", cols=1:6, rows=fn4)
addStyle(wb, "Sheet 4", note_s, rows=fn4, cols=1)
setRowHeights(wb, "Sheet 4", rows=fn4, heights=40)

# ── Sheet 5 ──────────────────────────────────────────────────────────────────
addWorksheet(wb, "Sheet 5")
writeData(wb, "Sheet 5", "Charge Analysis: Repair vs No Repair",
          startRow=1, startCol=1)
addStyle(wb, "Sheet 5", title_s, rows=1, cols=1)
writeData(wb, "Sheet 5", sheet5, startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sheet 5", header_s, rows=3, cols=1:5, gridExpand=TRUE)
grp5 <- sheet5$Characteristic[is.na(sheet5$`No Repair Median (IQR)`) &
                               !str_starts(sheet5$Characteristic,"   ")]
for (i in seq_len(nrow(sheet5))) {
  re <- i+3
  if (sheet5$Characteristic[i] %in% grp5) {
    addStyle(wb, "Sheet 5", group_s, rows=re, cols=1:5, gridExpand=TRUE)
  } else {
    addStyle(wb, "Sheet 5", data_s,   rows=re, cols=1)
    addStyle(wb, "Sheet 5", center_s, rows=re, cols=2:5, gridExpand=TRUE)
  }
}
addStyle(wb, "Sheet 5", bot_s, rows=nrow(sheet5)+3, cols=1:5, gridExpand=TRUE)
setColWidths(wb, "Sheet 5", cols=1, widths=35)
setColWidths(wb, "Sheet 5", cols=2:5, widths=28)
fn5 <- nrow(sheet5)+5
writeData(wb, "Sheet 5",
  paste0("Median (IQR) weighted national estimates. HCUP NASS 2016\u20132023. ",
         "Insurance type p-values: survey-weighted ANOVA on log-charges within ",
         "no-repair group and within repair group separately. ",
         "Teaching status p-values: survey-weighted t-test on log-charges ",
         "comparing Nonteaching vs Teaching within each repair group."),
  startRow=fn5, startCol=1)
mergeCells(wb, "Sheet 5", cols=1:5, rows=fn5)
addStyle(wb, "Sheet 5", note_s, rows=fn5, cols=1)
setRowHeights(wb, "Sheet 5", rows=fn5, heights=55)

# ── NIS 2023 Sheet ────────────────────────────────────────────────────────────
addWorksheet(wb, "NIS 2023 Inpatient Results")
writeData(wb, "NIS 2023 Inpatient Results",
          "NIS 2023 Facial Nerve Repair \u2014 Inpatient Database Results",
          startRow=1, startCol=1)
addStyle(wb, "NIS 2023 Inpatient Results", title_s, rows=1, cols=1)
mergeCells(wb, "NIS 2023 Inpatient Results", cols=1:4, rows=1)
writeData(wb, "NIS 2023 Inpatient Results", nis_sheet,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "NIS 2023 Inpatient Results", header_s, rows=3, cols=1:4,
         gridExpand=TRUE)
for (i in seq_len(nrow(nis_sheet))) {
  re       <- i + 3
  is_cns   <- i <= 3
  is_total <- i == 7
  bg <- if (is_total) "#EBF4FB" else "white"
  fc <- if (is_cns) "#999999" else "black"

  if (is_total) {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fontColour=fc, fgFill=bg,
                       textDecoration="bold")
  } else {
    sty <- createStyle(fontName="Arial", fontSize=11,
                       fontColour=fc, fgFill=bg)
  }
  addStyle(wb, "NIS 2023 Inpatient Results", sty,
           rows=re, cols=1:4, gridExpand=TRUE)
}
addStyle(wb, "NIS 2023 Inpatient Results", bot_s,
         rows=nrow(nis_sheet)+3, cols=1:4, gridExpand=TRUE)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=1, widths=20)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=2, widths=52)
setColWidths(wb, "NIS 2023 Inpatient Results", cols=3:4, widths=22)
fn_nis <- nrow(nis_sheet)+5
writeData(wb, "NIS 2023 Inpatient Results",
  paste0("Source: HCUP National Inpatient Sample (NIS) 2023. ",
         "Carried forward from primary analysis. ",
         "ICD-10-PCS codes 01QM0ZZ, 01QM3ZZ, 01QM4ZZ searched (PNS section). ",
         "CNS section codes 00QM0/3/4ZZ returned 0 encounters. ",
         "Individual PNS code breakdown suppressed per HCUP cell size restrictions (<11). ",
         "Weighted estimate uses NIS DISCWT discharge weight. ",
         "Finding confirms facial nerve repair is predominantly an ambulatory procedure."),
  startRow=fn_nis, startCol=1)
mergeCells(wb, "NIS 2023 Inpatient Results", cols=1:4, rows=fn_nis)
addStyle(wb, "NIS 2023 Inpatient Results", note_s, rows=fn_nis, cols=1)
setRowHeights(wb, "NIS 2023 Inpatient Results", rows=fn_nis, heights=60)

# ── IPTW Sheet ────────────────────────────────────────────────────────────────
addWorksheet(wb, "IPTW Charges Analysis")
writeData(wb, "IPTW Charges Analysis",
          "IPTW Analysis: Effect of Repair on Total Charges",
          startRow=1, startCol=1)
addStyle(wb, "IPTW Charges Analysis", title_s, rows=1, cols=1)
writeData(wb, "IPTW Charges Analysis", iptw_excel,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "IPTW Charges Analysis", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "IPTW Charges Analysis", data_s,
         rows=4:(nrow(iptw_excel)+3), cols=1:2, gridExpand=TRUE)
setColWidths(wb, "IPTW Charges Analysis", cols=1, widths=40)
setColWidths(wb, "IPTW Charges Analysis", cols=2, widths=60)
fn_iptw <- nrow(iptw_excel)+5
writeData(wb, "IPTW Charges Analysis",
  paste0("IPTW estimated using propensity score weighting (WeightIt, ATE estimand). ",
         "Outcome modeled as log(total charges) via survey-weighted linear regression. ",
         "Effect reflects percentage change in charges associated with repair after ",
         "balancing all measured confounders. HCUP NASS 2016\u20132023."),
  startRow=fn_iptw, startCol=1)
mergeCells(wb, "IPTW Charges Analysis", cols=1:2, rows=fn_iptw)
addStyle(wb, "IPTW Charges Analysis", note_s, rows=fn_iptw, cols=1)
setRowHeights(wb, "IPTW Charges Analysis", rows=fn_iptw, heights=55)

# ── Phase 1 vs Phase 2 Comparison (Excel only — internal QC) ─────────────────
addWorksheet(wb, "Phase 1 vs Phase 2 Comparison")
writeData(wb, "Phase 1 vs Phase 2 Comparison",
          "Phase 1 vs Phase 2 Comparison \u2014 Internal Quality Check",
          startRow=1, startCol=1)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", title_s, rows=1, cols=1)
mergeCells(wb, "Phase 1 vs Phase 2 Comparison", cols=1:3, rows=1)
writeData(wb, "Phase 1 vs Phase 2 Comparison", phase_compare,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", header_s, rows=3, cols=1:3,
         gridExpand=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", data_s,
         rows=4:(nrow(phase_compare)+3), cols=1:3, gridExpand=TRUE)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", bot_s,
         rows=nrow(phase_compare)+3, cols=1:3, gridExpand=TRUE)
setColWidths(wb, "Phase 1 vs Phase 2 Comparison", cols=1, widths=42)
setColWidths(wb, "Phase 1 vs Phase 2 Comparison", cols=2:3, widths=38)
fn_pc <- nrow(phase_compare)+5
writeData(wb, "Phase 1 vs Phase 2 Comparison",
  paste0("Investigator quality check only \u2014 does NOT appear in the Phase 2 ",
         "manuscript. Confirms Phase 2 produces a more complete cohort and that ",
         "the direction of significant predictors is consistent across definitions. ",
         "Phase 2 overall repair rate is expected to be lower than 29.2%; if higher, ",
         "review extraction logic. NOTE: HOSP_REGION not available in the 2023 NASS ",
         "Hospital file; regional ORs reflect 2016\u20132022 data only."),
  startRow=fn_pc, startCol=1)
mergeCells(wb, "Phase 1 vs Phase 2 Comparison", cols=1:3, rows=fn_pc)
addStyle(wb, "Phase 1 vs Phase 2 Comparison", note_s, rows=fn_pc, cols=1)
setRowHeights(wb, "Phase 1 vs Phase 2 Comparison", rows=fn_pc, heights=70)

# ── Sensitivity Analysis 1 ───────────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 1")
writeData(wb, "Sensitivity Analysis 1",
          "Sensitivity Analysis 1 \u2014 Criterion A Only",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 1", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 1", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 1", sens1_tbl,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 1", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 1", data_s,
         rows=4:(nrow(sens1_tbl)+3), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 1", bot_s,
         rows=nrow(sens1_tbl)+3, cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 1", cols=1, widths=32)
setColWidths(wb, "Sensitivity Analysis 1", cols=2, widths=22)
fn_s1 <- nrow(sens1_tbl)+5
writeData(wb, "Sensitivity Analysis 1",
  paste0("Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, ",
         "excluding the external-cause-only path AND excluding the Criterion B ",
         "procedural pathway. HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."),
  startRow=fn_s1, startCol=1)
mergeCells(wb, "Sensitivity Analysis 1", cols=1:2, rows=fn_s1)
addStyle(wb, "Sensitivity Analysis 1", note_s, rows=fn_s1, cols=1)
setRowHeights(wb, "Sensitivity Analysis 1", rows=fn_s1, heights=55)

# ── Sensitivity Analysis 2 ───────────────────────────────────────────────────
addWorksheet(wb, "Sensitivity Analysis 2")
writeData(wb, "Sensitivity Analysis 2",
          "Sensitivity Analysis 2 \u2014 Highest Mechanistic Specificity",
          startRow=1, startCol=1)
addStyle(wb, "Sensitivity Analysis 2", title_s, rows=1, cols=1)
mergeCells(wb, "Sensitivity Analysis 2", cols=1:2, rows=1)
writeData(wb, "Sensitivity Analysis 2", sens2_tbl,
          startRow=3, startCol=1, colNames=TRUE)
addStyle(wb, "Sensitivity Analysis 2", header_s, rows=3, cols=1:2,
         gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 2", data_s,
         rows=4:(nrow(sens2_tbl)+3), cols=1:2, gridExpand=TRUE)
addStyle(wb, "Sensitivity Analysis 2", bot_s,
         rows=nrow(sens2_tbl)+3, cols=1:2, gridExpand=TRUE)
setColWidths(wb, "Sensitivity Analysis 2", cols=1, widths=32)
setColWidths(wb, "Sensitivity Analysis 2", cols=2, widths=22)
fn_s2 <- nrow(sens2_tbl)+5
writeData(wb, "Sensitivity Analysis 2",
  paste0("Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
         "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
         "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight."),
  startRow=fn_s2, startCol=1)
mergeCells(wb, "Sensitivity Analysis 2", cols=1:2, rows=fn_s2)
addStyle(wb, "Sensitivity Analysis 2", note_s, rows=fn_s2, cols=1)
setRowHeights(wb, "Sensitivity Analysis 2", rows=fn_s2, heights=55)

# ── Save Excel ────────────────────────────────────────────────────────────────
xlsx_path <- file.path(xlsx_dir, "NASS Laceration Phase 2 Output 1.0.xlsx")
saveWorkbook(wb, file=xlsx_path, overwrite=TRUE)
message("Excel saved: ", xlsx_path)
```

# GT Tables for Manuscript

```{r gt-export}
#| cache: false

save_gt_to_word <- function(gt_tbl, filename) {
  filepath <- file.path(word_dir, filename)
  gt::gtsave(gt_tbl, filename=filepath)
  message("Saved: ", filename)
}

# ── Sheet 1 gt ───────────────────────────────────────────────────────────────
gt_sheet1 <- sheet1 %>%
  setNames(c("Characteristic","No Repair","Repair","Total","Repair Rate")) %>%
  mutate(across(c(`No Repair`,`Repair`,`Total`,`Repair Rate`),
                ~ifelse(is.na(.), "", .))) %>%
  gt() %>%
  tab_header(title=md("**National Estimates of Patient and Hospital Characteristics by Facial Nerve Repair Status, Including Repair Rates**")) %>%
  tab_footnote(footnote=md(paste0(
    "HCUP NASS 2016\u20132023. Total = All patients with traumatic facial nerve ",
    "laceration nationally. All values are weighted national estimates (DISCWT). ",
    "Repair Rate = Weighted repair / Weighted Group 1 x 100. Age mean \u00b1 SD ",
    "estimated using age group midpoints. NOTE: HOSP_REGION not available in 2023 ",
    "NASS Hospital file; regional estimates reflect 2016\u20132022 only."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_column_labels()) %>%
  tab_style(
    style=cell_text(weight="bold"),
    locations=cells_body(
      rows=(`No Repair`=="" & !str_starts(Characteristic,"   ")) |
            Characteristic=="Total cases" |
            Characteristic=="Age, mean \u00b1 SD"
    )
  ) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic="Characteristic",
             `No Repair`="No Repair (n, %)",
             Repair="Repair (n, %)",
             Total="Total (n)",
             `Repair Rate`="Repair Rate (%)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 2 gt ───────────────────────────────────────────────────────────────
gt_sheet2 <- sheet2 %>%
  setNames(c("Characteristic","Adjusted OR","p value")) %>%
  gt() %>%
  tab_header(title=md("**Multivariable Logistic Regression for Odds of Facial Nerve Repair**")) %>%
  tab_footnote(footnote=md(paste0(
    "Survey-weighted multivariable logistic regression (svyglm, quasibinomial). ",
    "HCUP NASS 2016\u20132023. Reference: Male, Age 18\u201334, White, ",
    "76th\u2013100th income quartile, Private insurance, South, Urban, Nonteaching. ",
    "OR = odds ratio; CI = 95% confidence interval. Significant predictors (p<0.05): ",
    sig_pred_text, "."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_body(
              rows=is.na(`Adjusted OR`) | `Adjusted OR`=="" )) %>%
  tab_style(style=cell_text(color="grey", style="italic"),
            locations=cells_body(
              rows=!is.na(`Adjusted OR`) & `Adjusted OR`=="Reference")) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic="Characteristic",
             `Adjusted OR`="Adjusted OR (95% CI)",
             `p value`="p value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 4 gt ───────────────────────────────────────────────────────────────
gt_sheet4 <- sheet4 %>%
  setNames(c("Year","Injury Unweighted N","Injury Weighted N",
             "Repair Unweighted N","Repair Weighted N","Repair Rate (%)")) %>%
  gt() %>%
  tab_header(title=md("**Repair Rates by Year**")) %>%
  tab_footnote(footnote=md(paste0("HCUP NASS 2016\u20132023. Weighted estimates use DISCWT discharge weight. ",
       "Repair rate = Weighted repair / Weighted total injury x 100. ",
       "Chi-square test for difference in repair rates across years: p = ", year_p, "."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  fmt_number(columns=c(`Injury Weighted N`,`Repair Weighted N`),
             decimals=0, use_seps=TRUE) %>%
  cols_label(`Injury Unweighted N`="Traumatic facial nerve laceration (unweighted N)",
             `Injury Weighted N`  ="Traumatic facial nerve laceration (weighted N)",
             `Repair Unweighted N`="Repair (unweighted N)",
             `Repair Weighted N`  ="Repair (weighted N)",
             `Repair Rate (%)`    ="Repair rate \u2014 weighted (%)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sheet 5 gt ───────────────────────────────────────────────────────────────
gt_sheet5 <- sheet5 %>%
  setNames(c("Characteristic","No Repair Median","Repair Median",
             "p value No Repair","p value Repair")) %>%
  gt() %>%
  tab_header(title=md("**Charge Analysis: Repair vs No Repair**")) %>%
  tab_footnote(footnote=md(paste0(
    "Median (IQR) weighted national estimates. HCUP NASS 2016\u20132023. ",
    "Insurance type p-values: survey-weighted ANOVA on log-charges within ",
    "no-repair group and within repair group separately. Teaching status p-values: ",
    "survey-weighted t-test comparing Nonteaching vs Teaching within each group."))) %>%
  tab_style(style=cell_text(weight="bold"),
            locations=cells_body(
              rows=is.na(`No Repair Median`) &
                   !str_starts(Characteristic,"   "))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Characteristic     ="Characteristic",
             `No Repair Median` ="No Repair Median (IQR)",
             `Repair Median`    ="Repair Median (IQR)",
             `p value No Repair`="p value (no repair: within-group)",
             `p value Repair`   ="p value (repair: within-group)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── NIS 2023 gt ──────────────────────────────────────────────────────────────
gt_nis <- nis_sheet %>%
  setNames(c("ICD-10-PCS Code","Description","Unweighted N","Weighted N")) %>%
  gt() %>%
  tab_header(
    title=md("**NIS 2023 Facial Nerve Repair \u2014 Inpatient Database Results**"),
    subtitle=md("*HCUP National Inpatient Sample, 2023*")) %>%
  tab_footnote(footnote=md(paste0(
    "ICD-10-PCS codes searched in PNS section (01QM0/3/4ZZ) and CNS section ",
    "(00QM0/3/4ZZ). Individual PNS code breakdown suppressed per HCUP cell size ",
    "restrictions (<11 per cell). Weighted estimate uses NIS DISCWT. Finding ",
    "confirms facial nerve repair is predominantly ambulatory; NASS is the ",
    "appropriate primary database."))) %>%
  tab_style(style=list(cell_text(color="grey", style="italic")),
            locations=cells_body(rows=1:3)) %>%
  tab_style(style=list(cell_text(weight="bold"),
                       cell_fill(color="#EBF4FB")),
            locations=cells_body(rows=7)) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  sub_missing(missing_text="\u2014") %>%
  cols_label(`ICD-10-PCS Code`="ICD-10-PCS Code",
             Description="Description",
             `Unweighted N`="Unweighted N",
             `Weighted N`="Weighted N (national estimate)") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sensitivity Analysis 1 gt ────────────────────────────────────────────────
gt_sens1 <- sens1_tbl %>%
  setNames(c("Metric","Value")) %>%
  gt() %>%
  tab_header(title=md("**Sensitivity Analysis 1 \u2014 Criterion A Only**")) %>%
  tab_footnote(footnote=md(paste0(
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.*, excluding ",
    "external-cause-only entries and excluding the Criterion B procedural ",
    "pathway. HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Metric="Metric", Value="Value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Sensitivity Analysis 2 gt ────────────────────────────────────────────────
gt_sens2 <- sens2_tbl %>%
  setNames(c("Metric","Value")) %>%
  gt() %>%
  tab_header(title=md("**Sensitivity Analysis 2 \u2014 Highest Mechanistic Specificity**")) %>%
  tab_footnote(footnote=md(paste0(
    "Group 1 restricted to S04.5 co-occurring with S01.* or S11.* AND an ",
    "external cause code (W25\u2013W27, W45, X99, Y28, V01\u2013V99). ",
    "HCUP NASS 2016\u20132023. Weighted estimates use DISCWT."))) %>%
  tab_style(style=cell_fill(color="#EBF4FB"),
            locations=cells_column_labels()) %>%
  cols_label(Metric="Metric", Value="Value") %>%
  opt_table_font(font="Arial") %>%
  tab_options(table.font.size=11, heading.title.font.size=12,
              column_labels.font.weight="bold",
              table.border.top.style="solid",
              table.border.bottom.style="solid")

# ── Save all Word tables ──────────────────────────────────────────────────────
save_gt_to_word(gt_sheet1, "Table1_Patient_Characteristics.docx")
save_gt_to_word(gt_sheet2, "Table2_Logistic_Regression.docx")
save_gt_to_word(gt_sheet4, "Table4_Repair_Rates_by_Year.docx")
save_gt_to_word(gt_sheet5, "Table5_Charges_Analysis.docx")
save_gt_to_word(gt_nis,    "Table_NIS2023_Inpatient_Results.docx")
save_gt_to_word(gt_sens1,  "Table_Sensitivity_Analysis_1.docx")
save_gt_to_word(gt_sens2,  "Table_Sensitivity_Analysis_2.docx")

message("\nAll Word tables saved to: ", word_dir)
```

# Draft Methods Paragraph

```{r draft-methods}
#| cache: false
g1_unw_fmt   <- format(nrow(enc_all),                  big.mark=",")
g1_w_fmt     <- format(round(total_g1_w),              big.mark=",")
g2_w_fmt     <- format(round(total_repair_w),          big.mark=",")
g3_w_fmt     <- format(round(total_norepair_w),        big.mark=",")
rate_pct     <- round(100*total_repair_w/total_g1_w,   1)
norate_pct   <- round(100*total_norepair_w/total_g1_w, 1)

draft_methods <- paste0(
  "We analyzed the HCUP National Ambulatory Surgery Sample (NASS) for 2016 ",
  "through 2023. A total of ", g1_w_fmt, " encounters with traumatic facial ",
  "nerve laceration were identified nationally (unweighted N = ", g1_unw_fmt,
  "), of whom ", g2_w_fmt, " (", rate_pct, "%) underwent surgical repair ",
  "(CPT 64864 or 64865). ", g3_w_fmt, " patients (", norate_pct,
  "%) were managed without operative repair."
)

cat(draft_methods)
```

# Copy Outputs to OneDrive

```{r copy-to-drive}
#| cache: false
# Copy the rendered HTML and the .qmd source into the Quarto Versions subfolder.
qmd_basename  <- "nass_phase2_analysis_1_0.qmd"
html_basename <- "nass_phase2_analysis_1_0.html"

# HTML: search common locations for the rendered output
html_candidates <- c(
  file.path(getwd(),     html_basename),
  file.path("~/Downloads", html_basename),
  file.path("~/Documents", html_basename)
)
html_src <- html_candidates[file.exists(path.expand(html_candidates))][1]
if (!is.na(html_src)) {
  file.copy(path.expand(html_src),
            file.path(qmd_dir, html_basename),
            overwrite=TRUE)
  message("HTML copied from: ", html_src)
} else {
  message("HTML not yet available; render the document and re-run this chunk.")
}

# QMD source: copy from the working directory if present
qmd_src <- file.path(getwd(), "nass_phase2_analysis_1_0.qmd")
qmd_dst <- file.path(qmd_dir, "nass_phase2_analysis_1_0.qmd")
if (file.exists(qmd_src) && normalizePath(qmd_src) != normalizePath(qmd_dst)) {
  file.copy(qmd_src, qmd_dst, overwrite=TRUE)
  message("QMD copied to Quarto Versions: ", qmd_dir)
} else {
  message("QMD is already in Quarto Versions — no copy needed.")
}

message("\nAll outputs saved under: ", output_dir)
message("  Figures      : ", fig_dir)
message("  Excel        : ", xlsx_dir)
message("  Word Tables  : ", word_dir)
message("  Quarto       : ", qmd_dir)
```

# Session Info

```{r session}
sessionInfo()
```
